---
title: "Archived Soil Incubations Manuscript Figures"
author: "J. Beem-Miller"
date: "08 Oct 2020"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
  html_notebook:
    css: "custom.css"
    toc: yes
    toc_depth: 2
  pdf_document:
    latex_engine: xelatex
header_includes:
- \usepackage[utf8]{inputenc}
- \usepackage{float}
---
```{r global_options, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.align = 'center', dev = c('cairo_pdf', 'png'), fig.width = 6.5)
```

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(SoilR)
library(corrplot)
library(openxlsx)
library(ISRaD)
library(utilities)
```
## Notes
* This workbook is intended to load and prepare the key data for analysis for the archive incubation project.
* This is an updated version of RMD file "./src/arc_inc_data-wrangling_2020-04-20.Rmd" with the explanatory text removed
* all code chunk options are set to "echo = FALSE"; see raw .Rmd file for data wrangling code.

# Methods and materials

*Radiocarbon corrections*

Although differences were small, radiocarbon data for archived samples were first corrected for decay since the year of collection before assessing treatment effects.

Decay correction formula:
$$1000 \cdot \left( (FM \cdot e^{\frac{-year_{sampled} + 1950}{8267}}) - 1 \right)$$

where FM is the fraction modern, and 8267 is the inverse of the product of the natural log of two and the true half life of ^14^C (5730 y).

```{r 14C helper fxs}
# 5. Create helper functions
# decay correction function  
decay.corr.fx <- function(year_sampled, fraction_modern) {
  ifelse(is.na(fraction_modern), NA, 1000*((fraction_modern*exp((-year_sampled + 1950)/8267))-1))
}

# d14C to fraction modern function
d14c.fm.fx <- function(year_sampled, d14c) {
  ((d14c / 1000) + 1) / exp(0.00012097 * (1950 - year_sampled))
  }

# define helper function for stripping out "NA" and non-Exploratories data
extra.clean.fx <- function(df) {
  df <- df[which(grepl("HEG", df$Probe) |
                   grepl("HEW", df$Probe) |
                   grepl("SEG", df$Probe) |
                   grepl("SEW", df$Probe)), ]
  if(length(grep("Experiment", names(df))) > 0) {
     df <- df[which(df$Experiment != "tme"), ] 
  }
  ix <- which(is.na(df$F14C) & is.na(df$`X.14C....`))
  if(length(ix) > 1) {
    df <- df[-ix, ]
  }
  return(df)
}

# extract storage duration data separately
tme.14c.fx <- function(df) {
  df$SampleName <- df$Probe
  df$Period <- "inc"
  if(length(grep("Experiment", names(df))) > 0) {
     df <- df[which(df$Experiment == "tme"), ] 
  }
  return(df)
}
```

```{r read wide ts data}
# 1. Load flux data from air-dry + storage (experiment 1) and storage duration (experiment 3) control samples, and convert from "wide" to "long" format so as to match other data. 
# Read air-dry + storage respiration control data
archive.ctl.ts.wide <- read.csv("../data/derived/arc-storage-t1-flux/arc-inc_xplr_arc-storage-t1-flux_2020-07-03.csv")

# Convert time data from wide to long
archive.ctl.ts.d <- select(archive.ctl.ts.wide, starts_with("inc_time_d")) %>%
  pivot_longer(everything(),
               names_to = "timepoint_cmtv",
               names_prefix = "inc_time_d_",
               values_to = "time_d")
# sort by timepoint_cmtv
archive.ctl.ts.d <- data.frame(archive.ctl.ts.d[order(archive.ctl.ts.d$timepoint_cmtv),])

# Convert mgCO2 data from wide to long
archive.ctl.ts.co2 <- select(archive.ctl.ts.wide, starts_with("inc_mgCO2_jar_")) %>%
  pivot_longer(everything(),
               names_to = "timepoint_cmtv",
               names_prefix = "inc_mgCO2_jar_",
               values_to = "mgCO2_jar")
# sort by timepoint_cmtv
archive.ctl.ts.co2 <- archive.ctl.ts.co2[order(archive.ctl.ts.co2$timepoint_cmtv),]

# assemble in long dataframe 
archive.ctl.ts <- data.frame(ID = rep(unique(archive.ctl.ts.wide$ID), 4),
                             Period = "inc",
                             timepoint_cmtv = rep(seq(from = 2, 
                                                      to = 5), 
                                                  each = length(archive.ctl.ts.wide$ID)),
                             time_d = archive.ctl.ts.d$time_d,
                             mgCO2_jar = archive.ctl.ts.co2$mgCO2_jar)

# add preincubation data
archive.ctl.ts.pre <- archive.ctl.ts.wide[1:length(unique(archive.ctl.ts.wide$ID)),
                                          c("ID", "preinc_time_d", "preinc_mgCO2_jar")]
archive.ctl.ts.pre <- rename(archive.ctl.ts.pre, time_d = preinc_time_d, mgCO2_jar = preinc_mgCO2_jar)
archive.ctl.ts.pre$timepoint_cmtv <- 1
archive.ctl.ts.pre$Period <- "pre"
archive.ctl.ts.pre$time_d_cmtv <- archive.ctl.ts.pre$time_d
  
# Calculate time_d_cmtv
archive.ctl.ts <- bind_rows(
  lapply(
    split(archive.ctl.ts, archive.ctl.ts$timepoint_cmtv),
    function(x) {
      x$time_d_cmtv <- x$time_d + 
        archive.ctl.ts.pre[match(x$ID, archive.ctl.ts.pre$ID), "time_d"]
      return(x)
    }
    )
  )

# rbind preinc and inc data
archive.ctl.ts <- rbind(archive.ctl.ts.pre, archive.ctl.ts)

# add SampleName column and restore Experiment column
archive.ctl.ts$SampleName <- archive.ctl.ts$ID
archive.ctl.ts$Experiment <- archive.ctl.ts.wide[match(archive.ctl.ts$ID, archive.ctl.ts.wide$ID), "Experiment"]

# remove intermediate files
rm(archive.ctl.ts.d, archive.ctl.ts.co2, archive.ctl.ts.pre)
```

```{r read-long-ts-data}
# 2. Load flux data from air-dry + storage samples and from air-dry experiment (ctl & treatment), C & N data for all the Exploratories samples (measured in 2011), and soil mass and moisture data for all experiments.
# Read Xplr C and N data
xplr.cn <- read.csv("../data/external/exploratories_ischoening_soilcn_2011/170407_CN_concentrations_Exploratories_EP_Plots_2011.csv")

# Read Xplr 14C data for control samples ("tme" and "arc" Experiments)
xplr.ctl.iso <- read.xlsx("../data/external/exploratories_ischoening_14c-co2_2011/170228_14C_Respiration.xlsx")

# Read soil mass and moisture info for air-dry + storage treatment samples
jarinfo.arc <- read.csv("../data/derived/arc-storage-t2-jarInfo/jar_information.csv")

# Read soil mass and moisture info for air-dry data
jarinfo.rewet <- read.csv("../data/derived/air-dry-2019-whc/air-dry-2019-whc_2020-07-03.csv")

# Read soil mass and moisture info for storage duration treatment samples
# remove unit row and convert to appropriate types
jarinfo.tme <- type.convert(read.xlsx("../data/raw/lab_jena_results-co2_2018-07-30/ArcInc_IncData_2018.xlsx", sheet = 1)[-1, ])

# Read Duke iso and flux data
tme.ctl.duk <- read.csv("../data/derived/arc-tme-duke/arc-tme-duke_2020-04-29.csv")

# Read Sierra Nevada (Jun's) soil data
tme.ctl.soil.sra <- read.csv("../data/derived/arc-tme-sierra/arc-tme-sierra-soil_2020-04-29.csv")

# Read Oak Ridge C data
tme.ctl.ornl <- read.csv("../data/derived/arc-tme-ornl/arc-tme-ornl_2020-04-30.csv")

# Read air-dry + storage respiration data
archive.arc.ts <- read.csv("../data/derived/arc-storage-t2-flux/archive-flux-t2_2020-04-24.csv")

# Read control respiration data from 2019 air-dry experiment
rewet.ctl.ts <- read.csv("../data/derived/fresh-2019-flux/fresh-2019-flux_2020-04-24.csv")

# Read treatment respiration data from 2019 air-dry experiment
rewet.dry.ts <- read.csv("../data/derived/dried-2019-flux/dried-2019-flux_2020-04-24.csv")

# Read Sierra Nevada (Jun's) flux data
tme.ctl.flx.sra <- read.csv("../data/derived/arc-tme-sierra/arc-tme-sierra-flux_2020-07-30.csv")

# Read storage duration respiration data
tme.trt.ts <- rbind(read.csv("../data/derived/arc-tme-t2-flux/arc-tme-t2-flux-1_2020-04-27.csv", stringsAsFactors = FALSE),
                        read.csv("../data/derived/arc-tme-t2-flux/arc-tme-t2-flux-2_2020-04-27.csv", stringsAsFactors = FALSE))

# Read miscellaneous storage treatment data
tme.misc <- read.csv("../data/derived/arc-tme-misc/arc-tme-misc-iso_2020-04-29.csv")
save(tme.misc, file = "tme.misc.RData")
```

```{r summarize data, include = FALSE}
# 3. Combine and summarize data in long format to calculate respiration rates and plot over time
# NB: set "include" to false to suppress bind_rows warning about coercing factor to char
# remove formula_check columns
invisible(list2env(lapply(list(archive.arc.ts = archive.arc.ts, 
                               rewet.ctl.ts = rewet.ctl.ts, 
                               rewet.dry.ts = rewet.dry.ts,
                               tme.trt.ts = tme.trt.ts),
                          function(x) {
                            x$formula_check <- NULL
                            return(x)
                            }),
                   envir = .GlobalEnv)
          )

# Add Treatment columns to each dataset
archive.arc.ts$Treatment <- "air-dry + storage"
archive.ctl.ts$Treatment <- "control"
tme.trt.ts$Treatment <- "storage duration"
rewet.ctl.ts$Treatment <- "control"
rewet.dry.ts$Treatment <- "air-dry"

# combine rewet data
rewet.ts <- rbind(rewet.ctl.ts, rewet.dry.ts)
rewet.ts$Experiment <- "rewet"

# add soil dry wt data
archive.arc.ts$dw_g <- jarinfo.arc[match(archive.arc.ts$SampleName, jarinfo.arc$Sample), "Soil.dry.weight_.g."]
archive.ctl.ts$dw_g <- archive.ctl.ts.wide[match(archive.ctl.ts$ID, archive.ctl.ts.wide$ID), "dry_soil_wt_corrected_g"]
rewet.ts <- merge(rewet.ts, jarinfo.rewet[ , c("ID", "Treatment", "dw_g")], by = c("ID", "Treatment"))
tme.trt.ts <- merge(tme.trt.ts, jarinfo.tme[ , c("Nr.", "dry.soil")], by.x = "SampleName", by.y = "Nr.")
colnames(tme.trt.ts)[which(names(tme.trt.ts) == "dry.soil")] <- "dw_g"

# add experiment col to archive.arc.ts, storage.trt.ts
archive.arc.ts$Experiment <- "arc"
tme.trt.ts$Experiment <- "tme"

# combine archive data and Exploratories storage duration data
archive.ts <- rbind(archive.ctl.ts[which(archive.ctl.ts$Experiment != "tme"), ], 
                    archive.arc.ts[ , names(archive.ctl.ts)])
tme.ts <- rbind(archive.ctl.ts[which(archive.ctl.ts$Experiment == "tme"), ],
                tme.trt.ts[ , names(archive.ctl.ts)])

# bind tme.ctl.flx.sra data to tme.ts
# all other tme.flx data is already summarized
tme.ctl.flx.sra$Experiment <- "tme"
tme.ctl.flx.sra$Treatment <- "control"
tme.ctl.flx.sra$dw_g <- tme.ctl.soil.sra[match(tme.ctl.flx.sra$Jar_ID, tme.ctl.soil.sra$Jar_ID), "dw_g"]
tme.ts <- rbind(tme.ts, tme.ctl.flx.sra[ , names(tme.ts)])

# Test that required names are present
nms <- c("SampleName", "ID", "Experiment", "Treatment", "Period", "timepoint_cmtv",  "time_d", "time_d_cmtv", "mgCO2_jar", "dw_g")
invisible(lapply(list(archive.ts = archive.ts,
                      rewet.ts = rewet.ts,
                      tme.ts),
       function(x) {
         ifelse(!is.na(match(nms, names(x))), "yes", "no")
       }
       ))

# combine all data and remove time points without CO2 measurements
ts <- bind_rows(archive.ts[ , nms], rewet.ts[ , nms], tme.ts[ , nms])
if(length(which(is.na(ts$mgCO2_jar))) > 0) {
  ts <- ts[-which(is.na(ts$mgCO2_jar)), ]
}

# add C content
# xplr
ts$gC_gS <- xplr.cn[match(ts$ID, xplr.cn$EP_Plotid), "Total_C"]*10^-3
# duke
ts[which(is.na(ts$gC_gS)), "gC_gS"] <- tme.ctl.duk[match(ts[which(is.na(ts$gC_gS)), "ID"], tme.ctl.duk$ID), "C_g_kg"]*10^-3
# sierras
ts[which(is.na(ts$gC_gS)), "gC_gS"] <- tme.ctl.soil.sra[match(ts[which(is.na(ts$gC_gS)), "ID"], tme.ctl.soil.sra$ID), "gC_kgS"]*10^-3
# oak ridge
ts[which(is.na(ts$gC_gS)), "gC_gS"] <- tme.ctl.ornl[match(ts[which(is.na(ts$gC_gS)), "ID"], tme.ctl.ornl$ID), "C_g_kg"] * 10^-3

# add Type
ts$Type <- ifelse(grepl("G", ts$ID), "G", "F")

# calculate CO2 respiration in mgCO2-C per g soil C
# NB: 12/44 = mass ratio of C in CO2
ts$mgCO2.C_gC <- (ts$mgCO2_jar*(12/44))/(ts$dw_g*ts$gC_gS)
ts$mgCO2.C_gC_d <- ts$mgCO2.C_gC/ts$time_d
# summarize CO2 release data for each timepoint, by experiment, treatment, and type
ts.avg <- data.frame(na.omit(ts) %>%
  select(Experiment, Treatment, Type, timepoint_cmtv, mgCO2.C_gC_d, time_d_cmtv)  %>%
  group_by(Experiment, Treatment, Type, timepoint_cmtv) %>%
  add_tally() %>%
  summarise_all(list(mean = mean, sd = sd), na.rm = TRUE) %>%
  mutate(se_slope = mgCO2.C_gC_d_sd/sqrt(n_mean),
         se_slope_l = mgCO2.C_gC_d_mean - se_slope,
         se_slope_u = mgCO2.C_gC_d_mean + se_slope))
# add timepoint zero data
tp0.f <- function(df) {
  tp0 <- df[1, ]
  tp0[ , -which(names(tp0) == "Experiment" | names(tp0) == "Treatment" | names(tp0) == "Type" | names(tp0) == "n")] <- 0
  rbind(tp0, df)
}
ts.avg$treat.Exp <- paste(ts.avg$Treatment, ts.avg$Experiment)
ts.avg <- lapply(split(ts.avg, ts.avg$treat.Exp), function(x) {
  lapply(split(x, x$Type), tp0.f)
  })
ts.avg <- bind_rows(lapply(ts.avg, bind_rows))
# restore classes
ts.avg <- as.data.frame(lapply(ts.avg, type.convert))
# save object
save(ts.avg, file = "./ts.avg.RData")
# average
ts.pre <- ts[ts$Period == "pre", ]
ts.pre.avg <- ts.pre %>%
  filter(time_d_cmtv > 3.5) %>%
  select(Experiment, Treatment, Type, timepoint_cmtv, SampleName, mgCO2.C_gC, time_d_cmtv)  %>%
  group_by(Experiment, Treatment, Type, timepoint_cmtv, SampleName) %>%
  summarise_all(list(mean = mean), na.rm = TRUE) %>%
  mutate(mgCO2.C_gC_d = mgCO2.C_gC_mean/time_d_cmtv_mean) %>%
  select(-SampleName, -mgCO2.C_gC_mean, Experiment, Treatment, Type, mgCO2.C_gC_d, time_d_cmtv_mean) %>%
  group_by(Experiment, Treatment, Type) %>%
  add_tally() %>%
  summarise_all(list(mean = mean, sd = sd), na.rm = TRUE) %>%
  mutate(se_slope = mgCO2.C_gC_d_sd/sqrt(n_mean),
         se_slope_l = mgCO2.C_gC_d_mean - se_slope,
         se_slope_u = mgCO2.C_gC_d_mean + se_slope)
ts.inc.avg <- ts %>%
  filter(Period == "inc") %>%
  select(Experiment, Treatment, Type, timepoint_cmtv, mgCO2.C_gC_d, time_d_cmtv)  %>%
  group_by(Experiment, Treatment, Type, timepoint_cmtv) %>%
  add_tally() %>%
  summarise_all(list(mean = mean, sd = sd), na.rm = TRUE) %>%
  mutate(se_slope = mgCO2.C_gC_d_sd/sqrt(n_mean),
         se_slope_l = mgCO2.C_gC_d_mean - se_slope,
         se_slope_u = mgCO2.C_gC_d_mean + se_slope)
colnames(ts.pre.avg)[which(names(ts.pre.avg) %in% c("timepoint_cmtv_mean",
                       "time_d_cmtv_mean_mean",
                       "time_d_cmtv_mean_sd"))] <- c("timepoint_cmtv",
                                                    "time_d_cmtv_mean",
                                                    "time_d_cmtv_sd")

ts.pre.avg <- ts.pre.avg[ , -which(!names(ts.pre.avg) %in% names(ts.inc.avg))]

ts.avg2 <- data.frame(rbind(ts.pre.avg, ts.inc.avg))
ts.avg2 <- ts.avg2[complete.cases(ts.avg2), ]

# add timepoint zero data
ts.avg2$treat.Exp <- paste(ts.avg2$Treatment, ts.avg2$Experiment)
ts.avg2 <- lapply(split(ts.avg2, ts.avg2$treat.Exp), function(x) {
  lapply(split(x, x$Type), tp0.f)
  })
ts.avg2 <- bind_rows(lapply(ts.avg2, bind_rows))

# restore classes
ts.avg2 <- as.data.frame(lapply(ts.avg2, type.convert))
save(ts.avg2, file = "./ts.avg2.Rdata")
```

# Results
## Respiration rates

*note that the text below was left because statistics were calculated on the fly*

```{r resp-rate-stats, include = FALSE}
# rewet max resp rates
t.test(
  tapply(ts[ts$Experiment == "rewet" & ts$Treatment == "air-dry" & ts$Period == "pre" & ts$Type == "G", "mgCO2.C_gC_d"], ts[ts$Experiment == "rewet" & ts$Treatment == "air-dry" & ts$Period == "pre" & ts$Type == "G", "ID"], max, na.rm = TRUE),
  tapply(ts[ts$Experiment == "rewet" & ts$Treatment == "air-dry" & ts$Period == "pre" & ts$Type == "F", "mgCO2.C_gC_d"],ts[ts$Experiment == "rewet" & ts$Treatment == "air-dry" & ts$Period == "pre" & ts$Type == "F", "ID"], max, na.rm = TRUE))
### ctl vs. trt pre-inc resp max
## Exp.1 
# G
t.test(
  unlist(
    ts %>%
      filter(Experiment == "rewet" & Treatment == "air-dry" & Period == "pre" & Type == "G") %>%
      group_by(ID) %>%
      summarize(across(mgCO2.C_gC_d, max)) %>%
      select(mgCO2.C_gC_d)),
  unlist(
    ts %>%
      filter(Experiment == "rewet" & Treatment == "control" & Period == "pre" & Type == "G") %>%
      group_by(ID) %>%
      summarize(across(mgCO2.C_gC_d, max)) %>%
      select(mgCO2.C_gC_d)),
  paired = TRUE)
# F
t.test(
  unlist(
    ts %>%
      filter(Experiment == "rewet" & Treatment == "air-dry" & Period == "pre" & Type == "F") %>%
      group_by(ID) %>%
      summarize(across(mgCO2.C_gC_d, max)) %>%
      select(mgCO2.C_gC_d)),
  unlist(
    ts %>%
      filter(Experiment == "rewet" & Treatment == "control" & Period == "pre" & Type == "F") %>%
      group_by(ID) %>%
      summarize(across(mgCO2.C_gC_d, max)) %>%
      select(mgCO2.C_gC_d)),
  paired = TRUE)
## Exp.2 
# G
t.test(
  unlist(
    ts %>%
      filter(Experiment == "rewet" & Treatment == "air-dry" & Period == "pre" & Type == "G") %>%
      group_by(ID) %>%
      summarize(across(mgCO2.C_gC_d, max)) %>%
      select(mgCO2.C_gC_d)),
  unlist(
    ts %>%
      filter(Experiment == "rewet" & Treatment == "control" & Period == "pre" & Type == "G") %>%
      group_by(ID) %>%
      summarize(across(mgCO2.C_gC_d, max)) %>%
      select(mgCO2.C_gC_d)),
  paired = TRUE)
# F
t.test(
  unlist(
    ts %>%
      filter(Experiment == "rewet" & Treatment == "air-dry" & Period == "pre" & Type == "F") %>%
      group_by(ID) %>%
      summarize(across(mgCO2.C_gC_d, max)) %>%
      select(mgCO2.C_gC_d)),
  unlist(
    ts %>%
      filter(Experiment == "rewet" & Treatment == "control" & Period == "pre" & Type == "F") %>%
      group_by(ID) %>%
      summarize(across(mgCO2.C_gC_d, max)) %>%
      select(mgCO2.C_gC_d)),
  paired = TRUE)
## 
```

*Experiment 1 (air-dry + storage treatment)*

Among the air-dry + storage samples, respiration rates were more than twice as high in grassland soils than in forest soils, reaching a maximum of `r {round(max(ts.avg[ts.avg$Experiment == "arc" & ts.avg$Treatment == "air-dry + storage" & ts.avg$Type == "G", "mgCO2.C_gC_d_mean"]),1)}` mg CO~2~ g soil C^-1^ d^-1^ after `r {round(ts.avg[which(ts.avg$mgCO2.C_gC_d_mean == max(ts.avg[ts.avg$Experiment == "arc" & ts.avg$Treatment == "air-dry + storage" & ts.avg$Type == "G", "mgCO2.C_gC_d_mean"])), "time_d_cmtv_mean"],1)}` d, followed by a sharp decline. Mean respiration rates in forest sites peaked at `r {round(max(ts.avg[ts.avg$Experiment == "arc" & ts.avg$Treatment == "air-dry + storage" & ts.avg$Type == "F", "mgCO2.C_gC_d_mean"]),1)}` mg CO~2~ g soil C^-1^ d^-1^ after `r {round(ts.avg[which(ts.avg$mgCO2.C_gC_d_mean == max(ts.avg[ts.avg$Experiment == "arc" & ts.avg$Treatment == "air-dry + storage" & ts.avg$Type == "F", "mgCO2.C_gC_d_mean"])), "time_d_cmtv_mean"],1)}` d, followed by a much more gradual decline than in grassland sites. Control samples responded more weakly and more gradually to rewetting, although as in the treatment samples respiration was greater in grassland soils than in forest soils. Peak respiration rates for control incubations were `r {round(max(ts.avg[ts.avg$Experiment == "arc" & ts.avg$Treatment == "control" & ts.avg$Type == "G", "mgCO2.C_gC_d_mean"]),1)}` and `r {round(max(ts.avg[ts.avg$Experiment == "arc" & ts.avg$Treatment == "control" & ts.avg$Type == "F", "mgCO2.C_gC_d_mean"]),1)}` mg CO~2~ g soil C^-1^ d^-1^ after `r {round(ts.avg[which(ts.avg$mgCO2.C_gC_d_mean == max(ts.avg[ts.avg$Experiment == "arc" & ts.avg$Treatment == "control" & ts.avg$Type == "F", "mgCO2.C_gC_d_mean"])), "time_d_cmtv_mean"],1)}` d for grassland and forest soils, respectively.

*Experiment 2 (air-dry treatment, 2019 samples)*

Peak respiration rates were not significantly different (p > 0.05) between forest and grassland soils in Experiment 2, peaking at `r {round(max(ts.avg2[ts.avg2$Experiment == "rewet" & ts.avg2$Treatment == "air-dry" & ts.avg2$Type == "G", "mgCO2.C_gC_d_mean"]),1)}`.0 and `r {round(max(ts.avg2[ts.avg2$Experiment == "rewet" & ts.avg2$Treatment == "air-dry" & ts.avg2$Type == "F", "mgCO2.C_gC_d_mean"]),1)}` mg CO~2~ g soil C^-1^ d^-1^ after `r {round(ts.avg2[which(ts.avg2$mgCO2.C_gC_d_mean == max(ts.avg2[ts.avg2$Experiment == "rewet" & ts.avg2$Treatment == "air-dry" & ts.avg2$Type == "F", "mgCO2.C_gC_d_mean"])), "time_d_cmtv_mean"]*24,0)}` h for grassland and forest soils, respectively **(Fig. 1)**.

```{r CO2-resp-rates, fig.height=4}
# plotted as measured
ts.avg %>%
  filter(Experiment != "tme") %>%
  ggplot(., aes(time_d_cmtv_mean, mgCO2.C_gC_d_mean)) +
  geom_vline(xintercept = 4, color="gray") +
  geom_ribbon(aes(ymin = se_slope_l, ymax = se_slope_u, fill = Type, linetype = Treatment, alpha = Treatment), show.legend = FALSE) +
  geom_point(aes(color = Type, shape = Treatment), size = 2) +
  geom_line(aes(color = Type, linetype = Treatment)) +
  facet_grid(rows = vars(Experiment),
             labeller = labeller(Experiment = c("arc" = "2011", "rewet" = "2019"))) +
  scale_x_continuous(limits = c(0,18)) +
  scale_color_manual(name = 'Ecosystem',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  scale_fill_manual(name = 'Standard error',
                    values =c('F'='#a35513','G'='#1361a3'),
                    labels = c('Forest','Grassland')) +
  scale_alpha_manual(name = 'Treatment',
                     values = c("air-dry + storage" = .2,
                                "control" = .4,
                                "air-dry" = .2)) +
  scale_linetype_manual(name = 'Treatment',
                        values = c("air-dry + storage" = 'dashed',
                                   "control" ='solid',
                                   "air-dry" = "dotted"),
                        labels = c("air-dry" = "air-dry/rewet",
                                   "air-dry + storage" = "air-dry/rewet + storage",
                                   "control" = "control-1 (2011)\ncontrol-2 (2019)")) +
  scale_shape_manual(name = 'Treatment',
                     values = c("air-dry + storage" = 0,
                                "control" = 16,
                                "air-dry" = 1)) +
  ylab(expression('Respiration Rate (mgCO'[2]*'-C gC'^-1*'d'^-1*')')) +
  xlab("Time (days)") +
  guides(color = guide_legend(order = 1),
         shape = guide_legend(order = 3)) +
  theme_bw() +
  theme(panel.grid = element_blank())
```

>**Fig. `r {(fig.n <- 1)}`. Respiration rates for Experiment 1 (air-dry + storage treatment, 2011) and Experiment 2 (air-dry only treatment, 2019)**

>*Caption:* Top panel shows data from samples collected in 2011 for Experiment 1 (air-dry + storage treatment), bottom panel shows data from samples collected in 2019 for Experiment 2 (air-dry only treatment). Vertical gray line at day 4 demarcates the end of the pre-incubation period and the start of the equilibrium respiration period. Points show measurements and lines show trends in mean respiration rate. Shaded ribbons represent one standard error. The final measurement points for a few samples which took >18 days to reach CO~2~ targets are excluded for display reasons; respiration rates for those samples remained flat. Note that headspace CO~2~ concentrations for Experiment 1 control samples were only measured once during the pre-incubation period (day 4) in contrast to daily measurements for all other samples. Consequently the respiration rate for those samples is the cumulative average rate over the first 4 d.

*Experiment 3 (storage duration)*

Grassland soils from the storage duration treatment group responded rapidly to rewetting and reached target CO~2~ levels after just `r {round(ts.avg2[which(ts.avg2$mgCO2.C_gC_d_mean == max(ts.avg2[ts.avg2$Experiment == "tme" & ts.avg2$Treatment == "storage duration" & ts.avg2$Type == "G", "mgCO2.C_gC_d_mean"])), "time_d_cmtv_mean"]*24,0)}` h of incubation. Only a single observation was made for the grassland treament samples due to the rapid respiration rates, which peaked at `r {round(max(ts.avg2[ts.avg2$Experiment == "tme" & ts.avg2$Treatment == "storage duration" & ts.avg2$Type == "G", "mgCO2.C_gC_d_mean"]),1)}` mg CO~2~ g soil C^-1^ d^-1^. The mean peak respiration rate for forest treatment samples was lower and lagged in comparison to the grassland soils, reaching `r {round(max(ts.avg2[ts.avg2$Experiment == "tme" & ts.avg2$Treatment == "storage duration" & ts.avg2$Type == "F", "mgCO2.C_gC_d_mean"]),1)}` mg CO~2~ g soil C^-1^ d^-1^ after `r {round(ts.avg2[which(ts.avg2$mgCO2.C_gC_d_mean == max(ts.avg2[ts.avg2$Experiment == "tme" & ts.avg2$Treatment == "storage duration" & ts.avg2$Type == "F", "mgCO2.C_gC_d_mean"])), "time_d_cmtv_mean"]*24,0)}` h.

Control-3 respiration rates peaked after the pre-incubation period (`r {round(ts.avg2[which(ts.avg2$mgCO2.C_gC_d_mean == max(ts.avg2[ts.avg2$Experiment == "tme" & ts.avg2$Treatment == "control" & ts.avg2$Type == "G", "mgCO2.C_gC_d_mean"])), "time_d_cmtv_mean"]*24,0)}` h), at `r {round(max(ts.avg2[ts.avg2$Experiment == "tme" & ts.avg2$Treatment == "control" & ts.avg2$Type == "G", "mgCO2.C_gC_d_mean"]),1)}` mg CO~2~ g soil C^-1^ d^-1^. Forest control samples were pre-incubated under various conditions, but respiration rates were only measured during the pre-incubation period for two samples from the Sierra Nevada mountains (USA). In general, respiration rates for forest control samples in Experiment 3 were much lower than in the treatment incubations, peaking at `r {round(max(ts.avg2[ts.avg2$Experiment == "tme" & ts.avg2$Treatment == "control" & ts.avg2$Type == "F", "mgCO2.C_gC_d_mean"]),1)}` mg CO~2~ g soil C^-1^ d^-1^ after `r {round(ts.avg2[which(ts.avg2$mgCO2.C_gC_d_mean == max(ts.avg2[ts.avg2$Experiment == "tme" & ts.avg2$Treatment == "control" & ts.avg2$Type == "F", "mgCO2.C_gC_d_mean"])), "time_d_cmtv_mean"]*24,0)}` h.

## Radiocarbon data

```{r load ams-jena-ingest fx}
# 1. Read in isotope data from various sources
# First load helper functions 'read_jena_ams_results.R', 'read_jena_iso_results.R' 
source("./utilities/jena_ams_ingest.R")
source("./utilities/jena_iso_ingest.R")
```

```{r read iso data, include = FALSE}
# 2. Next read in data from the appropriate directories in 'data/raw'
# 14C
# identify subdirectories in 'raw' directory with "ams_jena" in name
ams_jena_results_dirs <- list.files("../data/raw", pattern = "ams_jena_results", full.names = TRUE)
ams_results_ls <- lapply(seq_along(ams_jena_results_dirs), function(i) {
  read_jena_ams_results(ams_jena_results_dirs[i])
})
names(ams_results_ls) <- list.files("../data/raw", pattern = "ams_jena_results")

# 13C
# identify subdirectories in 'raw' directory with "iso_jena" in name
iso_jena_results_dirs <- list.files("../data/raw", pattern = "iso_jena_results", full.names = TRUE)
iso_results_ls <- lapply(seq_along(iso_jena_results_dirs), function(i) {
  read_jena_iso_results(iso_jena_results_dirs[i])
})
names(iso_results_ls) <- list.files("../data/raw", pattern = "iso_jena_results")
```

```{r arc and rewet experiment template}
# 3. Create a "tidy" style template for the data, i.e. variables in columns. 
# Key variables are as follows:
#   * SampleName (incorporates lab rep and treatment, e.g. "HEG10-1_dry")
#   * ID (plot IDs, e.g. for "HEG10" for Exploratory samples)
#   * Treatment (3 treatments: air-dry, air-dry + storage, storage duration; + controls)
#   * Type (2 levels: F = forest, G = grassland)
#   * Period (incubation period, 2 levels: pre = preincubation, inc = equilibrium incubation)
#   * Experiment (3 levels: arc = air-dry + storage, rewet = air-dry/rewet, time = storage duration)
# 
# * Observational columns include:
#   * d14c ($\Delta$^14^C-CO~2~)
#   * d13c ($\delta$^13^C-CO~2~)
#   * C_g_kg (C content)
#   * dw_g (dry weight)
#   * mgCO2.C_gS (mg CO~2~-C respired g^-1^ soil Period^-1^)
#   * time_d (days in incubation period prior to measurement)
#   * dH2O_grav (percent change in gravimetric water content due to laboratory moisture adjustment)
#   * dH2O_whc (percent change in water holding capacity due to laboratory moisture adjustment)
unique.id.n <- c(length(unique(archive.ts[archive.ts$Treatment == "control", "SampleName"])),
                 length(unique(archive.ts[archive.ts$Treatment == "air-dry + storage", "SampleName"])),
                 length(unique(rewet.ctl.ts$SampleName)),
                 length(unique(rewet.dry.ts$SampleName)))
pre.14c <- data.frame(SampleName = c(as.character(unique(archive.ts[archive.ts$Treatment == "control", "SampleName"])),
                                     as.character(unique(archive.ts[archive.ts$Treatment == "air-dry + storage", "SampleName"])),
                                     as.character(unique(rewet.ctl.ts$SampleName)),
                                     as.character(unique(rewet.dry.ts$SampleName))))
pre.14c$ID <- substr(pre.14c$SampleName, 1, 5)
pre.14c$Type <- ifelse(grepl("G", pre.14c$ID), "G", "F")
pre.14c$Experiment <- c(rep("arc", unique.id.n[1] + unique.id.n[2]),
                       rep("rewet", unique.id.n[3] + unique.id.n[4]))
pre.14c$Treatment <- c(rep("control", unique.id.n[1]),
                       rep("air-dry + storage", unique.id.n[2]),
                       rep("control", unique.id.n[3]),
                       rep("air-dry", unique.id.n[4]))
pre.14c$Period <- "pre"

# create inc period df as duplicate of pre-inc df, but test that it is the right shape first
if(nrow(pre.14c) == sum(unique.id.n) & ncol(pre.14c) == 6) {
  inc.14c <- pre.14c
  inc.14c$Period <- "inc"
} else {
  "pre.14c data frame malformed"
}
all.14c.template <- rbind(pre.14c, inc.14c)
```

```{r tme experiment template}
unique.id.tme.n <- c(length(unique(tme.ctl.duk$SampleName)),
                     length(unique(tme.ctl.ornl$SampleName)),
                     length(unique(tme.ctl.flx.sra$SampleName)),
                     length(unique(tme.misc$SampleName)),
                     length(unique(archive.ctl.ts.wide[archive.ctl.ts.wide$Experiment == "tme", "ID"])),
                     length(unique(tme.trt.ts$SampleName)))
pre.tme.ctl <- data.frame(SampleName = c(as.character(unique(tme.ctl.duk$SampleName)),
                                         as.character(unique(tme.ctl.ornl$SampleName)),
                                         as.character(unique(tme.ctl.flx.sra$SampleName)),
                                         as.character(unique(tme.misc$SampleName)),   as.character(unique(archive.ctl.ts.wide[archive.ctl.ts.wide$Experiment == "tme", "ID"]))))
pre.tme.ctl$Site <- c(rep("Duke", unique.id.tme.n[1]),
                      rep("ORNL", unique.id.tme.n[2]),
                      rep("Sierra", unique.id.tme.n[3]),
                      rep("misc", unique.id.tme.n[4]),
                      rep("Hainich", unique.id.tme.n[5]))


# Duke, ORNL, misc, and Xplr IDs
pre.tme.ctl[pre.tme.ctl$Site == "Duke" | 
            pre.tme.ctl$Site == "ORNL" |
            pre.tme.ctl$Site == "misc" |
            pre.tme.ctl$Site == "Hainich", "ID"] <- as.character(pre.tme.ctl[pre.tme.ctl$Site == "Duke" |
                                                                            pre.tme.ctl$Site == "ORNL" |
                                                                            pre.tme.ctl$Site == "misc" |
                                                                            pre.tme.ctl$Site == "Hainich", "SampleName"])

# Sierra IDs
pre.tme.ctl[pre.tme.ctl$Site == "Sierra", "ID"] <-
  substr(pre.tme.ctl[pre.tme.ctl$Site == "Sierra", "SampleName"], 10, 11)

# subdivide ORNL into WB and TVA
pre.tme.ctl[pre.tme.ctl$Site == "ORNL", "Site"] <- ifelse(grepl("TVA", pre.tme.ctl[pre.tme.ctl$Site == "ORNL", "ID"]), "Tennessee Valley", "Walker Branch")

# subdivide Sierra into Musick and Shaver
pre.tme.ctl[pre.tme.ctl$Site == "Sierra", "Site"] <- ifelse(grepl("M", pre.tme.ctl[pre.tme.ctl$Site == "Sierra", "ID"]), "Musick", "Shaver")

# add treatment col
pre.tme.ctl$Treatment <- "control"

pre.tme.trt <- data.frame(SampleName = unique(tme.trt.ts$SampleName))
pre.tme.trt$ID <- substr(pre.tme.trt$SampleName, 1, nchar(as.character(pre.tme.trt$SampleName))-2)
pre.tme.trt <- rbind(pre.tme.trt, pre.tme.ctl[pre.tme.ctl$Site == "misc", c("ID", "SampleName")])
pre.tme.trt$Treatment <- "storage duration"
pre.tme.trt$Site <- pre.tme.ctl[match(pre.tme.trt$ID, pre.tme.ctl$ID), "Site"]

pre.tme.14c <- rbind(pre.tme.ctl, pre.tme.trt)
pre.tme.14c$Type <- ifelse(grepl("G", pre.tme.14c$SampleName), "G", "F")
pre.tme.14c$Period <- "pre"
pre.tme.14c$Experiment <- "tme"

# create inc period df
inc.tme.14c <- pre.tme.14c
inc.tme.14c$Period <- "inc"
tme.14c.template <- rbind(pre.tme.14c, inc.tme.14c)
```

```{r combine soil obs data}
# 4. Summarize observational data from timeseries data by unique IDs (SampleName)
# summarize time and co2 data by experiment, period, treatment, and sample name
ts.epsn.avg <- ts %>%
  select(Experiment, ID, Period, Treatment, SampleName, gC_gS, dw_g, mgCO2.C_gC, time_d, time_d_cmtv)  %>%
  group_by(Experiment, ID, Period, Treatment, SampleName) %>%
  summarise_all(list(max = max), na.rm = TRUE)
ts.epsn.avg <- rename_at(ts.epsn.avg, vars(contains('max')), list(~sub("_max", "", .)))
```

```{r clean up 14C data}
# 6. Clean up ^14^C data and add external data points (tme experiment, Xplr control samples)
# Need to add SampleName col and variable columns Treatment, Period, and Experiment to all data frames in ams_results_ls 

# First add 14C and 13C data from Exploratories control samples to ams_results_ls
# note that neither 13C nor 14C was measured for the pre-incubation period for these samples
# get names for ams_results data frames
results.14c.nms <- names(ams_results_ls$`ams_jena_results-co2_2019-01-28`$`Beem-Miller_9.xlsx`)
# get names for relevant Experiment/Treatment/Period combination
arc_ctl_inc_SN <- as.character(all.14c.template[all.14c.template$Experiment == "arc" & all.14c.template$Treatment == "control" & all.14c.template$Period == "inc", "SampleName"])
tme_ctl_inc_xplr_SN <- unlist(lapply(ams_results_ls[1:2], function(x) unlist(lapply(x, "[[", "Probe"))))
tme_ctl_inc_xplr_SN <- tme_ctl_inc_xplr_SN[grep("HE", tme_ctl_inc_xplr_SN)]
if(length(tme_ctl_inc_xplr_SN) == 14) {
  tme_ctl_inc_xplr_SN <- unique(substr(tme_ctl_inc_xplr_SN, 1, nchar(tme_ctl_inc_xplr_SN)-2))
}

# add sublist for Exploratories 2011 (control) data and data frames for above subsets
ams_results_ls$exploratories_ischoening_14c.co2_2011$arc_ctl_inc <- data.frame(matrix(nrow = length(arc_ctl_inc_SN), ncol = length(results.14c.nms)))
ams_results_ls$exploratories_ischoening_14c.co2_2011$tme_ctl_inc_xplr <- data.frame(matrix(nrow = length(tme_ctl_inc_xplr_SN), ncol = length(results.14c.nms)))

# add names and fill SampleNames for new data frames
ams_results_ls$exploratories_ischoening_14c.co2_2011 <- lapply(
  ams_results_ls$exploratories_ischoening_14c.co2_2011, function(df) {
    names(df) <- results.14c.nms
    df <- data.frame(
      mapply(FUN = as,
             df, 
             sapply(ams_results_ls$`ams_jena_results-co2_2019-01-28`$`Beem-Miller_9.xlsx`, class),
             SIMPLIFY = FALSE),
      stringsAsFactors = FALSE)
  return(df)
})

ams_results_ls$exploratories_ischoening_14c.co2_2011$arc_ctl_inc[ , c("SampleName","Probe")] <- arc_ctl_inc_SN
ams_results_ls$exploratories_ischoening_14c.co2_2011$tme_ctl_inc_xplr[ , c("SampleName","Probe")] <- tme_ctl_inc_xplr_SN

# fill 14C data
ams_results_ls$exploratories_ischoening_14c.co2_2011 <- lapply(ams_results_ls$exploratories_ischoening_14c.co2_2011, function(df) {
  df$`X.14C....` <- xplr.ctl.iso[match(df$Probe, xplr.ctl.iso$EP_Plotid), "14c_CO2"]
  df$Treatment <- "control"
  df$Period <- "inc"
  return(df)
})

# add Experiment
ams_results_ls$exploratories_ischoening_14c.co2_2011$arc_ctl_inc$Experiment <- "arc"
ams_results_ls$exploratories_ischoening_14c.co2_2011$tme_ctl_inc_xplr$Experiment <- "tme"


# `ams_jena_results-co2_2018-10-12` and `ams_jena_results-co2_2018-11-09`
# `Beem-Miller_4.xlsx`, `Beem-Miller_5.xlsx`; `Beem-Miller_6.xlsx`, `Beem-Miller_7.xlsx`
# Add treatment, experiment, period cols
# Experiment = tme
# Treatment = storage duration
# Period = inc
ams_results_ls[1:2] <- lapply(ams_results_ls[1:2], function(ls) lapply(ls, function(df) {
  df$Experiment <- "tme"
  df$Treatment <- "storage duration"
  df$Period <- "inc"
  return(df)
  })
)

# `ams_jena_results-co2_2019-01-28`
# `Beem-Miller_9.xlsx`
# Experiment = arc
# Treatment = air-dry + storage
# Period = pre
ams_results_ls$`ams_jena_results-co2_2019-01-28`$`Beem-Miller_9.xlsx`$SampleName <- ams_results_ls$`ams_jena_results-co2_2019-01-28`$`Beem-Miller_9.xlsx`$Probe
ams_results_ls$`ams_jena_results-co2_2019-01-28`$`Beem-Miller_9.xlsx`$Treatment <- "air-dry + storage"
ams_results_ls$`ams_jena_results-co2_2019-01-28`$`Beem-Miller_9.xlsx`$Period <- "pre"
ams_results_ls$`ams_jena_results-co2_2019-01-28`$`Beem-Miller_9.xlsx`$Experiment <- "arc"

# ams_results_ls$`ams_jena_results-co2_2019-02-25`$
# Archived_Incubations_30.xlsx
# Experiment = arc
# Treatment = air-dry + storage
# Period = inc
ams_results_ls$`ams_jena_results-co2_2019-02-25`$Archived_Incubations_30.xlsx$SampleName <- substr(ams_results_ls$`ams_jena_results-co2_2019-02-25`$Archived_Incubations_30.xlsx$Probe, 1, 9)
ams_results_ls$`ams_jena_results-co2_2019-02-25`$Archived_Incubations_30.xlsx$Treatment <- "air-dry + storage"
ams_results_ls$`ams_jena_results-co2_2019-02-25`$Archived_Incubations_30.xlsx$Period <- "inc"
ams_results_ls$`ams_jena_results-co2_2019-02-25`$Archived_Incubations_30.xlsx$Experiment <- "arc"

# ams_results_ls$`ams_jena_results-co2_2019-04-30`
# `Rost_31-2.xlsx`
# Experiment = arc
# Treatment = air-dry + storage
# Period = inc, pre (SEW11-2-2)
ams_results_ls$`ams_jena_results-co2_2019-04-30`$`Rost_31-2.xlsx`$SampleName <- substr(ams_results_ls$`ams_jena_results-co2_2019-04-30`$`Rost_31-2.xlsx`$Probe, 1, 9)
ams_results_ls$`ams_jena_results-co2_2019-04-30`$`Rost_31-2.xlsx`$Treatment <- "air-dry + storage"
ams_results_ls$`ams_jena_results-co2_2019-04-30`$`Rost_31-2.xlsx`$Experiment <- "arc"
ams_results_ls$`ams_jena_results-co2_2019-04-30`$`Rost_31-2.xlsx`[2:9, "Period"] <- "inc"
ams_results_ls$`ams_jena_results-co2_2019-04-30`$`Rost_31-2.xlsx`[10, "Period"] <- "pre"

# ams_results_ls$`ams_jena_results-co2_2020-02-13`
# Rost_47.xlsx 
# Experiment = rewet
# Treatment = control
# Period = pre
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_47.xlsx$SampleName <- substr(ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_47.xlsx$Probe, 1, 7)
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_47.xlsx$Treatment <- "control"
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_47.xlsx$Experiment <- "rewet"
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_47.xlsx$Period <- "pre"
# Rost_50.xlsx
# Experiment = rewet
# Treatment = control
# Period = inc
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_50.xlsx$SampleName <- substr(ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_50.xlsx$Probe, 1, 7)
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_50.xlsx$Treatment <- "control"
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_50.xlsx$Experiment <- "rewet"
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_50.xlsx$Period <- "inc"
# Rost_52.xlsx
# Experiment = rewet
# Treatment = air-dry
# Period = pre, inc
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_52.xlsx$SampleName <- substr(ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_52.xlsx$Probe, 1, 11)
pre.ix <- which(grepl("PRE", ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_52.xlsx$Probe))
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_52.xlsx$Treatment <- "air-dry"
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_52.xlsx$Experiment <- "rewet"
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_52.xlsx[pre.ix, "Period"] <- "pre"
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_52.xlsx[-pre.ix, "Period"] <- "inc"
```

```{r combine 14C soil obs data}
# 7. Combine data.
# flatten list and remove AMS blanks, NA samples, and non-Exploratory data
ams_results_arc_rewet <- lapply(ams_results_ls[3:7], function(x) bind_rows(x))
ams_results_arc_rewet.df <- bind_rows(lapply(ams_results_arc_rewet, function(x) extra.clean.fx(x)))

# flatten external data list
ams_results_tme <- lapply(ams_results_ls[c(1:2,7)], function(x) lapply(x, tme.14c.fx))
ams_results_tme.df <- bind_rows(lapply(ams_results_tme, function(x) bind_rows(x)))

# cut out non-relevant AMS results columns
nms.ams.vars <- c("F14C", "err", "X.14C....", "err....",
                 "SampleName", "Treatment", "Period", "Experiment")
ams_results_arc_rewet.df <- ams_results_arc_rewet.df[ , nms.ams.vars]
colnames(ams_results_arc_rewet.df) <- c("fm", "fm_err", "d14c", "d14c_err", nms.ams.vars[5:8])

ams_results_tme.df <- ams_results_tme.df[ , nms.ams.vars]
colnames(ams_results_tme.df) <- c("fm", "fm_err", "d14c", "d14c_err", nms.ams.vars[5:8])
     
# complicated, but the following checks to see if:
# 1) The end of the sample name string is "-n" where "n" is a number
# 2) if true, the final two characters are trimmed
# 3) if false, the name is retained as is (e.g. for tme ctl samples)
ams_results_tme.df$ID <- ifelse(substr(substr(ams_results_tme.df$SampleName, 1, nchar(ams_results_tme.df$SampleName)-1),
              nchar(substr(ams_results_tme.df$SampleName, 1, nchar(ams_results_tme.df$SampleName)-1)),
              nchar(substr(ams_results_tme.df$SampleName, 1, nchar(ams_results_tme.df$SampleName)-1))) == "-",
       substr(ams_results_tme.df$SampleName, 1, nchar(ams_results_tme.df$SampleName)-2),
       ams_results_tme.df$SampleName)
ams_results_tme.df$Type <- ifelse(grepl("G", ams_results_tme.df$ID), "G", "F")

# merge arc and rewet 14C data with template
if(nrow(merge(all.14c.template, ams_results_arc_rewet.df, all.x = TRUE)) == nrow(all.14c.template)) {
  all.14c <- merge(all.14c.template, ams_results_arc_rewet.df, all.x = TRUE)
} else {
  cat("duplicate sample merge error")
}

# merge soil obs data with 14c data
if(nrow(merge(all.14c, ts.epsn.avg, by = c("SampleName", "Period"))) == nrow(all.14c)) {
  all.14c <- merge(all.14c, ts.epsn.avg, all.x = TRUE)
} else {
  cat("duplicate sample merge error")
}

# merge tme 14C data with template and ts.epsn.avg
tme.14c <- merge(tme.14c.template, ams_results_tme.df, all.x = TRUE)

# add ctl sample 14C data to tme.14c
# first remove NA values
tme.14c.clean <- tme.14c[which(!is.na(tme.14c$d14c) | !is.na(tme.14c$fm)), ]
# define merge fx
tme.mrg.fx <- function(x, y) {
  mrg.nms <- c("SampleName", "ID", "Treatment", "Period")
  var.nms <- c("YearSampled", "Site", "Horizon",  "d14c", "fm")
  if(length(grep("fm", names(y))) > 0) {
    df <- merge(x[ , mrg.nms], y[ , c(mrg.nms, var.nms)])
  } else {
    df <- merge(x[ , mrg.nms], y[ , c(mrg.nms, var.nms[1:4])])
    df$fm <- NA
  }
  return(df)
}

tme.14c.df <- suppressMessages(bind_rows(lapply(list(tme.ctl.duk,
                                                     tme.ctl.flx.sra[which(!is.na(tme.ctl.flx.sra$d14c)), ],
                                                     tme.ctl.ornl,
                                                     tme.misc), 
                                                function(i) tme.mrg.fx(tme.14c, i))))

# add storage duration data for treatment samples analyzed externally
tme.14c.df$Type <- "F"

# add YearSampled for tme.14c.clean samples
for(i in 1:nrow(tme.14c.clean)) {
  if(tme.14c.clean[i, "Site"] == "Duke") {
    tme.14c.clean[i, "YearSampled"] <- 2008
  } else if(tme.14c.clean[i, "Site"] == "Hainich") {
    tme.14c.clean[i, "YearSampled"] <- 2011
  } else if(tme.14c.clean[i, "Site"] == "Musick" | tme.14c.clean[i, "Site"] == "Shaver") {
    tme.14c.clean[i, "YearSampled"] <- 2009
  } else {
    tme.14c.clean[i, "YearSampled"] <- 2004
  }
}

# Add Horizon to tme.14c.clean
tme.14c.clean$Horizon <- ifelse(tme.14c.clean$ID == "MB" | tme.14c.clean$ID == "SB", "B", "A")

# combine tme data
tme.14c <- bind_rows(tme.14c.clean[ , names(tme.14c.df)], tme.14c.df)
tme.14c$Experiment <- "tme"

# remove outlier (MA-1)
# *** Need to clarify method for determining that this is an outlier *** 
tme.14c <- tme.14c[-which(tme.14c$SampleName == "MA-1" | tme.14c$SampleName == "HEW49-1" | tme.14c$SampleName == "HEW26-1"), ]


# Add missing variable columns to all.14c
# add YearSampled column
all.14c$YearSampled <- ifelse(all.14c$Experiment == "arc", 2011, 2019)

# add Site and Horizon columns
all.14c$Site <- ifelse(grepl("HE", all.14c$ID), "Hainich", "Schorfheide")
all.14c$Horizon <- "A"


# combine tme.14c and all.14c into a list and calculate fraction modern from d14C as needed
all.14c.ls <- lapply(list(tme.14c, all.14c), function(df) {
  df$fm <- ifelse(is.na(df$fm),
                  d14c.fm.fx(df$YearSampled, df$d14c),
                  df$fm)
  df$d14c_corr <- decay.corr.fx(df$YearSampled, df$fm)
  return(df)
})
save(all.14c.ls, file = "all.14c.ls.RData")

all.14c.sum <- lapply(all.14c.ls, function(df) {
  df.sum <- select(df, -SampleName) %>%
    group_by(ID, Type, Period, Treatment, Experiment, Site, Horizon, YearSampled) %>%
    summarise_all(list(mean = mean, max = max, min = min, sd = sd), na.rm = TRUE)
  df.sum[mapply(is.infinite, df.sum)] <- NA
  return(df.sum)
})
all.14c.sum <- rapply(all.14c.sum, f = function(x) ifelse(is.nan(x), NA , x), how = "replace")
all.14c.sum.df <- suppressMessages(bind_rows(all.14c.sum))
save(all.14c.sum.df, file = "all.14c.sum.df.RData")
```

```{r read-calc-13c}
# Clean up 13C data
# Note that iso data IDs should be stored as the "preparation" variable

# first create a library of possible IDs
all.ids <- unique(c(all.14c.template$ID, tme.14c.template$ID))

# function to strip out all values not corresponding to known IDs and return mean of duplicates
clean.iso <- function(df) {
  
  # strip out non arc-inc data
  ix <- vector(mode = "list", length = length(all.ids))
  for(i in seq_along(all.ids)) {
    if(any(grepl(all.ids[i], df$preparation))) {
      ix[[i]] <- which(grepl(all.ids[i], df$preparation))
    } 
  }
  
  ix <- unlist(ix)
  
  # convert 13C col to numeric
  df <- data.frame(df[ix, ])
  df$offset.corr.sample <- as.numeric(df$offset.corr.sample)
  
  # convert to data frame from tibble
  df <- data.frame(df)
  
  # average duplicates
  d13c <- vector()
  for(i in seq_along(df$preparation)) {
        if(i != nrow(df)) {
          d13c[i] <- ifelse(df[i, "preparation"] == df[i + 1, "preparation"],
                            mean(df[i, "offset.corr.sample"], df[i+1, "offset.corr.sample"], na.rm = TRUE),
                            NA)
      }
  }
  df <- df[complete.cases(df$offset.corr.sample), ]
  return(data.frame(SampleName = unique(df$preparation),
                    d13c = d13c[complete.cases(d13c)]))
}

iso_results_ls_clean <- lapply(iso_results_ls, function(df) {
  lapply(df, clean.iso)
})

# add in Experiment, treatment, period, etc.
# 1. iso_jena_results-co2_2019-01-22
## 13C_archivecInc_Pre_20181207-141691-141751.xls
iso_results_ls_clean[[1]][[1]]["Period"] <- "pre"
iso_results_ls_clean[[1]][[1]]["Experiment"] <- "arc"
iso_results_ls_clean[[1]][[1]]["Treatment"] <- "air-dry + storage"

## 13C_archivedInc_20190116-142012-142072.xls
iso_results_ls_clean[[1]][[2]]["Period"] <- "inc"
iso_results_ls_clean[[1]][[2]]["Experiment"] <- "arc"
iso_results_ls_clean[[1]][[2]]["Treatment"] <- "air-dry + storage"

# 2. iso_jena_results-co2_2019-01-23
## 20190122-ManuelRost-142082-141553.xls`
iso_results_ls_clean[[2]][[1]]["Period"] <- "inc"
iso_results_ls_clean[[2]][[1]]["Experiment"] <- "arc"
iso_results_ls_clean[[2]][[1]]["Treatment"] <- "air-dry + storage"

# 3. iso_jena_results-co2_2019-09-10
## 20190909-2Wh-Jeffrey_Beem-Miller-13613-13658.xls
iso_results_ls_clean[[3]][[1]]["Period"] <- "pre"
iso_results_ls_clean[[3]][[1]]["Experiment"] <- "rewet"
iso_results_ls_clean[[3]][[1]]["Treatment"] <- "control"

# 4. iso_jena_results-co2_2019-09-11
## 20190910-Jeffrey_Beem-Miller-13664-13702.xls
iso_results_ls_clean[[4]][[1]]["Period"] <- "inc"
iso_results_ls_clean[[4]][[1]]["Experiment"] <- "rewet"
iso_results_ls_clean[[4]][[1]]["Treatment"] <- "control"

# 5. iso_jena_results-co2_2019-12-04
## 20191126-MRost-BeemMiller-17509-17555.xls
iso_results_ls_clean[[5]][[1]]["Period"] <- "inc"
iso_results_ls_clean[[5]][[1]]["Experiment"] <- "rewet"
iso_results_ls_clean[[5]][[1]]["Treatment"] <- "air-dry"

# 6. iso_jena_results-co2_2019-12-21
## 20190213-ManuelRost-143192-143294.xls
iso_results_ls_clean[[6]][[1]]["Period"] <- "inc"
iso_results_ls_clean[[6]][[1]]["Experiment"] <- "arc"
iso_results_ls_clean[[6]][[1]]["Treatment"] <- "air-dry + storage"

## 20191106-MRost-16925-16952.xls
iso_results_ls_clean[[6]][[2]]["Period"] <- "inc"
iso_results_ls_clean[[6]][[2]]["Experiment"] <- "rewet"
iso_results_ls_clean[[6]][[2]]["Treatment"] <- "control"

## ArcInc_dry_13C.xls
iso_results_ls_clean[[6]][[3]]["Period"] <- "pre"
iso_results_ls_clean[[6]][[3]]["Experiment"] <- "rewet"
iso_results_ls_clean[[6]][[3]]["Treatment"] <- "air-dry"

# Flatten list
iso_df <- suppressWarnings(bind_rows(unlist(iso_results_ls_clean, recursive = FALSE)))
save(iso_df, file = "iso_df.RData")

# extract original IDs
og.sn <- vector()
for(i in seq_along(iso_df$SampleName)) {
  og.sn[i] <- ifelse(grepl("HE", iso_df$SampleName[i]),
                     substr(iso_df$SampleName[i], 
                            start = unlist(gregexpr("HE", iso_df$SampleName[i])), 
                            stop = unlist(gregexpr("HE", iso_df$SampleName[i])) + 4),
                     substr(iso_df$SampleName[i], 
                            start = unlist(gregexpr("SE", iso_df$SampleName[i])), 
                            stop = unlist(gregexpr("SE", iso_df$SampleName[i])) + 4))
}
iso_df$ID <- og.sn
iso_df$Type <- ifelse(grepl("G", iso_df$ID), "G", "F")

# add control sample 13C for arc experiment
arc.ctl.inc.IDs <- unique(iso_df[iso_df$Experiment == "arc" & iso_df$Period == "inc", "ID"])
arc.ctl.inc.13c <- data.frame(ID = arc.ctl.inc.IDs,
                              SampleName = arc.ctl.inc.IDs,
                              d13c = xplr.ctl.iso[which(!is.na(match(xplr.ctl.iso$EP_Plotid, arc.ctl.inc.IDs))), "13c_CO2"],
                              Experiment = "arc",
                              Period = "inc",
                              Treatment = "control",
                              Type = ifelse(grepl("G", arc.ctl.inc.IDs), "G", "F"))
iso_df <- rbind(iso_df, arc.ctl.inc.13c)
# ggplot(iso_df, aes(ID, d13c, color = Treatment, shape = Period)) +
#     geom_point(size = 3) +
#     facet_grid(rows = vars(Experiment)) +
#     theme_bw()
```


```{r counts for checking plots, include = FALSE}
# 8. Count number of ^14^C observations for checking plots.
# count !is.na(14C) cases by SampleName
all.14c[-which(is.na(all.14c$d14c_corr)),] %>%
  group_by(Experiment, Treatment, Period, Type) %>%
  count()
# by ID
all.14c[-which(is.na(all.14c$d14c_corr)),] %>%
  group_by(Experiment, Treatment, Period, Type) %>%
  distinct(ID) %>%
  count()
```

*Pre-incubation versus equilibrium respiration ^14^C-CO~2~*

```{r pre-inc-14c-plot}
all.14c.pre.inc <- merge(all.14c.sum.df[all.14c.sum.df$Period == "pre", c("ID", "Experiment", "Type", "Treatment", "d14c_corr_mean", "d14c_corr_min", "d14c_corr_max")],
                         all.14c.sum.df[all.14c.sum.df$Period == "inc", c("ID", "Experiment", "Type", "Treatment", "d14c_corr_mean", "d14c_corr_min", "d14c_corr_max")],
                         by = c("ID", "Experiment", "Type", "Treatment"),
                         suffixes = c("_pre", "_inc"))
all.14c.pre.inc$treat.bi <- ifelse(all.14c.pre.inc$Treatment == "control", "control", "treatment")

# save
save(all.14c.pre.inc, file = "./all.14c.pre.inc.RData")

all.14c.pre.inc %>%
  filter(Experiment != "tme") %>%
  ggplot(., aes(d14c_corr_mean_pre, d14c_corr_mean_inc, color = Type, shape = Treatment)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_vline(xintercept = 0, color = "gray") +
  geom_hline(yintercept = 0, color = "gray") +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = d14c_corr_min_inc, 
        ymax = d14c_corr_max_inc, 
        color = Type), 
    width = .25) +
  geom_errorbarh(
    aes(xmin = d14c_corr_min_pre, 
        xmax = d14c_corr_max_pre, 
        color = Type), 
    height = .9) +
  scale_color_manual(name = 'Ecosystem',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  scale_shape_manual(name = 'Treatment',
                     values = c("control" = 16,
                                "air-dry" = 1,
                                "air-dry + storage" = 0),
                     labels = c("control" = "control-2",
                                "air-dry" = "air-dry/rewet",
                                "air-dry + storage" = "air-dry/rewet + storage")) +
  scale_x_continuous(limits = c(-60, 115)) +
  scale_y_continuous(limits = c(-60, 115)) +
  xlab(expression('Rewetting pulse '*Delta*''^14*'C-CO'[2]*' ()')) +
  ylab(expression('Equilibrium respiration '*Delta*''^14*'C-CO'[2]*' ()')) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.key.height=unit(.8, "cm"),
        aspect.ratio = 1)
```
>**Fig. `r {fig.n + 1}`. $\Delta$^14^C-CO~2~ of the rewetting pulse and the equilibrium respiration period**

>*Caption:* Points are means of laboratory duplicates and error bars are the min and max (except for Experiment 1 control samples, which were not replicated). Note that rewetting pulse $\Delta$^14^C was not measured for control-1 samples; additionally samples from three of the forest plots of the air-dry + storage samples from Experiment 1 failed to accumulate enough CO~2~ during the pre-incubation period to measure $\Delta$^14^C. The outlier point with the substantially depleted pre-incubation $\Delta$^14^C is from Experiment 2 (control).

```{r pre-inc-13c-plot, include = FALSE}
iso.12.avg <- iso_df %>%
  select(ID, Experiment, Period, Treatment, d13c) %>%
  group_by(ID, Experiment, Period, Treatment) %>%
  add_count() %>%
  summarize_all(list(mean = mean, sd = sd), na.rm = TRUE) %>%
  mutate(d13c_se = d13c_sd/n_mean,
          d13c_u = d13c_mean + d13c_se*2,
          d13c_l = d13c_mean - d13c_se*2)

iso.12.avg$Type <- ifelse(grepl("G", iso.12.avg$ID), "G", "F")
iso.12.avg <- data.frame(iso.12.avg)

# stats
# air-dry + storage
# forest
t.test(iso.12.avg[iso.12.avg$Period == "inc" & iso.12.avg$Treatment == "air-dry + storage" & iso.12.avg$Type == "F", "d13c_mean"],
       iso.12.avg[iso.12.avg$Period == "inc" & iso.12.avg$Experiment == "arc" & iso.12.avg$Treatment == "control" & iso.12.avg$Type == "F", "d13c_mean"], 
       tails = 2, paired = TRUE)

# grassland
t.test(iso.12.avg[iso.12.avg$Period == "inc" & iso.12.avg$Treatment == "air-dry + storage" & iso.12.avg$Type == "G", "d13c_mean"],
       iso.12.avg[iso.12.avg$Period == "inc" & iso.12.avg$Experiment == "arc" & iso.12.avg$Treatment == "control" & iso.12.avg$Type == "G", "d13c_mean"],
       tails = 2, paired = TRUE)

# air-dry only
# forest
t.test(iso.12.avg[iso.12.avg$Period == "inc" & iso.12.avg$Treatment == "air-dry" & iso.12.avg$Type == "F", "d13c_mean"],
       iso.12.avg[iso.12.avg$Period == "inc" & iso.12.avg$Experiment == "rewet" & iso.12.avg$Treatment == "control" & iso.12.avg$Type == "F", "d13c_mean"],
       tails = 2, paired = TRUE)

# grassland
t.test(iso.12.avg[iso.12.avg$Period == "inc" & iso.12.avg$Treatment == "air-dry" & iso.12.avg$Type == "G", "d13c_mean"],
       iso.12.avg[iso.12.avg$Period == "inc" & iso.12.avg$Experiment == "rewet" & iso.12.avg$Treatment == "control" & iso.12.avg$Type == "G", "d13c_mean"], 
       tails = 2, paired = TRUE)

iso.12.pre.inc <- merge(iso.12.avg[iso.12.avg$Period == "pre", c("ID", "Experiment", "Type", "Treatment", "d13c_mean", "d13c_u", "d13c_l")],
                        iso.12.avg[iso.12.avg$Period == "inc", c("ID", "Experiment", "Type", "Treatment", "d13c_mean", "d13c_u", "d13c_l")],
                         by = c("ID", "Experiment", "Type", "Treatment"),
                         suffixes = c("_pre", "_inc"))
iso.12.pre.inc$treat.bi <- ifelse(iso.12.pre.inc$Treatment == "control", "control", "treatment")
save(iso.12.pre.inc, file = "iso.12.pre.inc.RData")

# range(c(iso.12.pre.inc$d13c_l_pre, iso.12.pre.inc$d13c_l_inc, iso.12.pre.inc$d13c_u_inc, iso.12.pre.inc$d13c_l_inc), na.rm = TRUE)
iso.12.pre.inc %>%
  ggplot(., aes(d13c_mean_pre, d13c_mean_inc, color = Type, shape = Treatment)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = d13c_l_inc, 
        ymax = d13c_u_inc, 
        color = Type), 
    width = .25) +
  geom_errorbarh(
    aes(xmin = d13c_l_pre, 
        xmax = d13c_u_pre,
        color = Type), 
    height = .25) +
  scale_color_manual(name = 'Ecosystem',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  scale_shape_manual(name = 'Treatment',
                     values = c("control" = 16,
                                "air-dry" = 1,
                                "air-dry + storage" = 0),
                     labels = c("control" = "control-1 (2011)\ncontrol-2 (2019)",
                                "air-dry" = "air-dry",
                                "air-dry + storage" = "air-dry + storage")) +
  scale_x_continuous(limits = c(-29.1, -22.25)) +
  scale_y_continuous(limits = c(-29.1, -22.25)) +
  xlab(expression('Pre-incubation '*delta*''^13*'C ()')) +
  ylab(expression('Equilibrium respiration '*delta*''^13*'C ()')) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.key.height=unit(.8, "cm"),
        aspect.ratio = 1)
```

```{r pre-equ-14c-stats, include = FALSE}
# gather pre and inc data from all reps, arc and rewet exp only
j.vars <- c("SampleName", "Site", "ID", "Experiment", "Treatment", "Type", "d14c_corr", "mgCO2.C_gC")
arc.rewet.pre.14c <- bind_rows(lapply(all.14c.ls, function(df) {
  if(!is.null(df$mgCO2.C_gC)) {
      df <- df[which(df$Period == "pre" & df$Experiment != "tme"), j.vars]
      df$treat.bi <- ifelse(df$Treatment == "control", "control", "treatment")
  return(df)
  }
}))
arc.rewet.inc.14c <- bind_rows(lapply(all.14c.ls, function(df) {
  if(!is.null(df$mgCO2.C_gC)) {
  df <- df[which(df$Period == "inc" & df$Experiment != "tme"), j.vars]
  df$treat.bi <- ifelse(df$Treatment == "control", "control", "treatment")
  return(df)
  }
}))
arc.rewet.14c <- merge(arc.rewet.pre.14c, arc.rewet.inc.14c, 
                       by = c(j.vars[1:6], "treat.bi"), suffixes = c("_pre", "_inc"))
arc.rewet.14c$YearSampled <- ifelse(grepl("arc", arc.rewet.14c$Experiment), 2011, 2019)
save(arc.rewet.14c, file = "arc.rewet.14c.RData")
```

```{r pre-equ-13c-stats, include = FALSE}
# gather pre and inc data from all reps, arc and rewet exp only
iso_df$treat.bi <- ifelse(iso_df$Treatment == "control", "control", "treatment")
arc.rewet.13c.stats.ls <- lapply(
  split(iso_df, iso_df$Experiment), function(exp) 
    lapply(split(exp, exp$Period), function(per) 
      lapply(split(per, per$treat.bi), function(trt) {
            # calculate mean and var of reps
              bind_rows(
                lapply(split(trt, trt$ID), function(id) {
                  data.frame(m = mean(id$d13c),
                             v = var(id$d13c),
                             Type = unique(id$Type))
                }), .id = "ID")})))

# calculate mean rep var
mean.rep.var.13c <- mean(unlist(lapply(arc.rewet.13c.stats.ls, function(exp)
  lapply(exp, function(per)
    lapply(per, function(trt) {
      mean(trt$v, na.rm = T)
    })))), na.rm = T)

# add rep var as needed
arc.rewet.13c.stats.ls <- lapply(arc.rewet.13c.stats.ls, function(exp)
  lapply(exp, function(per)
    lapply(per, function(trt) {
      trt$v <- ifelse(is.na(trt$v), mean.rep.var.13c, trt$v)
      return(trt)
      })))

# calculate grand mean and pooled sd within exp, trt, type
# for plotting 13C
arc.rewet.13c.pooled <- bind_rows(lapply(arc.rewet.13c.stats.ls, function(exp)
  bind_rows(lapply(exp, function(per)
    bind_rows(lapply(per, function(trt)
      bind_rows(lapply(split(trt, trt$Type), function(id.sum) {
        d <- sample.decomp(n = rep(2, length(id.sum$m)),
                           sample.mean = id.sum$m,
                           sample.var = id.sum$v,
                           include.sd = TRUE)
        return(d[grepl("pooled", rownames(d)), ])
      }), .id = "Type")), .id = "treat.bi")), .id = "Period")), .id = "Experiment")
save(arc.rewet.13c.pooled, file = "arc.rewet.13c.pooled.RData")

# Calculate mean differences between enclosure periods
arc.rewet.13c.pre.inc.ls <- lapply(arc.rewet.13c.stats.ls, function(exp) {
  per.ls <- lapply(exp, function(per) bind_rows(per, .id = "treat.bi"))
  pre.inc.df <- merge(
    per.ls[[1]], per.ls[[2]], by = c("ID", "treat.bi"), suffixes = names(per.ls))
  # calculate mean difference
  pre.inc.df$dif <- pre.inc.df$mpre - pre.inc.df$minc
  # sum variances
  pre.inc.df$vdif <- pre.inc.df$vpre + pre.inc.df$vinc
  pre.inc.df$Type <- pre.inc.df$Typepre
  return(pre.inc.df[ , c("ID", "dif", "vdif", "Type", "treat.bi")])
})

# calculate mean of mean difference and pooled variance
arc.rewet.13c.pre.inc.stats <- lapply(arc.rewet.13c.pre.inc.ls, function(exp)
  lapply(split(exp, exp$treat.bi), function(trt)
    lapply(split(trt, trt$Type), function(id.sum) {
      sample.decomp(n = rep(2, length(id.sum$dif)),
                    sample.mean = id.sum$dif,
                    sample.var = id.sum$vdif,
                    include.sd = TRUE)})))

# Calculate mean differences for 2nd enclosure period ("inc")
arc.rewet.13c.inc.dif.ls <- lapply(arc.rewet.13c.stats.ls, function(exp) {
  inc.ls <- unlist(exp[which(names(exp) == "inc")], recursive = FALSE) 
  inc.df <- merge(inc.ls[[1]], inc.ls[[2]], by = "ID", suffixes = names(exp$inc))
  # calculate mean difference
  inc.df$dif <- inc.df$mtreatment - inc.df$mcontrol
  # sum variances
  inc.df$vdif <- inc.df$vcontrol + inc.df$vtreatment
  inc.df$Type <- inc.df$Typecontrol
  return(inc.df[ , c("ID", "dif", "vdif", "Type")])
})

# calculate mean of mean difference and pooled variance
arc.rewet.13c.stats.inc.ls <- lapply(arc.rewet.13c.inc.dif.ls, function(exp) {
  lapply(split(exp, exp$Type), function(id.sum) {
    sample.decomp(n = rep(2, length(id.sum$dif)),
                  sample.mean = id.sum$dif,
                  sample.var = id.sum$vdif,
                  include.sd = TRUE)
  })
})
```

## Treatment effect on $\Delta$^14^C-CO~2~ for all samples (Experiments 1, 2, and 3)

```{r stats-2021-05-21}
# calculate means and variances within experiment, period, type, and ID
all.14c.df <- rbind(all.14c.ls[[1]],
                    all.14c.ls[[2]][, which(names(all.14c.ls[[2]]) %in% names(all.14c.ls[[1]]))])
all.14c.df$treat.bi <- ifelse(all.14c.df$Treatment == "control", "control", "treatment")

all.14c.stats.ls <- lapply(
  # split by Experiment and restrict to A horizon
  split(all.14c.df[all.14c.df$Horizon == "A", ], 
        all.14c.df[all.14c.df$Horizon == "A", "Experiment"]), function(exp) {
    # split by Period
          # exclude Oak Ridge
          # if (exp[1, "Experiment"] == "tme") {
          #   exp <- exp[-which(exp$Site == "Tennessee Valley" | exp$Site == "Walker Branch"), ]
          # }
          lapply(split(exp, exp$Period), function(per) {
            # split by Treatment
            lapply(split(per, per$treat.bi), function(trt) {
            # calculate mean and var of reps
              bind_rows(
                lapply(split(trt, trt$ID), function(id) {
                  data.frame(m = mean(id$d14c_corr),
                             v = var(id$d14c_corr),
                             Type = unique(id$Type))
                }), .id = "ID")
            })
        })
      })

# Use mean of replicate variances across all experiments for missing variances
mean.rep.var <- mean(c(all.14c.stats.ls$rewet$inc$control$v,
                       all.14c.stats.ls$rewet$inc$treatment$v,
                       all.14c.stats.ls$arc$inc$treatment$v,
                       all.14c.stats.ls$tme$inc$control$v), na.rm = TRUE)
# add var where needed
all.14c.stats.ls <- lapply(all.14c.stats.ls, function(exp) {
  lapply(exp, function(per) {
    lapply(per, function(trt) {
      trt$v <- ifelse(is.na(trt$v), mean.rep.var, trt$v)
      return(trt)
      })
    })
  })

# calculate grand means and pooled variance w/in treatment groups and types
arc.rewet.14c.stats.inc.ls2 <- lapply(all.14c.stats.ls[1:2], function(exp)
  lapply(exp, function(per)
    lapply(per, function(trt)
      lapply(split(trt, trt$Type), function(id.sum) {
        id.sum <- id.sum[!is.na(id.sum$m), ]
        df <- sample.decomp(n = rep(2, length(id.sum$m)),
                             sample.mean = id.sum$m,
                             sample.var = id.sum$v,
                             include.sd = TRUE)
        df[, "ID"] <- c(id.sum$ID, "pooled")
        return(df)
      }))))

# Calculate mean differences for 2nd enclosure period ("inc")
all.14c.stats.inc.dif.ls <- lapply(all.14c.stats.ls, function(exp) {
  inc.ls <- unlist(exp[which(names(exp) == "inc")], recursive = FALSE) 
  inc.df <- merge(inc.ls[[1]], inc.ls[[2]], by = "ID", suffixes = names(exp$inc))
  # calculate mean difference
  inc.df$dif <- inc.df$mtreatment - inc.df$mcontrol
  # sum variances
  inc.df$vdif <- inc.df$vcontrol + inc.df$vtreatment
  inc.df$Type <- inc.df$Typecontrol
  return(inc.df[ , c("ID", "dif", "vdif", "Type")])
})

# calculate mean of mean difference and pooled variance
# arc, rewet
arc.rewet.14c.stats.inc.ls <- lapply(all.14c.stats.inc.dif.ls[1:2], function(exp) {
  lapply(split(exp, exp$Type), function(id.sum) {
    df <- sample.decomp(n = rep(2, length(id.sum$dif)),
                        sample.mean = id.sum$dif,
                        sample.var = id.sum$vdif,
                        include.sd = TRUE)
    df[, "ID"] <- c(id.sum$ID, "pooled")
    return(df)
  })
})
save(arc.rewet.14c.stats.inc.ls, file = "arc.rewet.14c.stats.inc.ls.RData")

## tme
# all
tme.14c.stats <- sample.decomp(n = rep(2, nrow(all.14c.stats.inc.dif.ls$tme)),
                               sample.mean = all.14c.stats.inc.dif.ls$tme$dif,
                               sample.var = all.14c.stats.inc.dif.ls$tme$vdif,
                               include.sd = TRUE)
tme.ci95.int <- tme.14c.stats[nrow(tme.14c.stats), "sample.sd"] / sqrt(tme.14c.stats[nrow(tme.14c.stats), "n"]) * qt(0.975, tme.14c.stats[nrow(tme.14c.stats), "n"] - 1)
tme.ci95 <- c(tme.14c.stats[nrow(tme.14c.stats), "sample.mean"] - tme.ci95.int,
              tme.14c.stats[nrow(tme.14c.stats), "sample.mean"] + tme.ci95.int)

# Oak Ridge
tme.OR <- all.14c.stats.inc.dif.ls$tme[c(which(grepl("WB", all.14c.stats.inc.dif.ls$tme$ID)),
                                         which(grepl("TVA", all.14c.stats.inc.dif.ls$tme$ID))), ]
tme.OR.stats <- sample.decomp(n = rep(2, nrow(tme.OR)),
                               sample.mean = tme.OR$dif,
                               sample.var = tme.OR$vdif,
                               include.sd = TRUE)
tme.OR.ci95.int <- tme.OR.stats[nrow(tme.OR.stats), "sample.sd"] / sqrt(tme.OR.stats[nrow(tme.OR.stats), "n"]) * qt(0.975, tme.OR.stats[nrow(tme.OR.stats), "n"] - 1)
tme.OR.ci95 <- c(tme.OR.stats[nrow(tme.OR.stats), "sample.mean"] - tme.OR.ci95.int,
                 tme.OR.stats[nrow(tme.OR.stats), "sample.mean"] + tme.OR.ci95.int)
# Non Oak Ridge
tme.nonOR <- all.14c.stats.inc.dif.ls$tme[-as.numeric(row.names(tme.OR)), ]
tme.nonOR.stats <- sample.decomp(n = rep(2, nrow(tme.nonOR)),
                                 sample.mean = tme.nonOR$dif,
                                 sample.var = tme.nonOR$vdif,
                                 include.sd = TRUE)
tme.nonOR.ci95.int <- tme.nonOR.stats[nrow(tme.nonOR.stats), "sample.sd"] / sqrt(tme.nonOR.stats[nrow(tme.nonOR.stats), "n"]) * qt(0.975, tme.nonOR.stats[nrow(tme.nonOR.stats), "n"] - 1)
tme.nonOR.ci95 <- c(tme.nonOR.stats[nrow(tme.nonOR.stats), "sample.mean"] - tme.nonOR.ci95.int,
                    tme.nonOR.stats[nrow(tme.nonOR.stats), "sample.mean"] + tme.nonOR.ci95.int)
# nonOR F
tme.nonOR.F <- tme.nonOR[tme.nonOR$Type == "F", ]
tme.nonOR.F.stats <- sample.decomp(n = rep(2, nrow(tme.nonOR.F)),
                                   sample.mean = tme.nonOR.F$dif,
                                   sample.var = tme.nonOR.F$vdif,
                                   include.sd = TRUE)
tme.nonOR.F.ci95.int <- tme.nonOR.F.stats[nrow(tme.nonOR.F.stats), "sample.sd"] /
  sqrt(tme.nonOR.F.stats[nrow(tme.nonOR.F.stats), "n"]) * 
  qt(0.975, tme.nonOR.F.stats[nrow(tme.nonOR.F.stats), "n"] - 1)
tme.nonOR.F.ci95 <- c(tme.nonOR.F.stats[nrow(tme.nonOR.F.stats), "sample.mean"] - tme.nonOR.F.ci95.int,
                    tme.nonOR.F.stats[nrow(tme.nonOR.F.stats), "sample.mean"] + tme.nonOR.F.ci95.int)
# nonOR G
tme.nonOR.G <- tme.nonOR[tme.nonOR$Type == "G", ]
tme.nonOR.G.stats <- sample.decomp(n = rep(2, nrow(tme.nonOR.G)),
                                   sample.mean = tme.nonOR.G$dif,
                                   sample.var = tme.nonOR.G$vdif,
                                   include.sd = TRUE)
tme.nonOR.G.ci95.int <- tme.nonOR.G.stats[nrow(tme.nonOR.G.stats), "sample.sd"] /
  sqrt(tme.nonOR.G.stats[nrow(tme.nonOR.G.stats), "n"]) * 
  qt(0.975, tme.nonOR.G.stats[nrow(tme.nonOR.G.stats), "n"] - 1)
tme.nonOR.G.ci95 <- c(tme.nonOR.G.stats[nrow(tme.nonOR.G.stats), "sample.mean"] - tme.nonOR.G.ci95.int,
                    tme.nonOR.G.stats[nrow(tme.nonOR.G.stats), "sample.mean"] + tme.nonOR.G.ci95.int)

# calculate for aggregated dataset
arc.rewet.14c.stats.inc.ls2 <- lapply(
  split(
    bind_rows(lapply(all.14c.stats.inc.dif.ls, function(df) bind_rows(df, .id = "type"))),
    bind_rows(lapply(all.14c.stats.inc.dif.ls, function(df) bind_rows(df, .id = "type")))[["type"]]), 
  function(id.sum)
    sample.decomp(n = rep(2, length(id.sum$dif)),
                  sample.mean = id.sum$dif,
                  sample.var = id.sum$vdif,
                  include.sd = TRUE))

# # homogeneity of variance?
# arc.rewet.14c.var.inc.ls <- lapply(arc.rewet.14c.stats.inc.ls, function(df) {
#   var.df <- bind_rows(df, .id = "Type")
#   var.df <- var.df[which(var.df$n < 3), ]
#   var.test(sample.var ~ Type, data = var.df)
# })
# # control vs. treatment, Exp 2
# rewet.ctl.trt.var <- lapply(all.14c.stats.ls$rewet$inc, function(df) {
#     var.df <- bind_rows(df, .id = "treat.bi")
#     return(var.test(v ~ treat.bi, data = var.df))
# })

# extract mean differences
arc.14c.inc.meanDif <- bind_rows(
  lapply(arc.rewet.14c.stats.inc.ls, function(exp) {
    bind_rows(
      lapply(exp, function(md) {
        df <- round(md[nrow(md), c("n", "sample.mean", "sample.sd")], 1)
        t <- qt(.975, df = max(df$n) - 1)
        df$se <- round(df$sample.sd/sqrt(df$n), 1)
        df$ci95.l <- round(df$sample.mean - df$se * t, 1)
        df$ci95.u <- round(df$sample.mean + df$se * t, 1)
        return(df)
    }), .id = "type")
}), .id = "Experiment") %>%
  mutate(Type = substr(type, nchar(type), nchar(type))) %>%
  select(c(Experiment, Type, n, sample.mean, ci95.l, ci95.u, sample.sd, se))
ggplot(arc.14c.inc.meanDif, aes(Type, sample.mean, color = Type)) +
  geom_hline(yintercept = 0) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = ci95.l, ymax = ci95.u), width = .1) +
  scale_color_manual(name = 'Ecosystem',
                     values = c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  coord_flip() +
  facet_grid(rows = vars(Experiment), 
             labeller = labeller(Experiment = c("arc" = "Experiment 1", 
                                                "rewet" = "Experiment 2",
                                                "tme" = "Experiment 3"))) +
  ylab(expression("Control - Treatment "*Delta^14*"C-CO"[2]*" ()")) +
  xlab(NULL) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text.y = element_blank())

# extract mean differences for overall dataset
arc.14c.inc.meanDif2 <- bind_rows(
  lapply(arc.rewet.14c.stats.inc.ls2, function(md) {
        df <- round(md[nrow(md), c("n", "sample.mean", "sample.sd")], 1)
        t <- qt(.975, df = max(df$n) - 1)
        df$se <- round(df$sample.sd/sqrt(df$n), 1)
        df$ci95.l <- round(df$sample.mean - df$se * t, 1)
        df$ci95.u <- round(df$sample.mean + df$se * t, 1)
        return(df)
    }), .id = "type") %>%
  mutate(Type = substr(type, nchar(type), nchar(type))) %>%
  select(c(Type, n, sample.mean, ci95.l, ci95.u, sample.sd, se))
```

```{r stats-pooled-2}
# arc + rewet w/ reps
arc.rewet.14c.reps <- unsplit(
  lapply(split(arc.rewet.14c, arc.rewet.14c$Experiment), function(exp) {
    unsplit(lapply(split(exp, exp$treat.bi), function(trt) {
      unsplit(lapply(split(trt, trt$ID), function(df) {
        df$rep <- seq(1, nrow(df))
        return(df)
      }), trt$ID) 
    }), exp$treat.bi)
  }), arc.rewet.14c$Experiment)
arc.rewet.14c.reps$rep <- factor(arc.rewet.14c.reps$rep)

# pre vs. main inc
# 14C
arc.rewet.14c.reps.pre <- arc.rewet.14c.reps[-which(is.na(arc.rewet.14c.reps$d14c_corr_pre)), ]
pre.inc.dif <- lapply(
  split(arc.rewet.14c.reps.pre, arc.rewet.14c.reps.pre$Experiment), function(exp) {
    lapply(split(exp, exp$treat.bi), function(trt) {
      trt.dif <- bind_rows(
        lapply(split(trt, trt$ID), function(id) {
          pre <- data.frame(m = mean(id$d14c_corr_pre),
                            v = var(id$d14c_corr_pre))
          inc <- data.frame(m = mean(id$d14c_corr_inc),
                            v = var(id$d14c_corr_inc))
          dif <- data.frame(d = pre$m - inc$m,
                            vd = pre$v + inc$v,
                            Type = unique(id$Type))
          dif$vd <- ifelse(is.na(dif$vd), mean.rep.var, dif$vd)
          return(dif)
          }), .id = "ID")
      trt.dif.pool <- lapply(split(trt.dif, trt.dif$Type), function(id.sum) {
        decomp <- sample.decomp(n = rep(2, length(id.sum$d)),
                                sample.mean = id.sum$d,
                                sample.var = id.sum$vd,
                                include.sd = TRUE)
        pooled <- decomp[nrow(decomp), ]
        plusMinus <- pooled$sample.sd / sqrt(pooled$n) * qt(.95, pooled$n - 1)
        ci95 <- c(pooled$sample.mean - plusMinus, pooled$sample.mean + plusMinus)
        return(list(decomp = decomp, ci95 = ci95))
        })
      return(trt.dif.pool)
      })
    })
# remove HEW22 from rewet control
pre.inc.dif$rewet$control$F$decomp <- pre.inc.dif$rewet$control$F$decomp[-1, ]
pre.inc.dif.all <- bind_rows(lapply(pre.inc.dif, function(exp) 
  bind_rows(lapply(exp, function(trt) 
    bind_rows(lapply(trt, function(typ)
      typ$decomp[1:nrow(typ$decomp) - 1, ]), 
      .id = "Type")), .id = "treat.bi")), .id = "Experiment")
pre.inc.dif.all.pooled <- sample.decomp(n = pre.inc.dif.all$n,
                                        sample.mean = pre.inc.dif.all$sample.mean,
                                        sample.var = pre.inc.dif.all$sample.var,
                                        include.sd = TRUE)
pre.inc.dif.all.pooled.ci <- c(
  pre.inc.dif.all.pooled[nrow(pre.inc.dif.all.pooled), "sample.mean"] -
    pre.inc.dif.all.pooled[nrow(pre.inc.dif.all.pooled), "sample.sd"] / 
    sqrt(pre.inc.dif.all.pooled[nrow(pre.inc.dif.all.pooled), "n"]) * 
    qt(.95, pre.inc.dif.all.pooled[nrow(pre.inc.dif.all.pooled), "n"] - 1),
  pre.inc.dif.all.pooled[nrow(pre.inc.dif.all.pooled), "sample.mean"] +
    pre.inc.dif.all.pooled[nrow(pre.inc.dif.all.pooled), "sample.sd"] / 
    sqrt(pre.inc.dif.all.pooled[nrow(pre.inc.dif.all.pooled), "n"]) * 
    qt(.95, pre.inc.dif.all.pooled[nrow(pre.inc.dif.all.pooled), "n"] - 1))

# 14c ctl-trt
arc.rewet.14c.ctl.trt.ls <- lapply(
  split(arc.rewet.14c.reps, arc.rewet.14c.reps$Experiment), function(exp) {
    trt.dif <- bind_rows(
        lapply(split(exp, exp$ID), function(id) {
          ctl.14c <- id[id$treat.bi == "control", "d14c_corr_inc"]
          ctl <- data.frame(m = mean(ctl.14c),
                            v = var(ctl.14c))
          if (nrow(ctl) == 1) {
            ctl$v <- mean.rep.var
          }
          trt.14c <- id[id$treat.bi == "treatment", "d14c_corr_inc"]
          trt <- data.frame(m = mean(trt.14c),
                            v = var(trt.14c))
          if (nrow(trt) == 1) {
            trt$v <- mean.rep.var
          }
          dif <- data.frame(d = ctl$m - trt$m,
                            vd = ctl$v + trt$v,
                            Type = unique(id$Type))
          return(dif)
          }), .id = "ID")
          trt.dif.pool <- lapply(split(trt.dif, trt.dif$Type), function(id.sum) {
            decomp <- sample.decomp(n = rep(2, length(id.sum$d)),
                                    sample.mean = id.sum$d,
                                    sample.var = id.sum$vd,
                                    include.sd = TRUE)
            pooled <- decomp[nrow(decomp), ]
            plusMinus <- pooled$sample.sd / sqrt(pooled$n) * qt(.95, pooled$n - 1)
            ci95 <- c(pooled$sample.mean - plusMinus, pooled$sample.mean + plusMinus)
            return(list(decomp = decomp, ci95 = ci95))
            })
          return(trt.dif.pool)
      })

# 13C
load("iso_df.RData")
# calc difs
iso_df$treat.bi <- ifelse(iso_df$Treatment == "control", "control", "treatment")
mean.rep.var.13c <- mean(
  unlist(lapply(split(iso_df, iso_df$Experiment), function(exp)
    lapply(split(exp, exp$Period), function(per)
      lapply(split(per, per$treat.bi), function(trt)
        lapply(split(trt, trt$ID), function(id) {
          var(id$d13c)
        })
      )
    )
  )), na.rm = T)
# pre-inc
iso_df.pre.inc <- merge(
  iso_df[iso_df$Period == "pre", c("ID", "Experiment", "treat.bi", "Type", "d13c")],
  iso_df[iso_df$Period == "inc", c("ID", "Experiment", "treat.bi", "Type", "d13c")],
  by = c("ID", "Experiment", "treat.bi", "Type"), suffixes = c("_pre", "_inc"))
iso_df.pre.inc.ls <- lapply(
  split(iso_df.pre.inc, iso_df.pre.inc$Experiment), function(exp) {
    lapply(split(exp, exp$treat.bi), function(trt) {
      trt.dif <- bind_rows(
        lapply(split(trt, trt$ID), function(id) {
          pre <- data.frame(m = mean(id$d13c_pre),
                            v = var(id$d13c_pre))
          inc <- data.frame(m = mean(id$d13c_inc),
                            v = var(id$d13c_inc))
          dif <- data.frame(d = pre$m - inc$m,
                            vd = pre$v + inc$v,
                            Type = unique(id$Type))
          return(dif)
          }), .id = "ID")
      trt.dif.pool <- lapply(split(trt.dif, trt.dif$Type), function(id.sum) {
        decomp <- sample.decomp(n = rep(2, length(id.sum$d)),
                                sample.mean = id.sum$d,
                                sample.var = id.sum$vd,
                                include.sd = TRUE)
        pooled <- decomp[nrow(decomp), ]
        plusMinus <- pooled$sample.sd / sqrt(pooled$n) * qt(.95, pooled$n - 1)
        ci95 <- c(pooled$sample.mean - plusMinus, pooled$sample.mean + plusMinus)
        return(list(decomp = decomp, ci95 = ci95))
        })
      return(trt.dif.pool)
      })
    })

# inc
iso_df.inc.ls <- lapply(
  split(iso_df[iso_df$Period == "inc", ], 
        iso_df[iso_df$Period == "inc", "Experiment"]), function(exp) {
          trt.dif <- bind_rows(
            lapply(split(exp, exp$ID), function(id) {
              ctl <- data.frame(m = mean(id[id$treat.bi == "control", "d13c"]),
                                v = var(id[id$treat.bi == "control", "d13c"]))
              if (nrow(ctl) == 1) {
                ctl$v <- mean.rep.var.13c
              }
              trt <- data.frame(m = mean(id[id$treat.bi == "treatment", "d13c"]),
                                v = var(id[id$treat.bi == "treatment", "d13c"]))
              dif <- data.frame(d = ctl$m - trt$m,
                                vd = ctl$v + trt$v,
                                Type = unique(id$Type))
              return(dif)
              }), .id = "ID")
          trt.dif.pool <- lapply(split(trt.dif, trt.dif$Type), function(id.sum) {
            decomp <- sample.decomp(n = rep(2, length(id.sum$d)),
                                    sample.mean = id.sum$d,
                                    sample.var = id.sum$vd,
                                    include.sd = TRUE)
            pooled <- decomp[nrow(decomp), ]
            plusMinus <- pooled$sample.sd / sqrt(pooled$n) * qt(.95, pooled$n - 1)
            ci95 <- c(pooled$sample.mean - plusMinus, pooled$sample.mean + plusMinus)
            return(list(decomp = decomp, ci95 = ci95))
            })
          return(trt.dif.pool)
      })
```

```{r 14c-trt-difs}
# Load atmospheric data
Datm <- rbind(graven, future14C)
save(Datm, file = "Datm.RData")
yrs <- Datm$Date

# Filter data to "inc", split by binary treatment variable, then merge
# collapse treatment variable into binary control/treatment
all.14c.sum.df$treat.bi <- ifelse(all.14c.sum.df$Treatment == "control", "control", "treatment")

# inc data
all.14c.ctl <- all.14c.sum.df[all.14c.sum.df$treat.bi == "control" & all.14c.sum.df$Period == "inc", ]
all.14c.ctl <- rename_at(all.14c.ctl, vars(contains('d14c')), list(~paste0(., "_ctl")))
all.14c.trt <- all.14c.sum.df[all.14c.sum.df$treat.bi == "treatment" & all.14c.sum.df$Period == "inc", ]
all.14c.trt <- rename_at(all.14c.trt, vars(contains('d14c')), list(~paste0(., "_trt")))
all.14c.ctl.trt <- inner_join(data.frame(all.14c.ctl) %>% 
                                select(ID, Experiment, Type, Site, Horizon, contains("d14c_corr")),
                              data.frame(all.14c.trt) %>% 
                                select(ID, Experiment, Type, Site, Horizon, contains("d14c_corr")),
                              by = c("ID", "Experiment", "Type", "Site", "Horizon"))

# First add YearSampled and YearMeasured columns
all.14c.ctl.trt$YearSampled <- unlist(all.14c.sum.df[match(all.14c.ctl.trt$ID, all.14c.sum.df$ID), "YearSampled"])
# correct archive sample YearSampled
all.14c.ctl.trt[all.14c.ctl.trt$Experiment == "arc" & all.14c.ctl.trt$Site == "Hainich", "YearSampled"] <- 2011

# add year measured column, modify for tme.misc samples, rewet samples
all.14c.ctl.trt$YearMeasured <- 2018
all.14c.ctl.trt[all.14c.ctl.trt$Experiment == "rewet", "YearMeasured"] <- 2019
all.14c.ctl.trt[which(all.14c.ctl.trt$ID %in% tme.misc$ID), "YearMeasured"] <- 2009

# calculate storage duration numerically and as factor
all.14c.ctl.trt$tme <- unlist(all.14c.ctl.trt$YearMeasured - all.14c.ctl.trt$YearSampled)
all.14c.ctl.trt$tme.f <- factor(all.14c.ctl.trt$tme)

# add difference column
# (expressed as difference of means)
all.14c.ctl.trt$ctl.trt.d14 <- all.14c.ctl.trt$d14c_corr_mean_ctl - all.14c.ctl.trt$d14c_corr_mean_trt
# trt-ctl makes Fig. 4 (1:1 line) and Fig. 5 (storage duration) more compatible
all.14c.ctl.trt$trt.ctl.d14 <- all.14c.ctl.trt$d14c_corr_mean_trt - all.14c.ctl.trt$d14c_corr_mean_ctl

# add atm 14C for year sampled and calculate delta-delta
all.14c.ctl.trt$atm.yr.sampled <- Datm[match(all.14c.ctl.trt$YearSampled, floor(Datm$Date)), "NHc14"]
all.14c.ctl.trt$dd.ctl <- all.14c.ctl.trt$d14c_corr_mean_ctl - all.14c.ctl.trt$atm.yr.sampled
all.14c.ctl.trt$dd.trt <- all.14c.ctl.trt$d14c_corr_mean_trt - all.14c.ctl.trt$atm.yr.sampled
all.14c.ctl.trt$ctl.trt.dd14 <- all.14c.ctl.trt$dd.ctl - all.14c.ctl.trt$dd.trt

# save Rdat file
save(all.14c.ctl.trt, file = "./all.14c.ctl.trt.RData")

# difs
e12.14c.ctl.trt.sum <- all.14c.ctl.trt %>%
  filter(Experiment != "tme") %>%
  select(Experiment, Type, d14c_corr_mean_ctl, d14c_corr_mean_trt, trt.ctl.d14, dd.ctl, dd.trt, ctl.trt.dd14) %>%
  group_by(Experiment, Type) %>%
  add_tally() %>%
  summarize_all(list(mean = mean, sd = sd), na.rm = TRUE)
e12.14c.ctl.trt.sum$trt.ctl.d14_sd <- sqrt(e12.14c.ctl.trt.sum$d14c_corr_mean_ctl_sd^2/e12.14c.ctl.trt.sum$n_mean + e12.14c.ctl.trt.sum$d14c_corr_mean_trt_sd^2/e12.14c.ctl.trt.sum$n_mean)

# ctl-ctl dif vs. trt-trt
# difs by ID
hai.id.df <- all.14c.ctl.trt[all.14c.ctl.trt$Site == "Hainich" & all.14c.ctl.trt$Experiment != "tme", ]
hai.ctl.ctl.trt.trt <- bind_rows(
  lapply(split(hai.id.df, hai.id.df$ID), function(df) {
    dif.ctl <- df[df$YearSampled == 2011, "d14c_corr_mean_ctl"] -
      df[df$YearSampled == 2019, "d14c_corr_mean_ctl"]
    dif.trt <- df[df$YearSampled == 2011, "d14c_corr_mean_trt"] -
      df[df$YearSampled == 2019, "d14c_corr_mean_trt"]
    return(data.frame(dif = c(dif.ctl, dif.trt),
                      Type = df$Type,
                      treat = c("ctl", "trt")))
  })
)
hai.ctl.ctl.trt.trt.df <- hai.ctl.ctl.trt.trt %>% 
  group_by(Type, treat) %>%
  add_tally() %>%
  summarize(across(dif:n, list(mean = mean, sd = sd))) %>%
  mutate(se = dif_sd / n_mean,
         t = qt(p = .05/2, df = n_mean - 1),
         ci.95.u = dif_mean + t * se,
         ci.95.l = dif_mean - t * se)
hai.ctl.ctl.trt.trt.df %>%
  mutate(Treatment = ifelse(treat == "ctl", "control", "air-dried/rewetted")) %>%
  ggplot(., aes(Treatment, dif_mean, fill = Type, color = Type)) +
  geom_col(position = position_dodge(width = .9)) +
  geom_errorbar(aes(ymin = ci.95.l, ymax = ci.95.u),
                position = position_dodge(width = .9), 
                width = .5) +
  scale_color_manual(name = 'Ecosystem',
                     values = c('F' = '#a35513',
                                'G' = '#1361a3'),
                     labels = c('F' = 'Forest',
                                'G' = 'Grassland')) +
  scale_fill_manual(name = 'Ecosystem',
                    values = c('F' = '#a35513',
                               'G' = '#1361a3'),
                    labels = c('F' = 'Forest',
                               'G' = 'Grassland')) +
  ylab(expression('2011 '*Delta*''^14*'C - 2019 '*Delta*''^14*'C ()')) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```

```{r plot-14c-all-ctl-trt}
# list of locations for recoding
loc <- list(
    "Duke" = "Duke FACE control (#3)",
    "Howland" = "Howland (#3)",
    "Harvard" = "Harvard Forest (#3)",
    "Hainich" = "Central Germany",
    "Schorfheide" = "Central Germany",
    "Musick" = "Sierra Nevada, CA (#3)",
    "Shaver" = "Sierra Nevada, CA (#3)",
    "Walker Branch" = "Oak Ridge (#3)",
    "Tennessee Valley" = "Oak Ridge (#3)")

# Plot on 1:1 line w/ diagonal lines showing  offsets
# first collapse sites for better visuals
all.14c.ctl.trt %>%
  filter(Horizon == "A") %>%
  mutate(Location = recode(Site, !!!loc)) %>%
  mutate(Location = ifelse(Location == "Central Germany" & Experiment == "rewet", "Central Germany (#2)", ifelse(Location == "Central Germany" & Experiment == "arc", "Central Germany (#1)", ifelse(Location == "Central Germany" & Experiment == "tme", "Central Germany (#3)", Location)))) %>%
  ggplot(., aes(d14c_corr_mean_ctl, d14c_corr_mean_trt, color = Location, shape = Type)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_abline(slope = 1, intercept = -20, linetype = "dashed") +
  geom_abline(slope = 1, intercept = 20, linetype = "dashed") +
  geom_abline(slope = 1, intercept = -40, linetype = "dotted") +
  geom_abline(slope = 1, intercept = 40, linetype = "dotted") +
  geom_vline(xintercept = 0, color = "gray") +
  geom_hline(yintercept = 0, color = "gray") +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = d14c_corr_min_trt, 
        ymax = d14c_corr_max_trt, 
        color = Location), 
    width = .25) +
  geom_errorbarh(
    aes(xmin = d14c_corr_min_ctl, 
        xmax = d14c_corr_max_ctl, 
        color = Location), 
    height = .9) +
  scale_color_manual(name = "Location (Experiment #)",
                     values = c('Duke FACE control (#3)' = '#01BB97',
                                'Sierra Nevada, CA (#3)' = '#FFC107',
                                'Central Germany (#1)' = 'black',
                                'Central Germany (#2)' = '#999999',
                                'Central Germany (#3)' = '#c6c6c6',
                                'Harvard Forest (#3)' = '#1E88E5',
                                'Oak Ridge (#3)' = '#D81B60')) +
  scale_shape_manual(name = 'Ecosystem',
                     values = c('F'= 17,'G' = 16),
                     labels = c('Forest','Grassland')) +
  xlab(expression('Control '*Delta*''^14*'C ()')) +
  ylab(expression('Treatment '*Delta*''^14*'C ()')) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.key.height=unit(.8, "cm"),
        aspect.ratio = 1)

# for plotting with color = Type (currently suppressed)
# all.14c.ctl.trt %>%
#   filter(Horizon == "A") %>%
#   mutate(Location = factor(recode(Site, !!!loc), exclude = TRUE)) %>%
#   ggplot(., aes(d14c_corr_mean_ctl, d14c_corr_mean_trt, color = Type, shape = Location)) +
#   geom_abline(slope = 1, intercept = 0) +
#   geom_abline(slope = 1, intercept = -20, linetype = "dashed") +
#   geom_abline(slope = 1, intercept = 20, linetype = "dashed") +
#   geom_abline(slope = 1, intercept = -40, linetype = "dotdash") +
#   geom_abline(slope = 1, intercept = 40, linetype = "dotdash") +
#   geom_vline(xintercept = 0, color = "gray") +
#   geom_hline(yintercept = 0, color = "gray") +
#   geom_point(size = 3) +
#   geom_errorbar(
#     aes(ymin = d14c_corr_min_trt, 
#         ymax = d14c_corr_max_trt, 
#         color = Type), 
#     width = .25) +
#   geom_errorbarh(
#     aes(xmin = d14c_corr_min_ctl, 
#         xmax = d14c_corr_max_ctl, 
#         color = Type), 
#     height = .9) +
#   scale_color_manual(name = 'Ecosystem',
#                      values =c('F'='#a35513','G'='#1361a3'),
#                      labels = c('Forest','Grassland')) +
#   xlab(expression('Control '*Delta*''^14*'C ()')) +
#   ylab(expression('Treatment '*Delta*''^14*'C ()')) +
#   theme_bw() +
#   theme(panel.grid = element_blank(),
#         legend.key.height=unit(.8, "cm"),
#         aspect.ratio = 1)
```
>**Fig. `r {fig.n + 2}`. Overall treatment effect on $\Delta$^14^C-CO~2~**

>*Caption:* Points show data from Experiments 1 and 3 (air-dry + storage treatment) and Experiment 2 (air-dry only treatment). Points are the mean of laboratory replicates (for replicated samples); error bars are 2x standard error. Solid line is 1:1. For context, the dashed and dotted lines show differences of 20 and $40, equivalent to the decline in $\Delta$^14^C in atmospheric CO~2~ over 4 and 8 y respectively, during the period of 2000 to 2020 (Graven et al. 2017).

## Storage duration effect on ^14^C-CO~2~ (Experiment 3)

```{r plot-dur-stor, fig.height=4}
# new location codes
loc2 <- list(
    "Duke" = "Duke FACE control",
    "Howland" = "Howland",
    "Harvard" = "Harvard Forest",
    "Hainich" = "Central Germany",
    "Schorfheide" = "Central Germany",
    "Musick" = "Sierra Nevada, CA",
    "Shaver" = "Sierra Nevada, CA",
    "Walker Branch" = "Oak Ridge",
    "Tennessee Valley" = "Oak Ridge")

# Plot storage duration
suppressWarnings(
  all.14c.ctl.trt %>%
    filter(Horizon == "A") %>%
    filter(Experiment != "rewet") %>%
    select(Type, tme.f, Site, trt.ctl.d14, tme) %>%
    group_by(Type, tme.f, Site) %>%
    add_tally() %>%
    summarise_all(list(mean = mean, sd = sd), na.rm = TRUE) %>%
    mutate(dif_se = trt.ctl.d14_sd/n_mean,
           dif_u = trt.ctl.d14_mean + dif_se*2,
           dif_l = trt.ctl.d14_mean - dif_se*2) %>%
    mutate(Location = factor(recode(Site, !!!loc2), exclude = TRUE)) %>%
    ggplot(., aes(tme_mean, trt.ctl.d14_mean, color = Location, shape = Type)) +
    geom_hline(yintercept = 0) +
    geom_hline(yintercept = -20, linetype = "dashed") +
    geom_hline(yintercept = 20, linetype = "dashed") +
    geom_hline(yintercept = 40, linetype = "dotted") +
    geom_hline(yintercept = -40, linetype = "dotted") +
    geom_point(size = 3, position = position_dodge2(width = .7, preserve = "single")) +
    geom_errorbar(
      aes(ymin = dif_l,
          ymax = dif_u, 
          color = Location),
      width = .7,
      position = position_dodge2(width = .7, preserve = "single")) +
    scale_color_manual(values = c('Duke FACE control' = '#01BB97',
                                  'Sierra Nevada, CA' = '#FFC107',
                                  'Central Germany' = 'black',
                                  'Harvard Forest' = '#1E88E5',
                                  'Oak Ridge' = '#D81B60')) +
    scale_shape_manual(name = 'Ecosystem',
                       values = c('F'= 17,'G' = 16),
                       labels = c('Forest','Grassland')) +
    scale_x_continuous(breaks = seq(5,13,2)) +
    ylab(expression('Treatment - Control '*Delta*''^14*'C ()')) +
    xlab("Storage duration (years)") +
    theme_bw() +
    theme(panel.grid = element_blank())
  )
```

>**Fig. `r {fig.n + 3}`. Change in ^14^C-CO~2~ in relation to storage duration**

>*Caption:* Data are from both Experiment 1 (in black) and Experiment 3 (all other points), averaged by site and ecosystem type. Points are the mean, error bars are 2x standard error. For context, the dashed and dotted lines are the same is in Fig. `r {fig.n + 3}` and show a difference of 20 and 40, equivalent to the decline in $\Delta$^14^C in atmospheric CO~2~ over 4 and 8 y respectively, during the period of 2000 to 2020 (Graven et al. 2017). Position of points jittered to avoid overplotting; storage duration has been rounded down to the nearest whole year.

```{r treatment effects by time, fig.height=4}
# # code for plotting data by ID
# all.14c.sum %>%
#   filter(Period == "inc") %>%
#   filter(!grepl("S", ID)) %>%
#   mutate(year.ID = paste(YearSampled, ID)) %>%
#   ggplot(., aes(YearSampled, d14c_corr_mean, color = Type)) +
#   geom_point(aes(shape = Treatment), size = 3) +
#   geom_path(aes(group = year.ID), 
#             arrow = arrow(type = "open", length = unit(.1, "inches"), ends = "last"),
#             show.legend = FALSE) +
#   geom_path(data = Datm, aes(Date, NHc14), color = "darkgray") +
#   scale_color_manual(name = 'Ecosystem',
#                      values =c('F'='#a35513','G'='#1361a3'),
#                      labels = c('Forest','Grassland')) +
#   scale_shape_manual(name = 'Treatment',
#                      values = c("air-dry" = 1, 
#                                 "air-dry + storage" = 0,
#                                 "control" = 16)) +
#   scale_x_continuous(limits = c(2010, 2020), breaks = c(2013, 2016, 2019)) +
#   scale_y_continuous(limits = c(-15, 100)) +
#   xlab('Year sampled') +
#   ylab(expression('mean '*Delta*''^14*'C ()')) +
#   facet_wrap(~ ID, ncol = 3) +
#   theme_bw() +
#   theme(panel.grid = element_blank())

# Add "Type" to Datm to display legend
Datm$Atmosphere <- "atmosphere"
atm.d14c.2011 <- Datm[Datm$Date == 2011.5, "NHc14"]
atm.d14c.2019 <- Datm[Datm$Date == 2019.5, "NHc14"]

# plot over time, averaged by Type
# first need to calculate group means
arc.rewet.14c.inc.GM <- all.14c.sum.df %>%
  filter(Experiment != "tme" & Period == "inc") %>%
  group_by(Experiment, Type, YearSampled, Treatment) %>%
  summarize(across(d14c_corr_mean, mean, na.rm = TRUE)) %>%
  mutate(year.Type = paste(YearSampled, Type),
         dd14c = ifelse(YearSampled == 2011, 
                          atm.d14c.2011 - d14c_corr_mean, 
                          atm.d14c.2019 - d14c_corr_mean))
  
# # original plot
# all.14c.sum2 %>%
#   filter(Period == "inc") %>%
#   mutate(year.Type = paste(YearSampled, Type),
#          Eco = ifelse(Type == "F", "Forest", "Grassland")) %>%
#   ggplot(., aes(YearSampled, d14c_corr_mean, color = Type)) +
#   geom_point(aes(shape = Treatment), size = 3, stroke = 1.5) +
#   geom_path(aes(group = year.Type), 
#             arrow = arrow(type = "open", length = unit(.1, "inches"), ends = "first"),
#             show.legend = FALSE,
#             position = position_dodge(width = 2, preserve = "single"),
#             size = 1.2) +
#   geom_path(data = Datm, aes(Date, NHc14, linetype = Atmosphere), color = "black") +
#   geom_errorbar(
#     aes(ymin = d14c_corr_l, 
#         ymax = d14c_corr_u, 
#         color = Type), 
#     width = .5,
#     alpha = .3) +
#   scale_color_manual(name = 'Ecosystem',
#                      values = c('F' = '#a35513',
#                                 'G' = '#1361a3'),
#                      labels = c('Forest',
#                                 'Grassland')) +
#   scale_shape_manual(name = 'Treatment',
#                      values = c("air-dry" = 1, 
#                                 "air-dry + storage" = 0,
#                                 "control" = 16),
#                      labels = c("air-dry" = "air-dry", 
#                                 "air-dry + storage" = "air-dry + storage",
#                                 "control" = "control-1 (2011)\ncontrol-2 (2019)")) +
#   guides(linetype = guide_legend(title = NULL)) +
#   scale_x_continuous(limits = c(2010, 2020), breaks = c(2011, 2019)) +
#   scale_y_continuous(limits = c(-15, 100)) +
#   xlab('Year sampled') +
#   ylab(expression('mean '*Delta*''^14*'C-CO'[2]*' ()')) +
#   facet_grid(cols = vars(Eco)) +
#   theme_bw() +
#   theme(panel.grid = element_blank())

# Plot all points in background
all.14c.sum.df %>%
  filter(Experiment != "tme" & Period == "inc") %>%
  ggplot(., aes(YearSampled, d14c_corr_mean, color = Type)) +
  geom_point(data = arc.rewet.14c.inc.GM, 
             aes(YearSampled, d14c_corr_mean, shape = Treatment),
             size = 4, stroke = 2) +
  geom_point(aes(shape = Treatment), 
             size = 2, alpha = .5, 
             position = position_dodge(width = 1.11, preserve = "single"),
             show.legend = FALSE) +
  geom_errorbar(
    aes(ymin = d14c_corr_min, 
        ymax = d14c_corr_max), 
    width = .4, alpha = .3, 
    position = position_dodge(width = 1.2, preserve = "single")) +
  geom_path(data = arc.rewet.14c.inc.GM,
            aes(YearSampled, d14c_corr_mean,group = year.Type),
            arrow = arrow(type = "open", length = unit(.1, "inches"), ends = "first"),
            show.legend = FALSE,
            position = position_dodge(width = -1.4, preserve = "single"),
            size = 1.2) +
  geom_path(data = Datm, aes(Date, NHc14, linetype = Atmosphere), color = "black") +
  scale_color_manual(name = 'Ecosystem',
                     values = c('F' = '#a35513',
                                'G' = '#1361a3'),
                     labels = c('Forest',
                                'Grassland')) +
  scale_shape_manual(name = 'Treatment',
                     values = c("air-dry" = 1, 
                                "air-dry + storage" = 0,
                                "control" = 16),
                     labels = c("air-dry" = "air-dry", 
                                "air-dry + storage" = "air-dry + storage",
                                "control" = "control-1 (2011)\ncontrol-2 (2019)")) +
  guides(linetype = guide_legend(title = NULL)) +
  scale_x_continuous(limits = c(2010, 2020), breaks = c(2011, 2019)) +
  scale_y_continuous(limits = c(-15, 100)) +
  xlab('Year sampled') +
  ylab(expression('mean '*Delta*''^14*'C-CO'[2]*' ()')) +
  facet_grid(cols = vars(Type),
             labeller = labeller(Type = c("F" = "Forest", "G" = "Grassland"))) +
  theme_bw() +
  theme(panel.grid = element_blank())

# plot dd14c
all.14c.sum.df %>%
  filter(Experiment != "tme" & Period == "inc") %>%
  mutate(atm14c = ifelse(YearSampled == 2011, atm.d14c.2011, atm.d14c.2019)) %>%
  mutate(dd14c = atm14c - d14c_corr_mean,
         dd14c_min = atm14c -  d14c_corr_min,
         dd14c_max = atm14c -  d14c_corr_max) %>%
  ggplot(., aes(YearSampled, dd14c, color = Type)) +
  geom_hline(yintercept = 0) +
  geom_point(data = arc.rewet.14c.inc.GM, 
             aes(YearSampled, dd14c, shape = Treatment),
             size = 4, stroke = 2) +
  geom_point(aes(shape = Treatment), 
             size = 2, alpha = .5, 
             position = position_dodge(width = 1.11, preserve = "single"),
             show.legend = FALSE) +
  geom_errorbar(
    aes(ymin = dd14c_min, 
        ymax = dd14c_max), 
    width = .4, alpha = .3, 
    position = position_dodge(width = 1.2, preserve = "single")) +
  scale_color_manual(name = 'Ecosystem',
                     values = c('F' = '#a35513',
                                'G' = '#1361a3'),
                     labels = c('Forest',
                                'Grassland')) +
  scale_shape_manual(name = 'Treatment',
                     values = c("air-dry" = 1, 
                                "air-dry + storage" = 0,
                                "control" = 16),
                     labels = c("air-dry" = "air-dry", 
                                "air-dry + storage" = "air-dry + storage",
                                "control" = "control-1 (2011)\ncontrol-2 (2019)")) +
  guides(linetype = guide_legend(title = NULL)) +
  scale_x_continuous(limits = c(2010, 2020), breaks = c(2011, 2019)) +
  xlab('Year sampled') +
  ylab(expression('mean '*Delta*Delta*''^14*'C-CO'[2]*' ()')) +
  facet_grid(cols = vars(Type),
             labeller = labeller(Type = c("F" = "Forest", "G" = "Grassland"))) +
  theme_bw() +
  theme(panel.grid = element_blank())

# Plot Hainich points only, no arrows
all.14c.sum.df %>%
  filter(Experiment != "tme" & Period == "inc" & Site == "Hainich") %>%
  ggplot(., aes(YearSampled, d14c_corr_mean, color = Type)) +
  geom_point(data = arc.rewet.14c.inc.GM, 
             aes(YearSampled, d14c_corr_mean, shape = Treatment),
             size = 4, stroke = 2,
             position = position_dodge(width = -.5, preserve = "single")) +
  geom_point(aes(shape = Treatment), 
             size = 2, alpha = .5, 
             position = position_dodge(width = .5, preserve = "single"),
             show.legend = FALSE) +
  geom_errorbar(
    aes(ymin = d14c_corr_min, 
        ymax = d14c_corr_max), 
    width = .5, alpha = .3, 
    position = position_dodge(width = .6, preserve = "single")) +
  geom_path(data = Datm, aes(Date, NHc14, linetype = Atmosphere), color = "black") +
  scale_color_manual(name = 'Ecosystem',
                     values = c('F' = '#a35513',
                                'G' = '#1361a3'),
                     labels = c('Forest',
                                'Grassland')) +
  scale_shape_manual(name = 'Treatment',
                     values = c("air-dry" = 1, 
                                "air-dry + storage" = 0,
                                "control" = 16),
                     labels = c("air-dry" = "air-dry", 
                                "air-dry + storage" = "air-dry + storage",
                                "control" = "control-1 (2011)\ncontrol-2 (2019)")) +
  guides(linetype = guide_legend(title = NULL)) +
  scale_x_continuous(limits = c(2010, 2020), breaks = c(2011, 2019)) +
  scale_y_continuous(limits = c(-15, 100)) +
  xlab('Year sampled') +
  ylab(expression('mean '*Delta*''^14*'C-CO'[2]*' ()')) +
  facet_grid(cols = vars(Type),
             labeller = labeller(Type = c("F" = "Forest", "G" = "Grassland"))) +
  theme_bw() +
  theme(panel.grid = element_blank())

# plot trends over time by sample, treatment
all.14c.sum.df %>%
  filter(Experiment != "tme" & Period == "inc" & Site == "Hainich") %>%
  mutate(treat.ID = paste0(treat.bi, ID)) %>%
  ggplot(., aes(YearSampled, d14c_corr_mean, color = Type)) +
  geom_point(aes(shape = Treatment), 
             size = 2, alpha = .5) +
  geom_path(aes(group = treat.ID, linetype = treat.bi)) +
  geom_errorbar(
    aes(ymin = d14c_corr_min, 
        ymax = d14c_corr_max), 
    width = .5, alpha = .3) +
  geom_path(data = Datm, aes(Date, NHc14), color = "black") +
  scale_color_manual(name = 'Ecosystem',
                     values = c('F' = '#a35513',
                                'G' = '#1361a3'),
                     labels = c('Forest',
                                'Grassland')) +
  scale_shape_manual(name = 'Treatment',
                     values = c("air-dry" = 1, 
                                "air-dry + storage" = 0,
                                "control" = 16),
                     labels = c("air-dry" = "air-dry", 
                                "air-dry + storage" = "air-dry + storage",
                                "control" = "control-1 (2011)\ncontrol-2 (2019)")) +
  guides(linetype = guide_legend(title = NULL)) +
  scale_x_continuous(limits = c(2010, 2020), breaks = c(2011, 2019)) +
  scale_y_continuous(limits = c(-15, 100)) +
  xlab('Year sampled') +
  ylab(expression('mean '*Delta*''^14*'C-CO'[2]*' ()')) +
  facet_grid(cols = vars(Type),
             labeller = labeller(Type = c("F" = "Forest", "G" = "Grassland"))) +
  theme_bw() +
  theme(panel.grid = element_blank())

# plot SD instead of SE
arc.rewet.pooled <- bind_rows(lapply(
  split(
    all.14c.sum.df[all.14c.sum.df$Experiment != "tme" & all.14c.sum.df$Period == "inc", ],
    all.14c.sum.df[all.14c.sum.df$Experiment != "tme" & all.14c.sum.df$Period == "inc", "Experiment"]), function(exp) {
      bind_rows(lapply(split(exp, exp$Type), function(typ) {
        bind_rows(lapply(split(typ, typ$treat.bi), function(trt) {
          trt$d14c_corr_sd <- ifelse(
            is.na(trt$d14c_corr_sd), sqrt(mean.rep.var), trt$d14c_corr_sd)
          return(sample.decomp(n = rep(2, nrow(trt)),
                               sample.mean = trt$d14c_corr_mean,
                               sample.sd = trt$d14c_corr_sd,
                               include.sd = TRUE))
        }), .id = "treat.bi")
      }), .id = "Type")
    }), .id = "Experiment")
save(arc.rewet.pooled, file = "ts.pooled.Rdata")
# plot
arc.rewet.pooled[grep("pooled", row.names(arc.rewet.pooled)),] %>%
  mutate(YearSampled = ifelse(Experiment == "arc", 2011, 2019),
         Treatment = factor(ifelse(treat.bi == "treatment" & Experiment == "arc", 
                            "air-dry/rewet + storage", 
                            ifelse(treat.bi == "treatment" & Experiment == "rewet",
                            "air-dry/rewet", treat.bi)), 
                            levels = c(
                              "control", "air-dry/rewet + storage", "air-dry/rewet")),
         upper = sample.mean + sample.sd,
         lower = sample.mean - sample.sd) %>%
  ggplot(., aes(YearSampled, sample.mean, color = Type)) +
  geom_point(aes(shape = Treatment),
             size = 4, stroke = 1.5,
             position = position_dodge(width = 1)) +
  geom_errorbar(
    aes(ymin = upper, 
        ymax = lower,
        linetype = Treatment), 
    width = .4, alpha = .3, 
    position = position_dodge(width = 1),
    show.legend = FALSE) +
  geom_path(data = Datm, aes(Date, NHc14), color = "black") +
  scale_color_manual(name = 'Ecosystem',
                     values = c('F' = '#a35513',
                                'G' = '#1361a3'),
                     labels = c('Forest',
                                'Grassland')) +
  scale_shape_manual(name = 'Treatment',
                     values = c("air-dry/rewet" = 1, 
                                "air-dry/rewet + storage" = 0,
                                "control" = 16),
                     labels = c("air-dry/rewet" = "air-dry/rewet", 
                                "air-dry/rewet + storage" = "air-dry/rewet + storage",
                                "control" = "control-1 (2011)\ncontrol-2 (2019)")) +
  scale_linetype_manual(values = c("atmosphere" = 1,
                                   "air-dry/rewet" = 1, 
                                   "air-dry/rewet + storage" = 1,
                                   "control" = 1)) +
  guides(linetype = "none") +
  scale_x_continuous(limits = c(2010, 2020), breaks = c(2011, 2019)) +
  scale_y_continuous(limits = c(-15, 105)) +
  xlab('Year sampled') +
  ylab(expression('mean '*Delta*''^14*'C-CO'[2]*' ()')) +
  facet_grid(cols = vars(Type),
             labeller = labeller(Type = c("F" = "Forest", "G" = "Grassland"))) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        text = element_text(size = 14))
```
>**Fig `r {fig.n + 4}`. Time series of control and treatment $\Delta$^14^C-CO~2~ (Experiments 1 and 2)**
>*Caption:* Filled circles show $\Delta$^14^C-CO~2~ observed for both control-1 and control-2 samples (2011 and 2019 points, respectively). Open symbols show $\Delta$^14^C-CO~2~ observed for treament samples: open squares = air-dry + storage treatment, Experiment 1; open circles = air-dry only treatment, Experiment 2. Points are means and error bars show the pooled standard deviation. The black line shows $\Delta$^14^C of the atmosphere.

>**Supplemental Fig `r {sfig.n + 2}`. Time series of control and treatment $\delta$^13^C-CO~2~ (Experiments 1 and 2)**

>*Caption:* Filled circles show $\delta$^13^C-CO~2~ observed for control samples, while open symbols show $\delta$^13^C-CO~2~ observed for treament samples (open squares = air-dry + storage treatment, Experiment 1, 2011; open circles = air-dry only treatment, Experiment 2, 2019). Points are means and error bars show standard deviations. 

*Effect of cumulative respired carbon on ^14^C-CO~2~*

```{r ctl-trt-stats, include = FALSE}
# calculate difs ctl-trt, look at effect of C resp
arc.rewet.inc.14c$treat.bi <- ifelse(arc.rewet.inc.14c$Treatment == "control", "control", "treatment")
arc.rewet.inc.14c.sum <- arc.rewet.inc.14c %>%
  group_by(ID, Experiment, Type, treat.bi) %>%
  add_count() %>%
  summarise_if(.predicate = function(x) is.numeric(x),
               .funs = list(mean = "mean", sd = "sd"))
  
arc.rewet.inc.14c.sum$n_sd <- NULL

j.vars <- c("ID", "Experiment", "Type", "d14c_corr_mean", "mgCO2.C_gC_mean", "d14c_corr_sd", "mgCO2.C_gC_sd")
arc.rewet.14c.ctl.trt <- left_join(
  arc.rewet.inc.14c.sum[arc.rewet.inc.14c.sum$treat.bi == "control", j.vars],
  arc.rewet.inc.14c.sum[arc.rewet.inc.14c.sum$treat.bi != "control", j.vars],
  by = c(j.vars[1:3]),
  suffix = c(".ctl", ".trt"))


arc.rewet.14c.ctl.trt$d14c_dif <- arc.rewet.14c.ctl.trt$d14c_corr_mean.ctl - arc.rewet.14c.ctl.trt$d14c_corr_mean.trt
arc.rewet.14c.ctl.trt$c_dif <- arc.rewet.14c.ctl.trt$mgCO2.C_gC_mean.ctl - arc.rewet.14c.ctl.trt$mgCO2.C_gC_mean.trt

# save Rdat
save(arc.rewet.14c.ctl.trt, file = "./arc.rewet.14c.ctl.trt.Rdata")

## Look at effect of c_dif by experiment and interaction with type
# air-dry + storage
summary(lm(d14c_dif ~ c_dif * Type, arc.rewet.14c.ctl.trt[arc.rewet.14c.ctl.trt$Experiment == "arc", ]))
# effect of c_dif on d14c_dif ns; interaction also ns

# rewet
summary(lm(d14c_dif ~ c_dif * Type, arc.rewet.14c.ctl.trt[arc.rewet.14c.ctl.trt$Experiment == "rewet", ]))
# effect of c_dif marginally significant (p = 0.0632)

# Look at effect of c_dif and interaction with type overall
summary(lm(d14c_dif ~ Type * c_dif * Experiment, arc.rewet.14c.ctl.trt))
# no significance of c_dif overall
rm(j.vars)
```

```{r plot-14c-by-c-resp}
# 10. Plot $\Delta$^14^C against proportion of soil C respired by experiment, land cover, and sampling period.

# HEW22 control pre point excluded to make figure more readable 
all.14c.sum.df %>%
  filter(Experiment != "tme") %>%
  mutate(ID.treat = paste(ID, Treatment)) %>%
  mutate(ID.treat.Period = paste(ID.treat, Period)) %>%
  filter(ID.treat.Period != "HEW22 control pre") %>%
  mutate(treat.Period = paste(treat.bi, Period),
         Eco = ifelse(Type == "F", "Forest", "Grassland")) %>%
  ggplot(., aes(mgCO2.C_gC_mean, d14c_corr_mean, color = Type, shape = treat.Period)) +
  geom_point(size = 3) +
  geom_path(aes(group = ID.treat)) +
  geom_errorbar(
    aes(ymin = d14c_corr_min, 
        ymax = d14c_corr_max, 
        color = Type), 
    width = .25) +
  geom_errorbarh(
    aes(xmin = mgCO2.C_gC_min, 
        xmax = mgCO2.C_gC_max, 
        color = Type), 
    height = .9) +
  scale_color_manual(name = 'Ecosystem',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  scale_shape_manual(name = 'Sampling period (treatment)',
                     values =c(16, 17, 1, 2),
                     labels = c("control pre" = 'pre-incubation\n(control)', 
                                "control inc" = 'equilibrium respiration\n(control)',
                                "treatment pre" = 'pre-incubation\n(air-dry/air-dry + storage)',
                                "treatment inc" = 'equilibrium respiration\n(air-dry/air-dry + storage)')) +
  facet_grid(cols = vars(Eco), rows = vars(YearSampled)) +
  xlab(expression('Respired C (mg CO'[2]*'-C'*' g soil C'^-1*')')) +
  ylab(expression('mean '*Delta*''^14*'C ()')) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.key.height=unit(.8, "cm"))
```

>**Supplementary Fig. `r {sfig.n + 3}`. Change in ^14^C-CO~2~ in relation to cumulative soil carbon respired**

>*Caption:* Note that pre-incubation $\Delta$^14^C was not measured for the control-1 samples in 2011. Limits exclude outlier point (HEW22 control-2, pre-incubation) for improved legibility. Points are means, error bars show min and max of duplicate samples.

```{r plot 14c by c resp with outlier, include = FALSE}
# Code for a third plot with outlier data
# average control and treatment data
all.14c.sum3 <- select(all.14c.ls[[2]], Type, Period, Treatment, Experiment, YearSampled, d14c_corr, mgCO2.C_gC) %>%
  group_by(Type, Period, Treatment, Experiment, YearSampled) %>%
  add_count() %>%
  summarize_all(list(mean = mean, sd = sd), na.rm = TRUE) %>%
  mutate(d14c_corr_se = d14c_corr_sd/n_mean,
         d14c_corr_u = d14c_corr_mean + d14c_corr_se*2,
         d14c_corr_l = d14c_corr_mean - d14c_corr_se*2,
         mgCO2.C_gC_se = mgCO2.C_gC_sd/n_mean,
         mgCO2.C_gC_u = mgCO2.C_gC_mean + mgCO2.C_gC_se*2,
         mgCO2.C_gC_l = mgCO2.C_gC_mean - mgCO2.C_gC_se*2) 

all.14c.sum3$treat.bi <- ifelse(all.14c.sum3$Treatment == "control", "control","treatment")
all.14c.sum3 %>%
  mutate(treat.Period = paste(treat.bi, Period)) %>%
  ggplot(., aes(mgCO2.C_gC_mean, d14c_corr_mean, color = Type, shape = treat.Period)) +
  geom_point(size = 3) +
  geom_path(aes(group = Treatment)) +
  geom_errorbar(
    aes(ymin = d14c_corr_u, 
        ymax = d14c_corr_l, 
        color = Type), 
    width = .25) +
  geom_errorbarh(
    aes(xmin = mgCO2.C_gC_u, 
        xmax = mgCO2.C_gC_l, 
        color = Type), 
    height = .9) +
  scale_color_manual(name = 'Ecosystem',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  scale_shape_manual(name = 'Sampling period (treatment)',
                     values =c(16, 17, 1, 2),
                     labels = c("control pre" = 'pre-incubation\n(control)', 
                                "control inc" = 'equilibrium respiration\n(control)',
                                "treatment pre" = 'pre-incubation\n(air-dry/air-dry + storage)',
                                "treatment inc" = 'equilibrium respiration\n(air-dry/air-dry + storage)')) +
  facet_grid(cols = vars(Type), rows = vars(YearSampled)) +
  xlab(expression('Respired C (mgCO'[2]*'-C'*' g soil C'^-1*')')) +
  ylab(expression('mean '*Delta*''^14*'C ()')) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.key.height=unit(.8, "cm"))
```

# Discussion
```{r conceptual-model-calcs, include = FALSE}
# atm 14C
Datm <- rbind(graven, future14C)
save(Datm, file = "Datm.RData")
yrs <- Datm$Date

# helper fxs
lambda <- (1/8267)

k <- function(Fm) {
  (Fm*lambda)/(1-Fm)
}
fm <- function(k){
  k/(k+lambda)
}
fm_14c <- function(fm, date) {
  (fm*exp(lambda*(1950 - date)) - 1)*1000
}

# calculate resp flux points at 1991 and 2019 and put in data frame
arc.rewet.rsp <- arc.rewet.14c[arc.rewet.14c$Site == "Hainich" & 
                           arc.rewet.14c$Type == "F" & 
                           arc.rewet.14c$treat.bi == "control", 
                         "d14c_corr_inc"]
yrs.smp <- arc.rewet.14c[arc.rewet.14c$Site == "Hainich" & 
                                 arc.rewet.14c$Type == "F" & 
                                 arc.rewet.14c$treat.bi == "control", 
                               "YearSampled"] + .5
obs.rsp <- data.frame(time = yrs.smp,
                      resp = arc.rewet.rsp)

arc.rewet.rsp.t <- arc.rewet.14c[arc.rewet.14c$Site == "Hainich" & 
                                   arc.rewet.14c$Type == "F" & 
                                   arc.rewet.14c$treat.bi == "treatment", 
                                 "d14c_corr_inc"]
yrs.smp.t <- arc.rewet.14c[arc.rewet.14c$Site == "Hainich" & 
                           arc.rewet.14c$Type == "F" & 
                           arc.rewet.14c$treat.bi == "treatment", 
                         "YearSampled"] + .5
obs.rsp.t <- data.frame(time = yrs.smp.t,
                        resp = arc.rewet.rsp.t)
#####
# mod fit fxns
#####
cgam <- function(pars, In = 1, C = 38.96104, frc = .1) {
  
  # cstock
  Cfast <- frc * C
  Cslow <- C - Cfast

  # define pars
  fast.fm <- fm(pars[1])
  slow.fm <- fm(pars[2])
  F0_Delta14C <- fm_14c(c(fast.fm, slow.fm), date = c(1900, 1900))
  gam <- (C * Cfast * pars[1])/(C * Cfast * pars[1] + C * Cslow * pars[2])
  
  # model matrix
  A <- -diag(pars[1:2])
  
  # steady-state C stocks
  ss.cstock <- (-1 * solve(A) %*% c(In * gam, In * (1 - gam)))
  
  # return
  return(list(gam = gam, ss.cstock = ss.cstock))
}

modFun_2pp <- function(pars, In = 1, lag = 0, pass = TRUE, out = "modFit", 
                       gam = FALSE, yrs = Datm$Date) {
  # constants
  cstock <- 38.96104 # forest C pool, total (mean Hainich forest C  = 24 g kg^-1); nb @ ss = sum(-1*solve(A)%*%c(f.in*f.gam, f.in*(1-f.gam)))
  f.in <- In # inputs (12.5=mean equilibrium C flux, annual extrapolation, gC yr-1, 2011 incubation data); try much less...
  f.frc <- .1 # C-stock partitioning coefficient (from Schrumpf 2013, Hainich 0-5, free light)
  f.Cfast <- f.frc * cstock
  f.Cslow <- cstock - f.Cfast
  
  # define pars
  f.kfast <- pars[1] # unknown
  f.kslow <- pars[2] # unknown
  f.fast.fm <- fm(f.kfast)
  f.slow.fm <- fm(f.kslow)
  f.F0_Delta14C <- fm_14c(c(f.fast.fm, f.slow.fm), date = c(1900, 1900))
  f.gam <- (cstock*f.Cfast*f.kfast)/(cstock*f.Cfast*f.kfast+cstock*f.Cslow*f.kslow) # input partitioning coefficient (proportion to fast pool, function of ks)
  if (gam) {
    f.gam <- pars[3]
  }
  
  # model matrix
  A <- -diag(pars[1:2])
  
  # steady-state C stocks
  ss.cstock <- (-1 * solve(A) %*% c(In * f.gam, In * (1 - f.gam)))
  
  # time index
  ix.t <- c((lag + 1):length(yrs))
  
  # model
  mod <- TwopParallelModel14(t = yrs[ix.t],
                             ks = c(f.kfast, f.kslow),
                             C0 = c(f.Cfast, f.Cslow),
                             F0_Delta14C = f.F0_Delta14C,
                             In = f.in,
                             gam = f.gam,
                             inputFc = Datm,
                             lag = lag,
                             pass = pass)

  # get mod values
  C14m <- getF14C(mod)
  C14p <- getF14(mod)
  C14r <- getF14R(mod)
  Ctot <- getC(mod)
    
  if(out == "modFit") {
    # dataframe for modFit fx
    return(data.frame(
      time = Datm$Date[ix.t], 
      resp = C14r))
  } else {
    # data frame for plotting
    return(data.frame(
      years = rep(Datm$Date[ix.t], 5),
      d14C = c(C14p[,1], 
               C14p[,2], 
               C14m,
               C14r,
               Datm$NHc14[ix.t]),
      pool = rep(c("fast", "slow", "bulk C", "respiration", "atm"), each = nrow(C14p))))
  }
  }

# Use median values for ks and lag value from Schrumpf 2015
schrumpf.2015 <- modFun_2pp(par = c(1/4, 1/115), 
                            In = 1,
                            lag = 8,
                            out = "plot",
                            gam = FALSE)
# change factor levels
schrumpf.2015$pool <- factor(schrumpf.2015$pool, 
                             levels = c("atm", "bulk C", "respiration", "fast", "slow"))
cgam(c(1/4, 1/115))

# also for medians
par.25.25 <- c(1, 1/80)
par.25.75 <- c(1, 1/200)
par.75.25 <- c(1/7, 1/80)
par.75.75 <- c(1/7, 1/200)
L8.25.25 <- modFun_2pp(par = par.25.25,
                       In = 1,
                       lag = 8,
                       out = "plot",
                       gam = FALSE)
L8.25.75 <- modFun_2pp(par = par.25.75,
                       In = 1,
                       lag = 8,
                       out = "plot",
                       gam = FALSE)
L8.75.25 <- modFun_2pp(par = par.75.25,
                       In = 1,
                       lag = 8,
                       out = "plot",
                       gam = FALSE)
L8.75.75 <- modFun_2pp(par = par.75.75,
                       In = 1,
                       lag = 8,
                       out = "plot",
                       gam = FALSE)
schrumpf.2015$q25.25 <- L8.25.25$d14C - schrumpf.2015$d14C
schrumpf.2015$q25.75 <- L8.25.75$d14C - schrumpf.2015$d14C
schrumpf.2015$q75.25 <- L8.75.25$d14C - schrumpf.2015$d14C
schrumpf.2015$q75.75 <- L8.75.75$d14C - schrumpf.2015$d14C

# Observed respiration
# ctl
obs.rsp.m <- obs.rsp %>%
  group_by(time) %>%
  summarize(across(resp, .fns = c(mean = mean, sd = sd))) %>%
  rename(d14C = resp_mean) %>%
  mutate(years = time - .25,
         Treatment = "control")
save(obs.rsp.m, file = "/Users/jeff/arc-inc/src/obs.rsp.m.RData")
# trt
obs.rsp.t.m <- obs.rsp.t %>%
  group_by(time) %>%
  summarize(across(resp, .fns = c(mean = mean, sd = sd))) %>%
  rename(d14C = resp_mean) %>%
  mutate(years = time + .25,
         Treatment = ifelse(time == "2011.5", "air-dry/rewet + storage", "air-dry/rewet"))
save(obs.rsp.t.m, file = "/Users/jeff/arc-inc/src/obs.rsp.t.m.RData")

# plot medians w/ potential error ranges
schrumpf.2015 %>%
  filter(pool != "bulk C") %>%
  ggplot(., aes(years, d14C)) +
  geom_path(data = L8.25.25[which(L8.25.25$pool != "bulk C"), ],
            aes(years, d14C, color = pool), show.legend = FALSE, linetype = "dotted") +
  geom_path(data = L8.25.75[which(L8.25.75$pool != "bulk C"), ],
            aes(years, d14C, color = pool), show.legend = FALSE, linetype = "dashed") +
  geom_path(data = L8.75.25[which(L8.75.25$pool != "bulk C"), ],
    aes(years, d14C, color = pool), show.legend = FALSE, linetype = "dotdash") +
  geom_path(data = L8.75.75[which(L8.75.75$pool != "bulk C"), ],
    aes(years, d14C, color = pool), show.legend = FALSE, linetype = "longdash") +
  geom_path(aes(color = pool)) +
  geom_point(data = obs.rsp.m, aes(years, d14C, shape = Treatment), 
             color = "#FFC107", size = 3) +
  geom_point(data = obs.rsp.t.m, aes(years, d14C, shape = Treatment), 
             color = "#FFC107", size = 3) +
  geom_errorbar(data = obs.rsp.m, 
                aes(ymin = d14C - resp_sd, ymax = d14C + resp_sd),
                color = "#FFC107", width = .2) +
  geom_errorbar(data = obs.rsp.t.m, 
                aes(ymin = d14C - resp_sd, ymax = d14C + resp_sd),
                color = "#FFC107", width = .2) +
  scale_color_manual(
    name = "Model pool",
    values = c("atm" = "black",
               "fast" = "#D81B60", 
               "respiration" = "#FFC107", 
               "slow" = "#1E88E5"),
    labels = c("atm" = "atmosphere\n(measured)",
               "fast" = "fast",
               "slow" = "slow")) +
  scale_fill_manual(
    values = c("atm" = 8,
               "fast" = "#D81B60", 
               "respiration" = "#FFC107", 
               "slow" = "#1E88E5")) +
  scale_shape_manual(
    values = c("control" = 16,
               "air-dry/rewet" = 1, 
               "air-dry/rewet + storage" = 0),
    labels = c("control" = "control-1 (2011)\ncontrol-2 (2019)",
               "air-dry/rewet" = "air-dry/rewet", 
               "air-dry/rewet + storage" = "air-dry/rewet + storage")) +
  scale_x_continuous(limits = c(2000, 2022)) +
  scale_y_continuous(limits = c(-100, 180)) +
  xlab("Year") +
  ylab(expression(''*Delta*''^14*'C ()')) +
  theme_bw() +
  theme(panel.grid = element_blank())

# Save w/ error bars
schrumpf.2015$ymin <- ifelse(schrumpf.2015$pool == "respiration", 
                             schrumpf.2015$d14C + schrumpf.2015$q25.75, NA)
schrumpf.2015$ymax <- ifelse(schrumpf.2015$pool == "respiration", 
                             schrumpf.2015$d14C + schrumpf.2015$q75.25, NA)
save(schrumpf.2015, file = "schrumpf.2015.mod.RData")
```

```{r fig6-schrumpf15-median}
# plot whole bomb period
schrumpf.2015 %>%
  filter(pool != "bulk C") %>%
  ggplot(., aes(years, d14C)) +
  geom_path(aes(color = pool)) +
  geom_point(data = obs.rsp.m, aes(years, d14C, shape = Treatment), 
             color = "#FFC107", size = 3) +
  geom_point(data = obs.rsp.t.m, aes(years, d14C, shape = Treatment), 
             color = "#FFC107", size = 3) +
  geom_errorbar(data = obs.rsp.m, 
                aes(ymin = d14C - resp_sd, ymax = d14C + resp_sd),
                color = "#FFC107", width = .2) +
  geom_errorbar(data = obs.rsp.t.m, 
                aes(ymin = d14C - resp_sd, ymax = d14C + resp_sd),
                color = "#FFC107", width = .2) +
  scale_color_manual(
    name = "Model pool",
    values = c("atm" = "black",
               "fast" = "#D81B60", 
               "respiration" = "#FFC107", 
               "slow" = "#1E88E5"),
    labels = c("atm" = "atmosphere\n(measured)",
               "fast" = "fast",
               "slow" = "slow")) +
  scale_fill_manual(
    values = c("atm" = 8,
               "fast" = "#D81B60", 
               "respiration" = "#FFC107", 
               "slow" = "#1E88E5")) +
  scale_shape_manual(
    values = c("control" = 16,
               "air-dry/rewet" = 1, 
               "air-dry/rewet + storage" = 0),
    labels = c("control" = "control-1 (2011)\ncontrol-2 (2019)",
               "air-dry/rewet" = "air-dry/rewet", 
               "air-dry/rewet + storage" = "air-dry/rewet + storage")) +
  scale_x_continuous(limits = c(1950, 2022)) +
  xlab("Year") +
  ylab(expression(''*Delta*''^14*'C ()')) +
  theme_bw() +
  theme(panel.grid = element_blank())
```
>**Fig `r {(fig.n + 5)}`a. Trajectories of $\Delta$^14^C over time in soil C and respired CO~2~ for a 2-pool model fit to the Hainich-Dn forest sites in relation to atmospheric $\Delta$^14^C**

>*Caption:* Modeled curves derived from a two-pool parallel model parameterized with data from Schrumpf et al. (2015): (*k~fast~* = 1/4) and a more slowly cycling pool (*k~slow~* = 1/115). Gold points show observed $\Delta$^14^C-CO~2~ from this study. Solid symbols show control samples, open symbols show treatment samples. Atmospheric $\Delta$^14^C data up to the year 2015 are from Graven et al. (2017), while data points beyond 2015 use the extrapolation method from Sierra (2018). All atmospheric radiocarbon data is for the northern hemisphere (zone 2).

```{r conceptual-model-fig1b}
# plot zoom w/ 25th-75th quartile error on resp
schrumpf.2015 %>%
  filter(pool != "bulk C") %>%
  ggplot(., aes(years, d14C)) +
  geom_ribbon(aes(ymin = ymin, ymax = ymax, fill = pool),
              alpha = .1, show.legend = FALSE) +
  geom_path(aes(color = pool)) +
  geom_point(data = obs.rsp.m, aes(years, d14C, shape = Treatment), 
             color = "#FFC107", size = 3) +
  geom_point(data = obs.rsp.t.m, aes(years, d14C, shape = Treatment), 
             color = "#FFC107", size = 3) +
  geom_errorbar(data = obs.rsp.m, 
                aes(ymin = d14C - resp_sd, ymax = d14C + resp_sd),
                color = "#FFC107", width = .2) +
  geom_errorbar(data = obs.rsp.t.m, 
                aes(ymin = d14C - resp_sd, ymax = d14C + resp_sd),
                color = "#FFC107", width = .2) +
  scale_color_manual(
    name = "Model pool",
    values = c("atm" = "black",
               "fast" = "#D81B60", 
               "respiration" = "#FFC107", 
               "slow" = "#1E88E5"),
    labels = c("atm" = "atmosphere\n(measured)",
               "fast" = "fast",
               "slow" = "slow")) +
  scale_fill_manual(
    values = c("atm" = 8,
               "fast" = "#D81B60", 
               "respiration" = "#FFC107", 
               "slow" = "#1E88E5")) +
  scale_shape_manual(
    values = c("control" = 16,
               "air-dry/rewet" = 1, 
               "air-dry/rewet + storage" = 0),
    labels = c("control" = "control-1 (2011)\ncontrol-2 (2019)",
               "air-dry/rewet" = "air-dry/rewet", 
               "air-dry/rewet + storage" = "air-dry/rewet + storage")) +
  scale_x_continuous(limits = c(2000, 2022)) +
  scale_y_continuous(limits = c(-25, 180)) +
  xlab("Year") +
  ylab(expression(''*Delta*''^14*'C ()')) +
  theme_bw() +
  theme(panel.grid = element_blank())
```

>**Fig `r {(fig.n + 5)}`b. Potential shifts in $\Delta$^14^C of respired CO~2~ in response to treatment**

>*Caption:* Zooming in on the study period, we can see that the treatment sample $\Delta$^14^C-CO~2~ shifts in the direction of the slowly cycling soil C pool (blue line), indicating increased contribution from this pool to respiration following air-drying and rewetting. Note that the increased contribution from the slow cycling pool leads to depletion in $\Delta$^14^C-CO~2~ relative to the controls in 2011, but enrichment in 2019 due to the crossing of the slow and fast cycling pool curves in 2016. Shaded ribbon around the respiration curve shows the model parameterized with the 75th quartile estimates for *k~fast~* (1/7) and *k~slow~* (1/200).


```{r rep-var}
# calculate variability between reps
# calc difference between reps and compare to dif between ctl and treat
e12.d14 <- all.14c.ls[[2]]
e12.d14.inc <- e12.d14[e12.d14$Period == "inc", ]

# split by Experiment and Type to compare with ctl-trt difs
e12.d14.inc.ls <- lapply(split(e12.d14.inc, e12.d14.inc$Experiment), function(x) {
    lapply(split(x, x$Type), function(y) {
      y <- lapply(split(y, y$Treatment), function(z) {
        z <- lapply(split(z, z$ID), function(m) {
          if(nrow(m) > 1) {
            m <- abs(m[2, "d14c"] - m[1, "d14c"])
          } else {
            m <- NA
          }
        })
        mean <- mean(unlist(z))
        sd <- sd(unlist(z))
        var <- var(unlist(z))
        n <- length(z)
        z.ls <- list(mean = mean,
                     sd = sd,
                     var = var,
                     n = n,
                     se = sd/sqrt(n))
        })
  })
})

# reduce list
var.names <- c("Experiment", "Type", "Treatment", "dispersion")
values <- unlist(e12.d14.inc.ls)
e12.d14.sdif.mat <- matrix(ncol = length(var.names) + 1, 
                           nrow = length(values),
                           dimnames = list(NULL, c(var.names, "value")))
for(i in seq_along(var.names)) {
  e12.d14.sdif.mat[, i] <- unlist(lapply(strsplit(names(values), "\\."), "[[", i))
}

e12.d14.sdif <- data.frame(e12.d14.sdif.mat)
e12.d14.sdif$value <- values
 
# widen and remove NAs
e12.d14.sdif <- pivot_wider(e12.d14.sdif,
                            names_from = "dispersion",
                            values_from = "value")
e12.d14.sdif <- e12.d14.sdif[!is.na(e12.d14.sdif$mean), ]

# make further reduced df (collapse treatment)
e12.d14.sdif2 <- data.frame(Experiment = rep(c("arc", "rewet"), each = 2),
                            Type = rep(c("F", "G"), 2))
e12.d14.sdif2$mean <- unlist(c(e12.d14.sdif[e12.d14.sdif$Experiment == "arc" &e12.d14.sdif$Type == "F", "mean"],
                        e12.d14.sdif[e12.d14.sdif$Experiment == "arc" &e12.d14.sdif$Type == "G", "mean"],
                        mean(unlist(e12.d14.sdif[e12.d14.sdif$Experiment == "rewet" &e12.d14.sdif$Type == "F", "mean"])),
                        mean(unlist(e12.d14.sdif[e12.d14.sdif$Experiment == "rewet" &e12.d14.sdif$Type == "G", "mean"]))))

# function for calculating std err of difs
std.err.dif <- function(sds, ns) {sqrt(sds[1]^2/ns[1] + sds[2]^2/ns[2])}
# function for calculating mean std dev
std.dev.mean <- function(vars) {sqrt(mean(vars))}

e12.d14.sdif2$se <- unlist(c(e12.d14.sdif[e12.d14.sdif$Experiment == "arc" &e12.d14.sdif$Type == "F", "se"],
                        e12.d14.sdif[e12.d14.sdif$Experiment == "arc" &e12.d14.sdif$Type == "G", "se"],
                      std.err.dif(unlist(e12.d14.sdif[e12.d14.sdif$Experiment == "rewet" &e12.d14.sdif$Type == "F", "sd"]),
                                  unlist(e12.d14.sdif[e12.d14.sdif$Experiment == "rewet" &e12.d14.sdif$Type == "F", "n"])),
                      std.err.dif(unlist(e12.d14.sdif[e12.d14.sdif$Experiment == "rewet" &e12.d14.sdif$Type == "G", "sd"]),
                                  unlist(e12.d14.sdif[e12.d14.sdif$Experiment == "rewet" &e12.d14.sdif$Type == "G", "n"]))))

# calc mean variance
e12.d14.sdif2$var <- unlist(c(e12.d14.sdif[e12.d14.sdif$Experiment == "arc" &e12.d14.sdif$Type == "F", "var"],
                        e12.d14.sdif[e12.d14.sdif$Experiment == "arc" &e12.d14.sdif$Type == "G", "var"],
                      mean(unlist(e12.d14.sdif[e12.d14.sdif$Experiment == "rewet" &e12.d14.sdif$Type == "F", "var"])),
                      mean(unlist(e12.d14.sdif[e12.d14.sdif$Experiment == "rewet" &e12.d14.sdif$Type == "G", "var"]))))

# # mean std dev for forests
# std.dev.mean(c(e12.d14.sdif2[e12.d14.sdif2$Type == "F", "var"]))
# # mean std dev for grasslands
# std.dev.mean(c(e12.d14.sdif2[e12.d14.sdif2$Type == "G", "var"]))

# compare to ctl treat dif
e12.14c.ctl.trt <- all.14c.ctl.trt %>%
  select(Experiment, Type, ctl.trt.d14) %>%
  group_by(Experiment, Type) %>%
  add_tally() %>%
  summarize_all(list(mean = mean, sd = sd, var = var), na.rm = TRUE) %>%
  mutate(ctl.trt.se = ctl.trt.d14_sd/n_mean)

# combine
var.cf <- cbind.data.frame(e12.14c.ctl.trt[1:4, ], e12.d14.sdif2[ , c("mean", "se")]) %>%
  select(c(Experiment, Type, n_mean, ctl.trt.d14_mean, ctl.trt.se, mean, se))
# knitr::kable(var.cf[, c("Experiment", "Type", "ctl.trt.d14_mean", "ctl.trt.se", "mean", "se")],
#              col.names = c("Experiment", "Type", "Ctl-Trt mean dif", "Ctl-Trt SE", "Replicate mean dif", "Replicate SE"),
#              caption = "Comparison of treatment differences versus variability among replicates",
#              align = "c",
#              digits = 1)
```

```{r delta-delta-plot, include = FALSE}
# calculate dd14c
dd14c.sum <- e12.d14.inc
dd14c.sum$atm14c <- ifelse(dd14c.sum$YearSampled == 2011,
                             Datm[which(Datm$Date == 2011.5), "NHc14"],
                             Datm[which(Datm$Date == 2019.5), "NHc14"])
dd14c.sum$dd14c <- dd14c.sum$d14c_corr - dd14c.sum$atm14c
dd14c.sum <- dd14c.sum %>% 
  select(Experiment, Treatment, Type, dd14c) %>%
  group_by(Experiment, Treatment, Type) %>%
  add_tally() %>%
  summarize_all(list(mean = mean, sd = sd), na.rm = TRUE) %>%
  mutate(dd14c_se = dd14c_sd/n_mean)

dd14c.sum %>%
  mutate(treat.bi = ifelse(Treatment != "control", "treatment", "control"),
         dd14c_min = dd14c_mean - dd14c_se*2,
         dd14c_max = dd14c_mean + dd14c_se*2) %>%
  ggplot(., aes(treat.bi, dd14c_mean, color = Type)) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = dd14c_min, 
        ymax = dd14c_max, 
        color = Type), 
    width = .1) +
  scale_color_manual(name = 'Ecosystem',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  facet_grid(cols = vars(Experiment)) + 
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```


# Change in soil moisture content with moisture adjustment

```{r delta-moisture-calc}
# compare the change in moisture with control samples versus treatment samples
whcFresh <- merge(archive.ctl.ts.wide[archive.ctl.ts.wide$Experiment == "arc", c("ID", "whcFresh")],
                  jarinfo.rewet[jarinfo.rewet$Treatment == "control", c("ID", "whcFresh")],
                  by = "ID",
                  all.x = TRUE,
                  suffixes = c("_2011", "_2019"))
whcFresh$deltaWHC_2011 <- .6-whcFresh$whcFresh_2011
whcFresh$deltaWHC_2019 <- .6-whcFresh$whcFresh_2019
whcFresh$Type <- ifelse(grepl("G", whcFresh$ID), "G", "F")

# sumamrize
whcFresh_sum <- whcFresh %>%
  select(Type, deltaWHC_2011, deltaWHC_2019, whcFresh_2011, whcFresh_2019) %>%
  group_by(Type) %>%
  add_tally() %>%
  summarize_all(list(mean = mean, sd = sd), na.rm = TRUE) %>%
  mutate(delta_se_11 = deltaWHC_2011_sd/n_mean,
         delta_se_19 = deltaWHC_2019_sd/n_mean,
         fresh_se_11 = whcFresh_2011_sd/n_mean,
         fresh_se_19 = whcFresh_2019_sd/n_mean)

# switch whcFresh to long
whcFresh.2011 <- whcFresh[ , c("ID", "whcFresh_2011", "deltaWHC_2011")]
names(whcFresh.2011) <- c("ID", "whcFresh", "deltaWHC")
whcFresh.2011$Experiment <- "arc"
whcFresh.2019 <- whcFresh[ , c("ID", "whcFresh_2019", "deltaWHC_2019")]
names(whcFresh.2019) <- c("ID", "whcFresh", "deltaWHC")
whcFresh.2019$Experiment <- "rewet"
whcFresh.2019 <- whcFresh.2019[which(complete.cases(whcFresh.2019)), ]
whcFresh.long <- rbind(whcFresh.2011, whcFresh.2019)

# stats
whc.ctl.trt <- merge(arc.rewet.14c.ctl.trt, whcFresh.long, by = c("ID", "Experiment"))
save(whc.ctl.trt, file = "whc.ctl.trt.Rdata")
ggplot(whc.ctl.trt, aes(whcFresh, d14c_dif, color = Type, shape = Experiment)) +
  geom_hline(yintercept = 0) +
  geom_point(size = 3) +
  scale_color_manual(name = 'Ecosystem',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  theme_bw() +
  theme(panel.grid = element_blank())

# summary(lm(d14c_dif ~ whcFresh * Type, whc.ctl.trt))

# # calculate atm 14C change
# slopes <- data.frame(Datm[Datm$Date > 2000 & Datm$Date < 2016, ],
#                      slope = NA)
# for(i in seq_along(Datm[Datm$Date > 2000 & Datm$Date < 2016, "Date"])) {
#   if(i != nrow(slopes)) {
#     slopes$slope[i] <- (Datm[Datm$Date > 2000, "NHc14"][i+1]-Datm[Datm$Date > 2000, "NHc14"][i])/(Datm[Datm$Date > 2000, "Date"][i+1]-Datm[Datm$Date > 2000, "Date"][i])
#   }
# }
# mean(slopes$slope, na.rm = T)
```

>**Supplemental Fig `r {(sfig.n + 4)}`. Change in $\Delta$^14^C-CO~2~ (control - treatment) relative to field moisture**

>*Caption:* Data are from Experiment 1 ("arc") and Experiment 2 ("rewet"). All samples were moisture-adjusted prior to incubation, but control samples were adjusted from field moisture, "whcFresh" (percent of WHC), whereas treatment samples were moisture adjusted after air-drying, i.e. at approximately 0% of WHC.