---
title: "Archived Soil Incubations Manuscript Figures"
author: "J. Beem-Miller"
date: "30 Jul 2020"
output:
  pdf_document:
    latex_engine: xelatex
  html_notebook:
    css: "custom.css"
    toc: yes
    toc_depth: 2
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
header_includes:
- \usepackage[utf8]{inputenc}
- \usepackage{float}
---
```{r global_options, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.align = 'center', dev = 'cairo_pdf')
```

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(SoilR)
library(corrplot)
library(openxlsx)
library(ISRaD)
```
## Figures

# Results
## Respiration rates

```{r CO2-resp-rates}
# plotted as measured
load("/Users/jeff/arc-inc/src/ts.avg.RData")
ts.avg %>%
  filter(Experiment != "tme") %>%
  ggplot(., aes(time_d_cmtv_mean, mgCO2.C_gC_d_mean)) +
  geom_vline(xintercept = 4, color="gray") +
  geom_ribbon(aes(ymin = se_slope_l, ymax = se_slope_u, fill = Type, linetype = Treatment, alpha = Treatment), show.legend = FALSE) +
  geom_point(aes(color = Type, shape = Treatment), size = 2) +
  geom_line(aes(color = Type, linetype = Treatment)) +
  facet_grid(rows = vars(Experiment),
             labeller = labeller(Experiment = c("arc" = "2011", "rewet" = "2019"))) +
  scale_x_continuous(limits = c(0,18)) +
  scale_color_manual(name = 'Ecosystem',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  scale_fill_manual(name = 'Standard error',
                    values =c('F'='#a35513','G'='#1361a3'),
                    labels = c('Forest','Grassland')) +
  scale_alpha_manual(name = 'Treatment',
                     values = c("air-dry + storage" = .2,
                                "control" = .4,
                                "air-dry" = .2)) +
  scale_linetype_manual(name = 'Treatment',
                        values = c("air-dry + storage" = 'dashed',
                                   "control" ='solid',
                                   "air-dry" = "dashed")) +
  scale_shape_manual(name = 'Treatment',
                     values = c("air-dry + storage" = 0,
                                "control" = 16,
                                "air-dry" = 1)) +
  ylab(expression('Respiration Rate (mgCO'[2]*'-C gC'^-1*'d'^-1*')')) +
  xlab("Time (days)") +
  guides(color = guide_legend(order = 1),
         shape = guide_legend(order = 3)) +
  theme_bw() +
  theme(panel.grid = element_blank())
```

>**Fig. `r {(fig.n <- 1)}`. Respiration rates for Experiment 1 (air-dry + storage treatment, 2011) and Experiment 2 (air-dry only treatment, 2019)**

>*Caption:* Top panel shows data from samples collected in 2011 for Experiment 1 (air-dry + storage treatment), bottom panel shows data from samples collected in 2019 for Experiment 2 (air-dry only treatment). Vertical gray line at day 4 demarcates the end of the pre-incubation period and the start of the equilibrium respiration period. Points show measurements and lines show trends in mean respiration rate. Shaded ribbons represent one standard error. The final measurement points for a few samples which took >18 days to reach CO~2~ targets are excluded for display reasons; respiration rates for those samples remained flat. Note that headspace CO~2~ concentrations for Experiment 1 control samples were only measured once during the pre-incubation period (day 4) in contrast to daily measurements for all other samples. Consequently the respiration rate for those samples is the cumulative average rate over the first 4 d.

## Radiocarbon data

*Pre-incubation versus equilibrium respiration ^14^C-CO~2~*

```{r pre-inc-14c-plot}
load("/Users/jeff/arc-inc/src/all.14c.pre.inc.RData")

all.14c.pre.inc %>%
  filter(Experiment != "tme") %>%
  ggplot(., aes(d14c_corr_mean_pre, d14c_corr_mean_inc, color = Type, shape = Treatment)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_vline(xintercept = 0, color = "gray") +
  geom_hline(yintercept = 0, color = "gray") +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = d14c_corr_min_inc, 
        ymax = d14c_corr_max_inc, 
        color = Type), 
    width = .25) +
  geom_errorbarh(
    aes(xmin = d14c_corr_min_pre, 
        xmax = d14c_corr_max_pre, 
        color = Type), 
    height = .9) +
  scale_color_manual(name = 'Ecosystem',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  scale_shape_manual(name = 'Treatment',
                     values = c("control" = 16,
                                "air-dry" = 1,
                                "air-dry + storage" = 0),
                     labels = c("control" = "control-2",
                                "air-dry" = "air-dry",
                                "air-dry + storage" = "air-dry + storage")) +
  scale_x_continuous(limits = c(-60, 115)) +
  scale_y_continuous(limits = c(-60, 115)) +
  xlab(expression('Rewetting pulse '*Delta*''^14*'C-CO'[2]*' (‰)')) +
  ylab(expression('Equilibrium respiration '*Delta*''^14*'C-CO'[2]*' (‰)')) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.key.height=unit(.8, "cm"),
        aspect.ratio = 1)
```
>**Fig. `r {fig.n + 1}`. $\Delta$^14^C-CO~2~ of the rewetting pulse and the equilibrium respiration period**

>*Caption:* Points are means of laboratory duplicates and error bars are the min and max (except for Experiment 1 control samples, which were not replicated). Note that rewetting pulse $\Delta$^14^C was not measured for control-1 samples; additionally samples from three of the forest plots of the air-dry + storage samples from Experiment 1 failed to accumulate enough CO~2~ during the pre-incubation period to measure $\Delta$^14^C. The outlier point with the substantially depleted pre-incubation $\Delta$^14^C is from Experiment 2 (control).

## Treatment effect on ^14^C-CO~2~ for all samples (Experiments 1, 2, and 3)

```{r plot ctl vs trt}
load("/Users/jeff/arc-inc/src/all.14c.ctl.trt.RData")

# list of locations for recoding
loc <- list(
    "Duke" = "Duke, NC",
    "Howland" = "Howland, ME",
    "Harvard" = "Harvard, MA",
    "Hainich" = "Central Germany",
    "Schorfheide" = "Central Germany",
    "Musick" = "Sierra Nevada, CA",
    "Shaver" = "Sierra Nevada, CA",
    "Walker Branch" = "Oak Ridge, TN",
    "Tennessee Valley" = "Oak Ridge, TN")

# Plot on 1:1 line w/ diagonal lines showing ‰ offsets
# first collapse sites for better visuals
all.14c.ctl.trt %>%
  filter(Horizon == "A") %>%
  filter(Experiment != "rewet") %>%
  mutate(Location = recode(Site, !!!loc)) %>%
  ggplot(., aes(d14c_corr_mean_ctl, d14c_corr_mean_trt, color = Location, shape = Type)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_abline(slope = 1, intercept = -20, linetype = "dashed") +
  geom_abline(slope = 1, intercept = 20, linetype = "dashed") +
  geom_abline(slope = 1, intercept = -40, linetype = "dotted") +
  geom_abline(slope = 1, intercept = 40, linetype = "dotted") +
  geom_vline(xintercept = 0, color = "gray") +
  geom_hline(yintercept = 0, color = "gray") +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = d14c_corr_min_trt, 
        ymax = d14c_corr_max_trt, 
        color = Location), 
    width = .25) +
  geom_errorbarh(
    aes(xmin = d14c_corr_min_ctl, 
        xmax = d14c_corr_max_ctl, 
        color = Location), 
    height = .9) +
  scale_color_manual(values = c('Duke, NC' = '#01BB97',
                                'Sierra Nevada, CA' = '#FFC107',
                                'Central Germany' = 'black',
                                'Harvard, MA' = '#1E88E5',
                                'Oak Ridge, TN' = '#D81B60')) +
  scale_shape_manual(name = 'Land use',
                     values = c('F'= 17,'G' = 16),
                     labels = c('Forest','Grassland')) +
  xlab(expression('Control '*Delta*''^14*'C (‰)')) +
  ylab(expression('Air-dry + storage '*Delta*''^14*'C (‰)')) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.key.height=unit(.8, "cm"),
        aspect.ratio = 1)
```
>**Fig. `r {fig.n + 2}`. Overall effect of air-drying + storage on ^14^C-CO~2~**

>*Caption:* Points show data from Experiments 1 and 3 (air-dry + storage treatment). Points are the mean of laboratory replicates (for replicated samples); error bars are 2x standard error. Solid line is 1:1. For context, the dashed and dotted lines show differences of ±20‰ and ±$40‰, equivalent to the decline in $Delta/$^14^C in atmospheric CO~2~ over 4 and 8 y respectively, during the period of 2000 to 2020 (Graven et al. 2017).


```{r treatment effects by time}
load("/Users/jeff/arc-inc/src/Datm.RData")
load("/Users/jeff/arc-inc/src/all.14c.sum2.RData")

# plot over time, averaged by Type
all.14c.sum2 %>%
  filter(Period == "inc") %>%
  mutate(year.Type = paste(YearSampled, Type),
         Eco = ifelse(Type == "F", "Forest", "Grassland")) %>%
  ggplot(., aes(YearSampled, d14c_corr_mean, color = Type)) +
  geom_point(aes(shape = Treatment), size = 3, stroke = 1.5) +
  geom_path(aes(group = year.Type), 
            arrow = arrow(type = "open", length = unit(.1, "inches"), ends = "first"),
            show.legend = FALSE,
            position = position_dodge(width = 2, preserve = "single"),
            size = 1.2) +
  geom_path(data = Datm, aes(Date, NHc14, linetype = Atmosphere), color = "darkgray") +
  geom_errorbar(
    aes(ymin = d14c_corr_l, 
        ymax = d14c_corr_u, 
        color = Type), 
    width = .5,
    alpha = .3) +
  scale_color_manual(name = 'Ecosystem',
                     values = c('F'='#a35513',
                                'G'='#1361a3'),
                     labels = c('forest',
                                'grassland')) +
  scale_shape_manual(name = 'Treatment',
                     values = c("air-dry" = 1, 
                                "air-dry + storage" = 0,
                                "control" = 16),
                     labels = c("air-dry" = "air-dry", 
                                "air-dry + storage" = "air-dry + storage",
                                "control" = "control-1 (2011), control-2 (2019)")) +
  scale_x_continuous(limits = c(2010, 2020), breaks = c(2013, 2016, 2019)) +
  scale_y_continuous(limits = c(-15, 100)) +
  xlab('Year sampled') +
  ylab(expression('mean '*Delta*''^14*'C (‰)')) +
  facet_grid(cols = vars(Eco)) +
  theme_bw() +
  theme(panel.grid = element_blank())
```
>**Fig `r {fig.n + 3}`. Time series of control and treatment $\Delta$^14^C-CO~2~ (Experiments 1 and 2)**

>*Caption:* Filled circles show $\Delta$^14^C-CO~2~ observed for both control-1 and control-2 samples (2011 and 2019 points, respectively). Open symbols show $\Delta$^14^C-CO~2~ observed for treament samples: open squares = air-dry + storage treatment, Experiment 1; open circles = air-dry only treatment, Experiment 2. Arrows show the direction of change in $\Delta$^14^C-CO~2~ relative to the controls. Points are means and error bars show 2x standard error. The gray line shows $\Delta$^14^C of the atmosphere.

## Storage duration effect on ^14^C-CO~2~ (Experiment 3)

```{r plot-dur-stor}
# Plot storage duration
suppressWarnings(
  all.14c.ctl.trt %>%
  filter(Horizon == "A") %>%
  filter(Experiment != "rewet") %>%
  select(Type, tme.f, Site, trt.ctl.d14, tme) %>%
  group_by(Type, tme.f, Site) %>%
  add_tally() %>%
  summarise_all(list(mean = mean, sd = sd), na.rm = TRUE) %>%
  mutate(dif_se = trt.ctl.d14_sd/n_mean,
         dif_u = trt.ctl.d14_mean + dif_se*2,
         dif_l = trt.ctl.d14_mean - dif_se*2) %>%
  mutate(Location = factor(recode(Site, !!!loc), exclude = TRUE)) %>%
  ggplot(., aes(tme_mean, trt.ctl.d14_mean, color = Location, shape = Type)) +
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = -20, linetype = "dashed") +
  geom_hline(yintercept = 20, linetype = "dashed") +
  geom_hline(yintercept = 40, linetype = "dotted") +
  geom_hline(yintercept = -40, linetype = "dotted") +
  geom_point(size = 3, position = position_dodge2(width = .7, preserve = "single")) +
  geom_errorbar(
    aes(ymin = dif_l,
        ymax = dif_u, 
        color = Location),
    width = .7,
    position = position_dodge2(width = .7, preserve = "single")) +
  scale_color_manual(values = c('Duke, NC' = '#01BB97',
                                'Sierra Nevada, CA' = '#FFC107',
                                'Central Germany' = 'black',
                                'Harvard, MA' = '#1E88E5',
                                'Oak Ridge, TN' = '#D81B60')) +
  scale_shape_manual(name = 'Land use',
                     values = c('F'= 17,'G' = 16),
                     labels = c('Forest','Grassland')) +
  scale_x_continuous(breaks = seq(5,13,2)) +
  ylab(expression('Treatment - Control '*Delta*''^14*'C (‰)')) +
  xlab("Storage duration (years)") +
  theme_bw() +
  theme(panel.grid = element_blank())
  )
```

>**Fig. `r {fig.n + 4}`. Change in ^14^C-CO~2~ in relation to storage duration**

>*Caption:* Data are from both Experiment 1 (in black) and Experiment 3 (all other points), averaged by site and ecosystem type. Points are the mean, error bars are 2x standard error. For context, the dashed and dotted lines are the same is in Fig. `r {fig.n + 3}` and show a difference of 20‰ and 40‰, equivalent to the decline in $Delta/$^14^C in atmospheric CO~2~ over 4 and 8 y respectively, during the period of 2000 to 2020 (Graven et al. 2017). Position of points jittered to avoid overplotting; storage duration has been rounded down to the nearest whole year.

# Discussion

```{r conceptual-model-fig1}
ggplot(f.2p.C14.df, aes(years, d14C, color = pool)) +
  geom_path() +
  scale_color_manual(
    name = "Pool",
    values = c("atm" = 8,
               "fast" = "#D81B60", 
               "respiration" = "#FFC107", 
               "slow" = "#1E88E5"),
    labels = c("atm" = "atmosphere",
               "fast" = "fast",
               "slow" = "slow")) +
  scale_x_continuous(limits = c(1950, 2022)) +
  xlab("Year") +
  ylab(expression(''*Delta*''^14*'C (‰)')) +
  theme_bw() +
  theme(panel.grid = element_blank())
```

>**Fig `r {(fig.n + 5)}`a. Modeled trajectories of $\Delta$^14^C over time for a hypothetical 2-pool soil carbon model in relationship to observed atmospheric $\Delta$^14^C**

>*Caption:* Modeled curves derived from a hypothetical two-pool parallel model system in which inputs are partitioned between a fast cycling soil carbon pool (*k* = 1/6) and a more slowly cycling pool (*k* = 1/100) without any tranfers between the pools. Carbon stocks and pool sizes are based on density fraction data for the Hainich Forest (Schrumpf et al. 2013); decomposition rates are chosen to be realistic but are arbitrary. Atmospheric $\Delta$^14^C data up to the year 2015 are from Graven et al. (2017), while data points beyond 2015 use the extrapolation method from Sierra (2018). All atmospheric radiocarbon data is for the northern hemisphere (zone 2).

```{r conceptual-model-fig1b}
# zoom (w/ atm, to 2022)
f.2p.C14.df %>% 
  filter(years > 1987) %>%
  ggplot(., aes(years, d14C)) +
  geom_path(aes(color = pool)) +
  geom_point(data = obs.pts, aes(years, d14c, shape = Observation), color = "#FFC107", stroke = 1, size = 3) +
  scale_color_manual(
    name = "Pool",
    values = c("atm" = 8,
               "fast" = "#D81B60", 
               "respiration" = "#FFC107", 
               "slow" = "#1E88E5"),
    labels = c("atm" = "atmosphere",
               "fast" = "fast",
               "slow" = "slow")) +
  scale_shape_manual(
    name = "Observed respiration",
    values = c("control" = 16,
               "increased slow pool contribution" = 0,
               "increased fast pool contribution" = 1)) +
  scale_y_continuous(limits = c(-25, 265)) +
  xlab("Year") +
  ylab(expression(''*Delta*''^14*'C (‰)')) +
  theme_bw() +
  theme(panel.grid = element_blank())
```

>**Fig `r {(fig.n + 5)}`b. Potential shifts in $\Delta$^14^C of respired CO~2~ in response to treatment**

>*Caption:* Blue and magenta lines show modeled trajectories of $\Delta$^14^C in fast and slow cycling soil carbon pools (same model as Fig 1a), while $\Delta$^14^C of respired CO~2~ is shown in yellow. FIlled circles show hypothetical observations of $\Delta$^14^C-CO~2~ from control incubations, while the open symbols represent two possible scenarios in which an air-drying and rewetting treatment leads to shifts in $\Delta$^14^C-CO~2~. In the first scenario (open squares), an increased contribution from the slow pool shifts $\Delta$^14^C-CO~2~ toward the slow pool curve (blue line), while in the second scenario (open circles), an increased contribution from the fast pool shifts $\Delta$^14^C-CO~2~ toward the fast pool curve (magenta line). Due to the crossing of the slow and fast pool curves in 2009, an increased contribution of the slow pool to respiration following treatment leads to relative depletion of $\Delta$^14^C-CO~2~ in 1991, but relative enrichment of $\Delta$^14^C-CO~2~ in 2019, while the opposite is observed for the increased fast pool contribution scenario.

## Supplemental figures
*Respiration rates:*

Preincubation headspace CO~2~ concentations were only measured once for the control-1 samples, at the end of the four=day preincubation period. In order to compare the control-1 respiration rates more easily with the air-dry + storage treatment samples, Supplemental Figure 1 shows the air-dry + storage pre-incubation respiration rates plotted as a cumulative average for the preincubation period. However, as can be seen in Figure 1, headspace CO~2~ concentration was measured daily for the air-dry + storage samples.

```{r CO2-flux-plots-SI-1}
# Note: this code plots data with pre-incubation observations averaged
# Plot resp rates over time
ts.avg2 %>%
  filter(Experiment != "tme") %>%
  ggplot(., aes(time_d_cmtv_mean, mgCO2.C_gC_d_mean)) +
  geom_vline(xintercept = 4, color="gray") +
  geom_ribbon(aes(ymin = se_slope_l, ymax = se_slope_u, fill = Type, linetype = Treatment, alpha = Treatment)) +
  geom_point(aes(color = Type, shape = Treatment)) +
  geom_line(aes(color = Type, linetype = Treatment)) +
  facet_grid(rows = vars(Experiment),
             labeller = labeller(Experiment = c("arc" = "2011", "rewet" = "2019"))) +
  scale_x_continuous(limits = c(0,18)) +
  scale_color_manual(name = 'Ecosystem',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  scale_fill_manual(name = 'Std. Err.',
                    values =c('F'='#a35513','G'='#1361a3'),
                    labels = c('Forest','Grassland')) +
  scale_alpha_manual(name = 'Treatment',
                     values = c("air-dry + storage" = .2,
                                "control" = .4,
                                "air-dry" = .2)) +
  scale_linetype_manual(name = 'Treatment',
                        values = c("air-dry + storage" = 'dashed',
                                   "control" ='solid',
                                   "air-dry" = "dashed")) +
  scale_shape_manual(name = 'Treatment',
                     values = c("air-dry + storage" = 0,
                                "control" = 16,
                                "air-dry" = 1)) +
  ylab(expression('Respiration Rate (mgCO'[2]*'-C gC'^-1*'d'^-1*')')) +
  xlab("Time (days)") +
  theme_bw() +
  theme(panel.grid = element_blank())
```

>**Supplemental Fig `r {(sfig.n <- 1)}`. Respiration rates for Experiment 1 and Experiment 2 shown with all pre-incubation data calculated as cumulative averages**

>*Caption:* CO~2~ concentrations for Experiment 1 control samples were only measured once during the pre-incubation period, in contrast to daily measurements for all other samples. Pre-incubation respiration rates are shown here calculated as cumulative averages for the purpose of fair comparison across all treatments.


*Experiment 3 (storage duration)*

Grassland soils from the storage duration treatment group responded rapidly to rewetting and reached target CO~2~ levels after just `r {round(ts.avg2[which(ts.avg2$mgCO2.C_gC_d_mean == max(ts.avg2[ts.avg2$Experiment == "tme" & ts.avg2$Treatment == "storage duration" & ts.avg2$Type == "G", "mgCO2.C_gC_d_mean"])), "time_d_cmtv_mean"]*24,0)}` h of incubation. Only a single observation was made for the grassland treament samples due to the rapid respiration rates, which peaked at `r {round(max(ts.avg2[ts.avg2$Experiment == "tme" & ts.avg2$Treatment == "storage duration" & ts.avg2$Type == "G", "mgCO2.C_gC_d_mean"]),1)}` mg CO~2~ g soil C^-1^ d^-1^. The mean peak respiration rate for forest treatment samples was lower and lagged in comparison to the grassland soils, reaching `r {round(max(ts.avg2[ts.avg2$Experiment == "tme" & ts.avg2$Treatment == "storage duration" & ts.avg2$Type == "F", "mgCO2.C_gC_d_mean"]),1)}` mg CO~2~ g soil C^-1^ d^-1^ after `r {round(ts.avg2[which(ts.avg2$mgCO2.C_gC_d_mean == max(ts.avg2[ts.avg2$Experiment == "tme" & ts.avg2$Treatment == "storage duration" & ts.avg2$Type == "F", "mgCO2.C_gC_d_mean"])), "time_d_cmtv_mean"]*24,0)}` h.

Control-3 respiration rates peaked after the pre-incubation period (`r {round(ts.avg2[which(ts.avg2$mgCO2.C_gC_d_mean == max(ts.avg2[ts.avg2$Experiment == "tme" & ts.avg2$Treatment == "control" & ts.avg2$Type == "G", "mgCO2.C_gC_d_mean"])), "time_d_cmtv_mean"]*24,0)}` h), at `r {round(max(ts.avg2[ts.avg2$Experiment == "tme" & ts.avg2$Treatment == "control" & ts.avg2$Type == "G", "mgCO2.C_gC_d_mean"]),1)}` mg CO~2~ g soil C^-1^ d^-1^. Forest control samples were pre-incubated under various conditions, but respiration rates were only measured during the pre-incubation period for two samples from the Sierra Nevada mountains (USA). In general, respiration rates for forest control samples in Experiment 3 were much lower than in the treatment incubations, peaking at `r {round(max(ts.avg2[ts.avg2$Experiment == "tme" & ts.avg2$Treatment == "control" & ts.avg2$Type == "F", "mgCO2.C_gC_d_mean"]),1)}` mg CO~2~ g soil C^-1^ d^-1^ after `r {round(ts.avg2[which(ts.avg2$mgCO2.C_gC_d_mean == max(ts.avg2[ts.avg2$Experiment == "tme" & ts.avg2$Treatment == "control" & ts.avg2$Type == "F", "mgCO2.C_gC_d_mean"])), "time_d_cmtv_mean"]*24,0)}` h.

```{r CO2-flux-plots-SI-2}
# respiration rates for experiment 3
ts.avg %>%
  filter(Experiment == "tme") %>%
  ggplot(., aes(time_d_cmtv_mean, mgCO2.C_gC_d_mean)) +
  geom_ribbon(aes(ymin = se_slope_l, ymax = se_slope_u, fill = Type, linetype = Treatment, alpha = Treatment)) +
  geom_point(aes(color = Type, shape = Treatment)) +
  geom_line(aes(color = Type, linetype = Treatment)) +
  facet_grid(rows = vars(Experiment)) +
  scale_x_continuous(limits = c(0,18)) +
  scale_color_manual(name = 'Ecosystem',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  scale_fill_manual(name = 'Std. Err.',
                    values =c('F'='#a35513','G'='#1361a3'),
                    labels = c('Forest','Grassland')) +
  scale_alpha_manual(name = 'Treatment',
                     values = c("air-dry + storage" = .2,
                                "control" = .4,
                                "air-dry" = .2,
                                "storage duration" = .2)) +
  scale_linetype_manual(name = 'Treatment',
                        values = c("air-dry + storage" = 'dashed',
                                   "control" ='solid',
                                   "air-dry" = "dashed",
                                   "storage duration" = "dashed")) +
  scale_shape_manual(name = 'Treatment',
                     values = c("air-dry + storage" = 1,
                                "control" = 16,
                                "air-dry" = 1,
                                "storage duration" = 1)) +
  ylab(expression('Respiration Rate (mgCO'[2]*'-C gC'^-1*'d'^-1*')')) +
  xlab("Time (days)") +
  theme_bw() +
  theme(panel.grid = element_blank())
```

>**Supplemental Fig `r {(sfig.n + 1)}`. Respiration rates for Experiment 3**

>*Caption:* Experiment 3 storage duration treatment samples were only incubated for a single enclosure period, as the results of Experiment 1 and Experiment 2 showed no significant difference in $Delta/$^14^C-CO~2~ between the rewetting pulse CO~2~ released during the pre-incubation period and the CO~2~ respired during the equilibrium respiration period. The grassland storage duration treatment samples (blue dashed line) respired an equivalent amount of CO~2~ in just 3 d as the corresponding control-3 samples respired during the pre-incubation period and the equilibrium respiration period combined. Consequently those incubations were stopped after the first CO~2~ measurement point. Control-3 samples did undergo pre-incubation, but as the CO~2~ release was not measured nor was $Delta/$^14^C-CO~2~ for the majority of the samples, all data were averaged by day of measurement. 


# Change in soil moisture content with moisture adjustment

```{r delta-moisture-calc}
# compare the change in moisture with control samples versus treatment samples
whcFresh <- merge(archive.ctl.ts.wide[archive.ctl.ts.wide$Experiment == "arc", c("ID", "whcFresh")],
                  jarinfo.rewet[jarinfo.rewet$Treatment == "control", c("ID", "whcFresh")],
                  by = "ID",
                  all.x = TRUE,
                  suffixes = c("_2011", "_2019"))
whcFresh$deltaWHC_2011 <- .6-whcFresh$whcFresh_2011
whcFresh$deltaWHC_2019 <- .6-whcFresh$whcFresh_2019
whcFresh$Type <- ifelse(grepl("G", whcFresh$ID), "G", "F")

# sumamrize
whcFresh_sum <- whcFresh %>%
  select(Type, deltaWHC_2011, deltaWHC_2019, whcFresh_2011, whcFresh_2019) %>%
  group_by(Type) %>%
  add_tally() %>%
  summarize_all(list(mean = mean, sd = sd), na.rm = TRUE) %>%
  mutate(delta_se_11 = deltaWHC_2011_sd/n_mean,
         delta_se_19 = deltaWHC_2019_sd/n_mean,
         fresh_se_11 = whcFresh_2011_sd/n_mean,
         fresh_se_19 = whcFresh_2019_sd/n_mean)

# switch whcFresh to long
whcFresh.2011 <- whcFresh[ , c("ID", "whcFresh_2011", "deltaWHC_2011")]
names(whcFresh.2011) <- c("ID", "whcFresh", "deltaWHC")
whcFresh.2011$Experiment <- "arc"
whcFresh.2019 <- whcFresh[ , c("ID", "whcFresh_2019", "deltaWHC_2019")]
names(whcFresh.2019) <- c("ID", "whcFresh", "deltaWHC")
whcFresh.2019$Experiment <- "rewet"
whcFresh.2019 <- whcFresh.2019[which(complete.cases(whcFresh.2019)), ]
whcFresh.long <- rbind(whcFresh.2011, whcFresh.2019)

# stats
whc.ctl.trt <- merge(arc.rewet.14c.ctl.trt, whcFresh.long, by = c("ID", "Experiment"))
ggplot(whc.ctl.trt, aes(whcFresh, d14c_dif, color = Type, shape = Experiment)) +
  geom_hline(yintercept = 0) +
  geom_point(size = 3) +
  scale_color_manual(name = 'Ecosystem',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  theme_bw() +
  theme(panel.grid = element_blank())

# summary(lm(d14c_dif ~ whcFresh * Type, whc.ctl.trt))

# # calculate atm 14C change
# slopes <- data.frame(Datm[Datm$Date > 2000 & Datm$Date < 2016, ],
#                      slope = NA)
# for(i in seq_along(Datm[Datm$Date > 2000 & Datm$Date < 2016, "Date"])) {
#   if(i != nrow(slopes)) {
#     slopes$slope[i] <- (Datm[Datm$Date > 2000, "NHc14"][i+1]-Datm[Datm$Date > 2000, "NHc14"][i])/(Datm[Datm$Date > 2000, "Date"][i+1]-Datm[Datm$Date > 2000, "Date"][i])
#   }
# }
# mean(slopes$slope, na.rm = T)
```

>**Supplemental Fig `r {(sfig.n + 4)}`. Change in $\Delta$^14^C-CO~2~ (control - treatment) relative to field moisture**

>*Caption:* Data are from Experiment 1 ("arc") and Experiment 2 ("rewet"). All samples were moisture-adjusted prior to incubation, but control samples were adjusted from field moisture, "whcFresh" (percent of WHC), whereas treatment samples were moisture adjusted after air-drying, i.e. at approximately 0% of WHC.

# Rewetting pulse 13C vs. equilibrium respiration 13C

```{r pre-inc-13c-plot, include = FALSE}
iso.12.avg <- iso_df %>%
  select(ID, Experiment, Period, Treatment, d13c) %>%
  group_by(ID, Experiment, Period, Treatment) %>%
  add_count() %>%
  summarize_all(list(mean = mean, sd = sd), na.rm = TRUE) %>%
  mutate(d13c_se = d13c_sd/n_mean,
          d13c_u = d13c_mean + d13c_se*2,
          d13c_l = d13c_mean - d13c_se*2)

iso.12.avg$Type <- ifelse(grepl("G", iso.12.avg$ID), "G", "F")

# stats
# air-dry + storage
# forest
t.test(iso_df[iso_df$Period == "inc" & iso_df$Experiment == "arc" & iso_df$Treatment == "control" & iso_df$Type == "F", "d13c"], 
       iso_df[iso_df$Period == "inc" & iso_df$Experiment == "arc" & iso_df$Treatment == "air-dry + storage" & iso_df$Type == "F", "d13c"], tails = 2)

# grassland
t.test(iso_df[iso_df$Period == "inc" & iso_df$Experiment == "arc" & iso_df$Treatment == "control" & iso_df$Type == "G", "d13c"], 
       iso_df[iso_df$Period == "inc" & iso_df$Experiment == "arc" & iso_df$Treatment == "air-dry + storage" & iso_df$Type == "G", "d13c"], tails = 2)

# air-dry only
# forest
t.test(iso_df[iso_df$Period == "inc" & iso_df$Experiment == "rewet" & iso_df$Treatment == "control" & iso_df$Type == "F", "d13c"], 
       iso_df[iso_df$Period == "inc" & iso_df$Experiment == "rewet" & iso_df$Treatment == "air-dry" & iso_df$Type == "F", "d13c"], tails = 2)

# grassland
t.test(iso_df[iso_df$Period == "inc" & iso_df$Experiment == "rewet" & iso_df$Treatment == "control" & iso_df$Type == "G", "d13c"], 
       iso_df[iso_df$Period == "inc" & iso_df$Experiment == "rewet" & iso_df$Treatment == "air-dry" & iso_df$Type == "G", "d13c"], tails = 2)



iso.12.pre.inc <- merge(iso.12.avg[iso.12.avg$Period == "pre", c("ID", "Experiment", "Type", "Treatment", "d13c_mean", "d13c_u", "d13c_l")],
                        iso.12.avg[iso.12.avg$Period == "inc", c("ID", "Experiment", "Type", "Treatment", "d13c_mean", "d13c_u", "d13c_l")],
                         by = c("ID", "Experiment", "Type", "Treatment"),
                         suffixes = c("_pre", "_inc"))
iso.12.pre.inc$treat.bi <- ifelse(iso.12.pre.inc$Treatment == "control", "control", "treatment")

# range(c(iso.12.pre.inc$d13c_l_pre, iso.12.pre.inc$d13c_l_inc, iso.12.pre.inc$d13c_u_inc, iso.12.pre.inc$d13c_l_inc), na.rm = TRUE)
iso.12.pre.inc %>%
  ggplot(., aes(d13c_mean_pre, d13c_mean_inc, color = Type, shape = Treatment)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = d13c_l_inc, 
        ymax = d13c_u_inc, 
        color = Type), 
    width = .25) +
  geom_errorbarh(
    aes(xmin = d13c_l_pre, 
        xmax = d13c_u_pre,
        color = Type), 
    height = .25) +
  scale_color_manual(name = 'Ecosystem',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  scale_shape_manual(name = 'Treatment',
                     values = c("control" = 16,
                                "air-dry" = 1,
                                "air-dry + storage" = 0),
                     labels = c("control" = "control-1 (2011), control-2 (2019)",
                                "air-dry" = "air-dry",
                                "air-dry + storage" = "air-dry + storage")) +
  scale_x_continuous(limits = c(-29.1, -22.25)) +
  scale_y_continuous(limits = c(-29.1, -22.25)) +
  xlab(expression('Pre-incubation '*delta*''^13*'C (‰)')) +
  ylab(expression('Equilibrium respiration '*delta*''^13*'C (‰)')) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.key.height=unit(.8, "cm"),
        aspect.ratio = 1)
```

# Time series of 13C, Experiments 1 & 2
```{r ctl-trt-difs-13c}
# average control and treatment data
# note that upper/lower limits = 2xSE

# Plot side-by-side w/ 2xSE
iso.12.avg2 %>%
  filter(Period == "inc") %>%
  mutate(YearSampled = ifelse(Experiment == "arc", "2011", "2019")) %>%
  mutate(Eco = ifelse(Type == "F", "Forest", "Grassland")) %>%
  ggplot(., aes(YearSampled, d13c_mean, color = Type)) +
  geom_point(aes(shape = Treatment), size = 3) +
  geom_errorbar(
    aes(ymin = d13c_u, 
        ymax = d13c_l, 
        color = Type), 
    width = .1) +
  scale_color_manual(name = 'Ecosystem',
                     values =c('F'='#a35513','G'='#1361a3')) +
  scale_shape_manual(name = 'Treatment',
                     values = c("air-dry" = 1, 
                                "air-dry + storage" = 0,
                                "control" = 16)) +
  ylab(expression('mean '*delta*''^13*'C (‰)')) +
  facet_grid(cols = vars(Eco)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.title.x = element_blank())

# iso.12.avg2 %>%
#   mutate(YearSampled = ifelse(Experiment == "arc", "2011", "2019"),
#          treat.period = paste(Treatment, Period, sep = "."),
#          Eco = ifelse(Type == "F", "Forest", "Grassland")) %>%
#   ggplot(., aes(YearSampled, d13c_mean, color = Type)) +
#   geom_point(aes(shape = treat.period), size = 3) +
#   geom_errorbar(
#     aes(ymin = d13c_u,
#         ymax = d13c_l,
#         color = Type),
#     width = .1) +
#   scale_color_manual(name = 'Ecosystem',
#                      values = c('F'='#a35513','G'='#1361a3')) +
#   scale_shape_manual(name = 'Treatment',
#                      values = c("air-dry.pre" = 2,
#                                 "air-dry.inc" = 1,
#                                 "air-dry + storage.inc" = 0,
#                                 "air-dry + storage.pre" = 6,
#                                 "control.inc" = 16,
#                                 "control.pre" = 17)) +
#   ylab(expression('mean '*delta*''^13*'C (‰)')) +
#   facet_grid(cols = vars(Eco)) +
#   theme_bw() +
#   theme(panel.grid = element_blank(),
#         axis.title.x = element_blank())
```
>**Supplemental Fig `r {sfig.n + 2}`. Time series of control and treatment $\delta$^13^C-CO~2~ (Experiments 1 and 2)**

>*Caption:* Filled circles show $\delta$^13^C-CO~2~ observed for control samples, while open symbols show $\delta$^13^C-CO~2~ observed for treament samples (open squares = air-dry + storage treatment, Experiment 1, 2011; open circles = air-dry only treatment, Experiment 2, 2019). Points are means and error bars show 2x standard error. 

*Effect of cumulative respired carbon on ^14^C-CO~2~*

We looked at the possible effect of the difference in the amount of carbon respired (mg CO~2~-C g soil C^-1^) on the differences between control and treatment ^14^C-CO~2~ using a linear regression model, but it was not significant overall. When data from Experiment 1 and Experiment 2 were considered separately, we observed a slight positive trend between the difference in respired carbon and the difference in ^14^C-CO~2~ within Experiment 2, but it was only marginally significant (p = `r {round(summary(lm(d14c_dif ~ c_dif * Type, arc.rewet.14c.ctl.trt[arc.rewet.14c.ctl.trt$Experiment == "rewet", ]))$coef[2,4],3)}`). 

```{r plot-14c-by-c-resp}
# 10. Plot $\Delta$^14^C against proportion of soil C respired by experiment, land cover, and sampling period.
# collapse treatment variable into binary control/treatment
all.14c.sum.df$treat.bi <- ifelse(all.14c.sum.df$Treatment == "control", "control", "treatment")

# HEW22 control pre point excluded to make figure more readable 
all.14c.sum.df %>%
  filter(Experiment != "tme") %>%
  mutate(ID.treat = paste(ID, Treatment)) %>%
  mutate(ID.treat.Period = paste(ID.treat, Period)) %>%
  filter(ID.treat.Period != "HEW22 control pre") %>%
  mutate(treat.Period = paste(treat.bi, Period),
         Eco = ifelse(Type == "F", "Forest", "Grassland")) %>%
  ggplot(., aes(mgCO2.C_gC_mean, d14c_corr_mean, color = Type, shape = treat.Period)) +
  geom_point(size = 3) +
  geom_path(aes(group = ID.treat)) +
  geom_errorbar(
    aes(ymin = d14c_corr_min, 
        ymax = d14c_corr_max, 
        color = Type), 
    width = .25) +
  geom_errorbarh(
    aes(xmin = mgCO2.C_gC_min, 
        xmax = mgCO2.C_gC_max, 
        color = Type), 
    height = .9) +
  scale_color_manual(name = 'Ecosystem',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  scale_shape_manual(name = 'Sampling period (treatment)',
                     values =c(16, 17, 1, 2),
                     labels = c("control pre" = 'pre-incubation\n(control)', 
                                "control inc" = 'equilibrium respiration\n(control)',
                                "treatment pre" = 'pre-incubation\n(air-dry/air-dry + storage)',
                                "treatment inc" = 'equilibrium respiration\n(air-dry/air-dry + storage)')) +
  facet_grid(cols = vars(Eco), rows = vars(YearSampled)) +
  xlab(expression('Respired C (mg CO'[2]*'-C'*' g soil C'^-1*')')) +
  ylab(expression('mean '*Delta*''^14*'C (‰)')) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.key.height=unit(.8, "cm"))
```

>**Supplementary Fig. `r {sfig.n + 3}`. Change in ^14^C-CO~2~ in relation to cumulative soil carbon respired**

>*Caption:* Note that pre-incubation $\Delta$^14^C was not measured for the control-1 samples in 2011. Limits exclude outlier point (HEW22 control-2, pre-incubation) for improved legibility. Points are means, error bars show min and max of duplicate samples.
