---
title: "Archived Soil Incubations Project"
author: "J. Beem-Miller"
date: "30 Apr 2020"
output:
  pdf_document:
    latex_engine: xelatex
  html_notebook:
    toc: yes
    toc_depth: 2
header_includes:
  - \usepackage[utf8]{inputenc}
  - \usepackage{float}
---
```{r global_options, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.align = 'center', dev = 'cairo_pdf')
```

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(SoilR)
library(corrplot)
library(openxlsx)
library(ISRaD)
```
# Notes
* This workbook is intended to load and prepare the key data for analysis for the archive incubation project.
* in general, this is an updated version of script "./src/arc_inc_master.R"
* all code chunk options are set to "echo = FALSE"; see raw .Rmd file for data wrangling code.

# Introduction
The laboratory soil incubation is a commonly used technique for understanding soil carbon dynamics. Soil carbon is a heterogeneous mixture of organic matter, some of which persists in the soil for months or years, while some persists for centuries or millenia. The persistence of soil carbon can be understood through the concept of different "pools" of carbon that are defined by the mechanism by which they persist in the soil and are characterized by distinct age distributions. 

Natural abundance radiocarbon provides information about carbon dynamics on the scale of centuries to millenia, while insight into decadal scale dynamics can be gained from tracing the pulse of radiocarbon introduced into the biosphere from nuclear weapons testing ("bomb-C") in the mid-20^th^ century. The bomb-C pulse peaked in the atmosphere in the 1950s (Fig. 1), but due to differential rates of abiotic incorporation and biological processing, the peak is offset forward in time in soils. The relative enrichment in bomb-C in different pools of soil carbon is a useful tool for inferring the relative rate at which carbon enters and leaves the pool, and for a homogenous pool it is functionally equivalent to the intrinsic decomposition rate.

Extracting and measuring the radiocarbon content of specific soil carbon pools is hampered by spatial and temporal heterogeneity of the mechanisms that lead to soil carbon persistence, such as physical occlusion in aggregates, association with minerals, or thermodynamics. Defining soil carbon pools empirically with techniques such as density, size, or resistance to chemical attack can be useful, but these methods also introduce artifacts and likely result in mixtures of pools with different age distributions. In contrast, although they also introduce artifacts due to disturbance and potential alteration of the microbial community, laboratory soil incubations make use of the same fractionation agent as is found *in situ*: the microbial community. Measuring the radiocarbon signal of CO^2^ ($\Delta$^14^C-CO~2~) released in laboratory incubations of bulk soils is a powerful tool for understanding the relative processing rate of carbon in soil (or transit time) as it provides an integrated measure of the weighted contribution to the release flux from pools of soil carbon with distinct processing rates. 

Interpreting $\Delta$^14^C-CO~2~ from laboratory soil incubations requires the use of a model. However, parameterizing these models is challenging, both due to the uncertainty of the persistence mechanisms themselves as well as a lack of observational constraints. Radiocarbon observations at a single point in time are very useful, but due to the curvature of the bomb-C peak there are two points in time with the same atmospheric radiocarbon value, leading to multiple model solutions. Adding additional observations of $\Delta$^14^C-CO~2~ at multiple points in time can greatly reduce model uncertainty.

Soil archives have already proven to be a valuable source of data for constraining soil carbon models by providing time series of the change in soil carbon ^14^C content. The promise of improving models further by obtaining ^14^C-CO~2~ measurements from archived soils is tantalizing, but first the possible effects of air-drying and rewetting, as well as the effect of the duration of storage, must be quantified.

We designed an experiment to assess these effects, preliminary results of which are presented in the following report.

# Conceptual understanding
If the null hypothesis is disproved, i.e. air-drying and rewetting (air-dry) or air-drying, rewetting, and storage (air-dry + storage) have a significant effect on $\Delta$^14^C-CO~2~ relative to control samples, there are two possible outcomes: enrichment of the respiration flux or depletion. 

In a simple two-pool soil carbon model, the treatment effect can be conceptualized as a relative change in the contribution of "slower" or "faster" cycling carbon to the respiration flux observed in control samples.

Trajectories of the change in $\Delta$^14^C over time for "slow" and "fast" soil carbon pools are shown below (Fig. 1) in relationship to the $\Delta$^14^C of the atmosphere and respired CO~2~ for a theoretical model system.
```{r conceptual-model-calcs, include = FALSE}
# initial parameter set
Datm <- rbind(graven, future14C)
Datm <- Datm[Datm$Date > 1900, c("Date", "NHc14")]
yrs <- Datm$Date

# helper fxs
lambda <- (1/8267)

k <- function(Fm) {
  (Fm*lambda)/(1-Fm)
}
fm <- function(k){
  k/(k+lambda)
}
fm_14c <- function(fm, date) {
  (fm*exp(lambda*(1950 - date)) - 1)*1000
}

# forests
f.c <- 38.96104 # forest C pool, total (mean Hainich forest C  = 24 g kg^-1); nb @ ss = sum(-1*solve(A)%*%c(f.in*f.gam, f.in*(1-f.gam)))
f.in <- 1 # inputs (12.5=mean equilibrium C flux, annual extrapolation, gC yr-1, 2011 incubation data); try much less...
f.frc <- .1 # C-stock partitioning coefficient (from Schrumpf 2013, Hainich 0-5, free light)
f.Cfast <- f.frc*f.c
f.Cslow <- f.c-f.Cfast
f.kfast <- 1/6 # unknown
f.kslow <- 1/100 # unknown
f.fast.fm <- fm(f.kfast)
f.slow.fm <- fm(f.kslow)
f.F0_Delta14C <- fm_14c(c(f.fast.fm, f.slow.fm), date = c(1900, 1900))
f.gam <- (f.c*f.Cfast*f.kfast)/(f.c*f.Cfast*f.kfast+f.c*f.Cslow*f.kslow) # input partitioning coefficient (proportion to fast pool, function of ks)
  
# grasslands
g.c <- 42 # C pool, total (mean Hainich grassland C g kg^-1)
g.in <- 63.4 # inputs (mean equilibrium C flux, annual extrapolation, 2011 incubation data)
g.frc <- .3 # C-stock partitioning coefficient (from Schrumpf 2013, average of Lacqueville, Easter Bush, Bugac, 0-5, free light)
g.Cfast <- g.frc*g.c
g.Cslow <- g.c-g.Cfast
g.kfast <- 1/10 # unknown
g.kslow <- 1/100 # unknown
g.fast.fm <- fm(g.kfast)
g.slow.fm <- fm(g.kslow)
g.F0_Delta14C <- fm_14c(c(g.fast.fm, g.slow.fm), date = c(1900, 1900))
g.gam <- (g.c*g.Cfast*g.kfast)/(g.c*g.Cfast*g.kfast+g.c*g.Cslow*g.kslow) # input partitioning coefficient (proportion to fast pool, function of ks)

# 2pool parallel model
f.2p <- TwopParallelModel14(t = yrs,
                            ks = c(f.kfast, f.kslow),
                            C0 = c(f.Cfast, f.Cslow),
                            F0_Delta14C = f.F0_Delta14C,
                            In = f.in,
                            gam = f.gam,
                            inputFc = Datm)

f.2p.C14 <- getF14(f.2p)
f.2p.HR <- getF14R(f.2p)
f.2p.Ctot <- getC(f.2p)

# solve for steady-state C stocks
A <- -1*diag(c(f.kfast, f.kslow))
sum(-1*solve(A)%*%c(f.in*f.gam, f.in*(1-f.gam)))

f.2p.C14.df <- data.frame(
  years = rep(Datm$Date, 4),
  d14C = c(f.2p.C14[,1], f.2p.C14[,2], f.2p.HR, Datm$NHc14),
  pool = rep(c("fast", "slow", "respiration", "atm"), each = nrow(f.2p.C14))
  )

# # Plot C stocks to confirm flat
# plot(yrs, xlim = c(1900,2022), ylim = c(0,60))
# lines(yrs, f.2p.Ctot[, 1], col = 2)
# lines(yrs, f.2p.Ctot[, 2], col = 4)

# calculate resp flux points at 1991 and 2019 and put in data frame
obs.pts <- data.frame(pool = rep("respiration", 6),
                      years = rep(c(1991,2019),3),
                      d14c = c(mean(f.2p.C14.df[f.2p.C14.df$years > 1990 & f.2p.C14.df$years < 1992 & f.2p.C14.df$pool == "respiration", "d14C"]),
                               mean(f.2p.C14.df[f.2p.C14.df$years > 2018 & f.2p.C14.df$years < 2020 & f.2p.C14.df$pool == "respiration", "d14C"]),
                               150, 60,
                               190, 30),
                      Observation = rep(c("control",
                                          "increased slow contribution",
                                          "increased fast contribution"),
                                        each = 2))
obs.pts.r <- obs.pts[obs.pts$Observation == "control", ]
```

```{r conceptual-model-fig1a}
ggplot(f.2p.C14.df, aes(years, d14C, color = pool)) +
  geom_path() +
  scale_color_manual(
    name = "Model pool",
    values = c("atm" = 8,
               "fast" = "#D81B60", 
               "respiration" = "#FFC107", 
               "slow" = "#1E88E5"),
    labels = c("atm" = "atmosphere",
               "fast" = "fast",
               "slow" = "slow")) +
  scale_x_continuous(limits = c(1950, 2022)) +
  xlab("Year") +
  ylab(expression(''*Delta*''^14*'C (‰)')) +
  theme_bw() +
  theme(panel.grid = element_blank())
```


The simplest two-pool model is a parallel system: one in which carbon entering the soil is partitioned between the two pools without any transfers from one pool to the other. However, even in this simple system, the same treatment effect can result in either enrichment or depletion depending on the year the sample was collected. The direction of change willl be dependent on system dynamics, specifically the intrinsic decomposition rates of the slow and fast pools, while in more complicated sytems the amount of mixing between the pools will play a role as well. Potential treatment effects for the same parallel two-pool model system shown in Fig. 1 are shown below at two points in time (Fig. 2).
```{r conceptual-model-fig1b}
# zoom (w/ atm, to 2022)
f.2p.C14.df %>% 
  filter(years > 1987) %>%
  filter(pool != "atm") %>%
  ggplot(., aes(years, d14C)) +
  geom_path(aes(color = pool)) +
  geom_point(data = obs.pts, aes(years, d14c, shape = Observation), color = "#FFC107", stroke = 1, size = 3) +
  scale_color_manual(
    name = "Model pool",
    values = c("atm" = 8,
               "fast" = "#D81B60", 
               "respiration" = "#FFC107", 
               "slow" = "#1E88E5"),
    labels = c("atm" = "atmosphere",
               "fast" = "fast",
               "slow" = "slow")) +
  scale_shape_manual(
    name = "Observed respiration",
    values = c("control" = 16,
               "increased slow contribution" = 0,
               "increased fast contribution" = 1)) +
  scale_y_continuous(limits = c(0, 265)) +
  xlab("Year") +
  ylab(expression(''*Delta*''^14*'C (‰)')) +
  theme_bw() +
  theme(panel.grid = element_blank())
```

In the above plot (Fig. 2), increased contribution from the slow pool leads to depletion in $\Delta$^14^C of respiration relative to the control observation in 1991 but relative enrichment in 2019.


# CO~2~ fluxes and soil data 
1. Load flux data from air-dry + storage (experiment 1) and storage duration (experiment 3) control samples, and convert from "wide" to "long" format so as to match other data. 
```{r read wide ts data}
# Read air-dry + storage respiration control data
archive.ctl.ts.wide <- read.csv("../data/derived/arc-storage-t1-flux/arc-inc_xplr_arc-storage-t1-flux_2020-04-29.csv")

# Convert time data from wide to long
archive.ctl.ts.d <- select(archive.ctl.ts.wide, starts_with("inc_time_d")) %>%
  pivot_longer(everything(),
               names_to = "timepoint_cmtv",
               names_prefix = "inc_time_d_",
               values_to = "time_d")
# sort by timepoint_cmtv
archive.ctl.ts.d <- data.frame(archive.ctl.ts.d[order(archive.ctl.ts.d$timepoint_cmtv),])

# Convert mgCO2 data from wide to long
archive.ctl.ts.co2 <- select(archive.ctl.ts.wide, starts_with("inc_mgCO2_jar_")) %>%
  pivot_longer(everything(),
               names_to = "timepoint_cmtv",
               names_prefix = "inc_mgCO2_jar_",
               values_to = "mgCO2_jar")
# sort by timepoint_cmtv
archive.ctl.ts.co2 <- archive.ctl.ts.co2[order(archive.ctl.ts.co2$timepoint_cmtv),]

# assemble in long dataframe 
archive.ctl.ts <- data.frame(ID = rep(unique(archive.ctl.ts.wide$ID), 4),
                             Period = "inc",
                             timepoint_cmtv = rep(seq(from = 2, 
                                                      to = 5), 
                                                  each = length(archive.ctl.ts.wide$ID)),
                             time_d = archive.ctl.ts.d$time_d,
                             mgCO2_jar = archive.ctl.ts.co2$mgCO2_jar)

# add preincubation data
archive.ctl.ts.pre <- archive.ctl.ts.wide[1:length(unique(archive.ctl.ts.wide$ID)),
                                          c("ID", "preinc_time_d", "preinc_mgCO2_jar")]
archive.ctl.ts.pre <- rename(archive.ctl.ts.pre, time_d = preinc_time_d, mgCO2_jar = preinc_mgCO2_jar)
archive.ctl.ts.pre$timepoint_cmtv <- 1
archive.ctl.ts.pre$Period <- "pre"
archive.ctl.ts.pre$time_d_cmtv <- archive.ctl.ts.pre$time_d
  
# Calculate time_d_cmtv
archive.ctl.ts <- bind_rows(
  lapply(
    split(archive.ctl.ts, archive.ctl.ts$timepoint_cmtv),
    function(x) {
      x$time_d_cmtv <- x$time_d + 
        archive.ctl.ts.pre[match(x$ID, archive.ctl.ts.pre$ID), "time_d"]
      return(x)
    }
    )
  )

# rbind preinc and inc data
archive.ctl.ts <- rbind(archive.ctl.ts.pre, archive.ctl.ts)

# add SampleName column and restore Experiment column
archive.ctl.ts$SampleName <- archive.ctl.ts$ID
archive.ctl.ts$Experiment <- archive.ctl.ts.wide[match(archive.ctl.ts$ID, archive.ctl.ts.wide$ID), "Experiment"]

# remove intermediate files
rm(archive.ctl.ts.d, archive.ctl.ts.co2, archive.ctl.ts.pre)
```

2. Load flux data from air-dry + storage samples and from air-dry experiment (ctl & treatment), C & N data for all the Exploratories samples (measured in 2011), and soil mass and moisture data for all experiments.
```{r read long ts data}
# Read Xplr C and N data
xplr.cn <- read.csv("../data/external/exploratories_ischoening_soilcn_2011/170407_CN_concentrations_Exploratories_EP_Plots_2011.csv")

# Read Xplr 14C data for control samples ("tme" and "arc" Experiments)
xplr.ctl.iso <- read.xlsx("../data/external/exploratories_ischoening_14c-co2_2011/170228_14C_Respiration.xlsx")

# Read soil mass and moisture info for air-dry + storage treatment samples
jarinfo.arc <- read.csv("../data/derived/arc-storage-t2-jarInfo/jar_information.csv")

# Read soil mass and moisture info for air-dry data
jarinfo.rewet <- read.csv("../data/derived/air-dry-2019-whc/air-dry-2019-whc_2020-04-29.csv")

# Read soil mass and moisture info for storage duration treatment samples
# remove unit row and convert to appropriate types
jarinfo.tme <- type.convert(read.xlsx("../data/raw/lab_jena_results-co2_2018-07-30/ArcInc_IncData_2018.xlsx", sheet = 1)[-1, ])

# Read Duke iso and flux data
tme.ctl.duk <- read.csv("../data/derived/arc-tme-duke/arc-tme-duke_2020-04-29.csv")

# Read Sierra Nevada (Jun's) soil data
tme.ctl.soil.sra <- read.csv("../data/derived/arc-tme-sierra/arc-tme-sierra-soil_2020-04-29.csv")

# Read Oak Ridge C data
tme.ctl.ornl <- read.csv("../data/derived/arc-tme-ornl/arc-tme-ornl_2020-04-30.csv")

# Read air-dry + storage respiration data
archive.arc.ts <- read.csv("../data/derived/arc-storage-t2-flux/archive-flux-t2_2020-04-24.csv")

# Read control respiration data from 2019 air-dry experiment
rewet.ctl.ts <- read.csv("../data/derived/fresh-2019-flux/fresh-2019-flux_2020-04-24.csv")

# Read treatment respiration data from 2019 air-dry experiment
rewet.dry.ts <- read.csv("../data/derived/dried-2019-flux/dried-2019-flux_2020-04-24.csv")

# Read Sierra Nevada (Jun's) flux data
tme.ctl.flx.sra <- read.csv("../data/derived/arc-tme-sierra/arc-tme-sierra-flux_2020-04-29.csv")

# Read storage duration respiration data
tme.trt.ts <- rbind(read.csv("../data/derived/arc-tme-t2-flux/arc-tme-t2-flux-1_2020-04-27.csv", stringsAsFactors = FALSE),
                        read.csv("../data/derived/arc-tme-t2-flux/arc-tme-t2-flux-2_2020-04-27.csv", stringsAsFactors = FALSE))

# Read miscellaneous storage treatment data
tme.misc <- read.csv("../data/derived/arc-tme-misc/arc-tme-misc-iso_2020-04-29.csv")
```

3. Combine and summarize data in long format to calculate respiration rates and plot over time.
```{r summarize data, include = FALSE}
# NB: set "include" to false to suppress bind_rows warning about coercing factor to char
# remove formula_check columns
invisible(list2env(lapply(list(archive.arc.ts = archive.arc.ts, 
                               rewet.ctl.ts = rewet.ctl.ts, 
                               rewet.dry.ts = rewet.dry.ts,
                               tme.trt.ts = tme.trt.ts),
                          function(x) {
                            x$formula_check <- NULL
                            return(x)
                            }),
                   envir = .GlobalEnv)
          )

# Add Treatment columns to each dataset
archive.arc.ts$Treatment <- "air-dry + storage"
archive.ctl.ts$Treatment <- "control"
tme.trt.ts$Treatment <- "storage duration"
rewet.ctl.ts$Treatment <- "control"
rewet.dry.ts$Treatment <- "air-dry"

# combine rewet data
rewet.ts <- rbind(rewet.ctl.ts, rewet.dry.ts)
rewet.ts$Experiment <- "rewet"

# add soil dry wt data
archive.arc.ts$dw_g <- jarinfo.arc[match(archive.arc.ts$SampleName, jarinfo.arc$Sample), "Soil.dry.weight_.g."]
archive.ctl.ts$dw_g <- archive.ctl.ts.wide[match(archive.ctl.ts$ID, archive.ctl.ts.wide$ID), "dry_soil_wt_corrected_g"]
rewet.ts <- merge(rewet.ts, jarinfo.rewet[ , c("ID", "Treatment", "dw_g")], by = c("ID", "Treatment"))
tme.trt.ts <- merge(tme.trt.ts, jarinfo.tme[ , c("Nr.", "dry.soil")], by.x = "SampleName", by.y = "Nr.")
colnames(tme.trt.ts)[which(names(tme.trt.ts) == "dry.soil")] <- "dw_g"

# add experiment col to archive.arc.ts, storage.trt.ts
archive.arc.ts$Experiment <- "arc"
tme.trt.ts$Experiment <- "tme"

# combine archive data and Exploratories storage duration data
archive.ts <- rbind(archive.ctl.ts[which(archive.ctl.ts$Treatment != "storage duration"), ], 
                    archive.arc.ts[ , names(archive.ctl.ts)])
tme.ts <- rbind(archive.ctl.ts[which(archive.ctl.ts$Treatment == "storage duration"), ],
                tme.trt.ts[ , names(archive.ctl.ts)])

# bind tme.ctl.flx.sra data to tme.ts
# all other tme.flx data is already summarized
tme.ctl.flx.sra$Experiment <- "tme"
tme.ctl.flx.sra$Treatment <- "control"
tme.ctl.flx.sra$dw_g <- tme.ctl.soil.sra[match(tme.ctl.flx.sra$Jar_ID, tme.ctl.soil.sra$Jar_ID), "dw_g"]
tme.ts <- rbind(tme.ts, tme.ctl.flx.sra[ , names(tme.ts)])

# Test that required names are present
nms <- c("SampleName", "ID", "Experiment", "Treatment", "Period", "timepoint_cmtv",  "time_d", "time_d_cmtv", "mgCO2_jar", "dw_g")
invisible(lapply(list(archive.ts = archive.ts,
                      rewet.ts = rewet.ts,
                      tme.ts),
       function(x) {
         ifelse(!is.na(match(nms, names(x))), "yes", "no")
       }
       ))

# combine all data and remove time points without CO2 measurements
ts <- bind_rows(archive.ts[ , nms], rewet.ts[ , nms], tme.ts[ , nms])
if(length(which(is.na(ts$mgCO2_jar))) > 0) {
  ts <- ts[-which(is.na(ts$mgCO2_jar)), ]
}

# add C content
# xplr
ts$gC_gS <- xplr.cn[match(ts$ID, xplr.cn$EP_Plotid), "Total_C"]*10^-3
# duke
ts[which(is.na(ts$gC_gS)), "gC_gS"] <- tme.ctl.duk[match(ts[which(is.na(ts$gC_gS)), "ID"], tme.ctl.duk$ID), "C_g_kg"]*10^-3
# sierras
ts[which(is.na(ts$gC_gS)), "gC_gS"] <- tme.ctl.soil.sra[match(ts[which(is.na(ts$gC_gS)), "ID"], tme.ctl.soil.sra$ID), "gC_kgS"]*10^-3
# oak ridge
ts[which(is.na(ts$gC_gS)), "gC_gS"] <- tme.ctl.ornl[match(ts[which(is.na(ts$gC_gS)), "ID"], tme.ctl.ornl$ID), "C_g_kg"]*10^-3

# add Type
ts$Type <- ifelse(grepl("G", ts$ID), "G", "F")

# calculate CO2 respiration in mgCO2-C per g soil C
# NB: 12/44 = mass ratio of C in CO2
ts$mgCO2.C_gC <- (ts$mgCO2_jar*(12/44))/(ts$dw_g*ts$gC_gS)
ts$mgCO2.C_gC_d <- ts$mgCO2.C_gC/ts$time_d

# summarize CO2 release data for each timepoint, by experiment, treatment, and type
ts.avg <- data.frame(na.omit(ts) %>%
  select(Experiment, Treatment, Type, timepoint_cmtv, mgCO2.C_gC_d, time_d_cmtv)  %>%
  group_by(Experiment, Treatment, Type, timepoint_cmtv) %>%
  add_tally() %>%
  summarise_all(list(mean = mean, sd = sd), na.rm = TRUE) %>%
  mutate(se_slope = mgCO2.C_gC_d_sd/sqrt(n_mean),
         se_slope_l = mgCO2.C_gC_d_mean - se_slope,
         se_slope_u = mgCO2.C_gC_d_mean + se_slope))

# add timepoint zero data
tp0.f <- function(df) {
  tp0 <- df[1, ]
  tp0[ , -which(names(tp0) == "Experiment" | names(tp0) == "Treatment" | names(tp0) == "Type" | names(tp0) == "n")] <- 0
  rbind(tp0, df)
}
ts.avg$treat.Exp <- paste(ts.avg$Treatment, ts.avg$Experiment)
ts.avg <- lapply(split(ts.avg, ts.avg$treat.Exp), function(x) {
  lapply(split(x, x$Type), tp0.f)
  })
ts.avg <- bind_rows(lapply(ts.avg, bind_rows))

# restore classes
ts.avg <- as.data.frame(lapply(ts.avg, type.convert))
```


4. Plots of CO~2~ fluxes over time. The final measurement points for a few samples which took >18 days to reach CO~2~ targets are excluded for display reasons. Respiration rates for those samples remained flat. 
* While daily measurements were made during the pre-incubation period for the majority of samples, only the cumulative pre-incubation flux was measued for the 2011 control samples
* Consequently, the respiration rate for the pre-incubation period was calculated cumulatively for all samples (Fig. 1)
```{r CO2-flux-plots}
ts.pre <- ts[ts$Period == "pre", ]
ts.pre.avg <- ts.pre %>%
  filter(time_d_cmtv > 3.5) %>%
  select(Experiment, Treatment, Type, timepoint_cmtv, SampleName, mgCO2.C_gC, time_d_cmtv)  %>%
  group_by(Experiment, Treatment, Type, timepoint_cmtv, SampleName) %>%
  summarise_all(list(mean = mean), na.rm = TRUE) %>%
  mutate(mgCO2.C_gC_d = mgCO2.C_gC_mean/time_d_cmtv_mean) %>%
  select(-SampleName, -mgCO2.C_gC_mean, Experiment, Treatment, Type, mgCO2.C_gC_d, time_d_cmtv_mean) %>%
  group_by(Experiment, Treatment, Type) %>%
  add_tally() %>%
  summarise_all(list(mean = mean, sd = sd), na.rm = TRUE) %>%
  mutate(se_slope = mgCO2.C_gC_d_sd/sqrt(n_mean),
         se_slope_l = mgCO2.C_gC_d_mean - se_slope,
         se_slope_u = mgCO2.C_gC_d_mean + se_slope)

ts.inc.avg <- ts %>%
  filter(Period == "inc") %>%
  select(Experiment, Treatment, Type, timepoint_cmtv, mgCO2.C_gC_d, time_d_cmtv)  %>%
  group_by(Experiment, Treatment, Type, timepoint_cmtv) %>%
  add_tally() %>%
  summarise_all(list(mean = mean, sd = sd), na.rm = TRUE) %>%
  mutate(se_slope = mgCO2.C_gC_d_sd/sqrt(n_mean),
         se_slope_l = mgCO2.C_gC_d_mean - se_slope,
         se_slope_u = mgCO2.C_gC_d_mean + se_slope)

colnames(ts.pre.avg)[which(names(ts.pre.avg) %in% c("timepoint_cmtv_mean",
                       "time_d_cmtv_mean_mean",
                       "time_d_cmtv_mean_sd"))] <- c("timepoint_cmtv",
                                                    "time_d_cmtv_mean",
                                                    "time_d_cmtv_sd")

ts.pre.avg <- ts.pre.avg[ , -which(!names(ts.pre.avg) %in% names(ts.inc.avg))]

ts.avg2 <- data.frame(rbind(ts.pre.avg, ts.inc.avg))
ts.avg2 <- ts.avg2[complete.cases(ts.avg2), ]

# add timepoint zero data
ts.avg2$treat.Exp <- paste(ts.avg2$Treatment, ts.avg2$Experiment)
ts.avg2 <- lapply(split(ts.avg2, ts.avg2$treat.Exp), function(x) {
  lapply(split(x, x$Type), tp0.f)
  })
ts.avg2 <- bind_rows(lapply(ts.avg2, bind_rows))

# restore classes
ts.avg2 <- as.data.frame(lapply(ts.avg2, type.convert))

ts.avg2 %>% 
  filter(Experiment != "tme") %>%
  ggplot(., aes(time_d_cmtv_mean, mgCO2.C_gC_d_mean)) +
  geom_vline(xintercept = 4, color="gray") +
  geom_ribbon(aes(ymin = se_slope_l, ymax = se_slope_u, fill = Type, linetype = Treatment, alpha = Treatment)) +
  geom_line(aes(color = Type, linetype = Treatment)) +
  facet_grid(rows = vars(Experiment),
             labeller = labeller(Experiment = c("arc" = "2011", "rewet" = "2019"))) +
  scale_x_continuous(limits = c(0,18)) +
  scale_color_manual(name = 'Landuse',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  scale_fill_manual(name = 'Std. Err.',
                    values =c('F'='#a35513','G'='#1361a3'),
                    labels = c('Forest','Grassland')) +
  scale_alpha_manual(name = 'Treatment',
                     values = c("air-dry + storage" = .2, 
                                "control" = .4,
                                "air-dry" = .2)) +
  scale_linetype_manual(name = 'Treatment',
                        values = c("air-dry + storage" = 'dashed', 
                                   "control" ='solid',
                                   "air-dry" = "dashed")) +
  ylab(expression('Respiration Rate (mgCO'[2]*'-C gC'^-1*'d'^-1*')')) +
  xlab("Time (days)") +
  theme_bw() +
  theme(panel.grid = element_blank())
```


* However, daily measurements made for the remaining samples shows that the initial rewetting pulse was much substantially greater for the treatment samples in both 2011 and 2019 (Fig. 2)
```{r CO2 fluxes plots}
ts.avg %>% 
  filter(Experiment != "tme") %>%
  ggplot(., aes(time_d_cmtv_mean, mgCO2.C_gC_d_mean)) +
  geom_vline(xintercept = 4, color="gray") +
  geom_ribbon(aes(ymin = se_slope_l, ymax = se_slope_u, fill = Type, linetype = Treatment, alpha = Treatment)) +
  geom_line(aes(color = Type, linetype = Treatment)) +
  facet_grid(rows = vars(Experiment),
             labeller = labeller(Experiment = c("arc" = "2011", "rewet" = "2019"))) +
  scale_x_continuous(limits = c(0,18)) +
  # scale_y_continuous(limits = c(0,30)) +
  scale_color_manual(name = 'Landuse',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  scale_fill_manual(name = 'Std. Err.',
                    values =c('F'='#a35513','G'='#1361a3'),
                    labels = c('Forest','Grassland')) +
  scale_alpha_manual(name = 'Treatment',
                     values = c("air-dry + storage" = .2, 
                                "control" = .4,
                                "air-dry" = .2)) +
  scale_linetype_manual(name = 'Treatment',
                        values = c("air-dry + storage" = 'dashed', 
                                   "control" ='solid',
                                   "air-dry" = "dashed")) +
  ylab(expression('Respiration Rate (mgCO'[2]*'-C gC'^-1*'d'^-1*')')) +
  xlab("Time (days)") +
  theme_bw() +
  theme(panel.grid = element_blank())
```

# Isotope data

1. Read in isotope data from various sources. First load helper function 'read_jena_ams_results.R"
```{r load ams-jena-ingest fx}
source("./utilities/jena_ams_ingest.R")
```


2. Next read in data from the appropriate directories in 'data/raw'.
```{r read iso data}
# identify subdirectories in 'raw' directory with "ams_jena" in name
ams_jena_results_dirs <- list.files("../data/raw", pattern = "ams_jena_results", full.names = TRUE)
ams_results_ls <- lapply(seq_along(ams_jena_results_dirs), function(i) {
  read_jena_ams_results(ams_jena_results_dirs[i])
})
names(ams_results_ls) <- list.files("../data/raw", pattern = "ams_jena_results")
```


3. Create a "tidy" style template for the data, i.e. variables in columns. 
* Key variables are as follows:
  * SampleName (incorporates lab rep and treatment, e.g. "HEG10-1_dry")
  * ID (plot IDs, e.g. for "HEG10" for Exploratory samples)
  * Treatment (3 treatments: air-dry, air-dry + storage, storage duration; + controls)
  * Type (2 levels: F = forest, G = grassland)
  * Period (incubation period, 2 levels: pre = preincubation, inc = equilibrium incubation)
  * Experiment (3 levels: arc = air-dry + storage, rewet = air-dry/rewet, time = storage duration)

* Observational columns include:
  * d14c ($\Delta$^14^C-CO~2~)
  * d13c ($\delta$^13^C-CO~2~)
  * C_g_kg (C content)
  * dw_g (dry weight)
  * mgCO2.C_gS (mg CO~2~-C respired g^-1^ soil Period^-1^)
  * time_d (days in incubation period prior to measurement)
  * dH2O_grav (percent change in gravimetric water content due to laboratory moisture adjustment)
  * dH2O_whc (percent change in water holding capacity due to laboratory moisture adjustment)
  
```{r arc and rewet experiment template}
unique.id.n <- c(length(unique(archive.ctl.ts$ID)),
                 length(unique(archive.arc.ts$SampleName)),
                 length(unique(rewet.ctl.ts$SampleName)),
                 length(unique(rewet.dry.ts$SampleName)))
pre.14c <- data.frame(SampleName = c(as.character(unique(archive.ctl.ts$ID)),
                                     as.character(unique(archive.arc.ts$SampleName)),
                                     as.character(unique(rewet.ctl.ts$SampleName)),
                                     as.character(unique(rewet.dry.ts$SampleName))))
pre.14c$ID <- substr(pre.14c$SampleName, 1, 5)
pre.14c$Type <- ifelse(grepl("G", pre.14c$ID), "G", "F")
pre.14c$Experiment <- c(rep("arc", unique.id.n[1] + unique.id.n[2]),
                       rep("rewet", unique.id.n[3] + unique.id.n[4]))
pre.14c$Treatment <- c(rep("control", unique.id.n[1]),
                       rep("air-dry + storage", unique.id.n[2]),
                       rep("control", unique.id.n[3]),
                       rep("air-dry", unique.id.n[4]))
pre.14c$Period <- "pre"

# create inc period df as duplicate of pre-inc df, but test that it is the right shape first
if(nrow(pre.14c) == sum(unique.id.n) & ncol(pre.14c) == 6) {
  inc.14c <- pre.14c
  inc.14c$Period <- "inc"
} else {
  "pre.14c data frame malformed"
}
all.14c.template <- rbind(pre.14c, inc.14c)
```

```{r tme experiment template}
unique.id.tme.n <- c(length(unique(tme.ctl.duk$SampleName)),
                     length(unique(tme.ctl.ornl$SampleName)),
                     length(unique(tme.ctl.flx.sra$SampleName)),
                     length(unique(tme.misc$SampleName)),
                     length(unique(archive.ctl.ts.wide[archive.ctl.ts.wide$Experiment == "tme", "ID"])),
                     length(unique(tme.trt.ts$SampleName)))
pre.tme.ctl <- data.frame(SampleName = c(as.character(unique(tme.ctl.duk$SampleName)),
                                         as.character(unique(tme.ctl.ornl$SampleName)),
                                         as.character(unique(tme.ctl.flx.sra$SampleName)),
                                         as.character(unique(tme.misc$SampleName)),   as.character(unique(archive.ctl.ts.wide[archive.ctl.ts.wide$Experiment == "tme", "ID"]))))
pre.tme.ctl$Site <- c(rep("Duke", unique.id.tme.n[1]),
                      rep("ORNL", unique.id.tme.n[2]),
                      rep("Sierra", unique.id.tme.n[3]),
                      rep("misc", unique.id.tme.n[4]),
                      rep("Xplr", unique.id.tme.n[5]))


# Duke, ORNL, misc, and Xplr IDs
pre.tme.ctl[pre.tme.ctl$Site == "Duke" | 
            pre.tme.ctl$Site == "ORNL" |
            pre.tme.ctl$Site == "misc" |
            pre.tme.ctl$Site == "Xplr", "ID"] <- as.character(pre.tme.ctl[pre.tme.ctl$Site == "Duke" |
                                                                            pre.tme.ctl$Site == "ORNL" |
                                                                            pre.tme.ctl$Site == "misc" |
            
                                                                            pre.tme.ctl$Site == "Xplr", "SampleName"])

# Sierra IDs
pre.tme.ctl[pre.tme.ctl$Site == "Sierra", "ID"] <-
  substr(pre.tme.ctl[pre.tme.ctl$Site == "Sierra", "SampleName"], 10, 11)

# subdivide ORNL into WB and TVA
pre.tme.ctl[pre.tme.ctl$Site == "ORNL", "Site"] <- ifelse(grepl("TVA", pre.tme.ctl[pre.tme.ctl$Site == "ORNL", "ID"]), "Tennessee Valley", "Walker Branch")

# subdivide Sierra into Musick and Shaver
pre.tme.ctl[pre.tme.ctl$Site == "Sierra", "Site"] <- ifelse(grepl("M", pre.tme.ctl[pre.tme.ctl$Site == "Sierra", "ID"]), "Musick", "Shaver")

# add treatment col
pre.tme.ctl$Treatment <- "control"

pre.tme.trt <- data.frame(SampleName = unique(tme.trt.ts$SampleName))
pre.tme.trt$ID <- substr(pre.tme.trt$SampleName, 1, nchar(as.character(pre.tme.trt$SampleName))-2)
pre.tme.trt <- rbind(pre.tme.trt, pre.tme.ctl[pre.tme.ctl$Site == "misc", c("ID", "SampleName")])
pre.tme.trt$Treatment <- "storage duration"
pre.tme.trt$Site <- pre.tme.ctl[match(pre.tme.trt$ID, pre.tme.ctl$ID), "Site"]

pre.tme.14c <- rbind(pre.tme.ctl, pre.tme.trt)
pre.tme.14c$Type <- ifelse(grepl("G", pre.tme.14c$SampleName), "G", "F")
pre.tme.14c$Period <- "pre"
pre.tme.14c$Experiment <- "tme"

# create inc period df
inc.tme.14c <- pre.tme.14c
inc.tme.14c$Period <- "inc"
tme.14c.template <- rbind(pre.tme.14c, inc.tme.14c)
```


4. Summarize observational data from timeseries data by unique IDs (SampleName).
```{r combine soil obs data}
# summarize time and co2 data by experiment, period, treatment, and sample name
ts.epsn.avg <- ts %>%
  select(Experiment, ID, Period, Treatment, SampleName, gC_gS, dw_g, mgCO2.C_gC, time_d, time_d_cmtv)  %>%
  group_by(Experiment, ID, Period, Treatment, SampleName) %>%
  summarise_all(list(max = max), na.rm = TRUE)
ts.epsn.avg <- rename_at(ts.epsn.avg, vars(contains('max')), list(~sub("_max", "", .)))
```

5. Create helper functions for decay correction, converting $\Delta$^14^C to fraction modern, and cleaning up extraneous values in raw ^14^C data. Archived sample $\Delta$^14^C data should to be corrected for decay since the year of collection. (Although the correction is very small and likely insignificant, I will do it anyway).
* decay correction formula is:
$$1000 \cdot \left( (FM \cdot e^{\frac{-year_{sampled} + 1950}{8267}}) - 1 \right)$$
```{r 14C helper fxs}
# create decay correction function  
decay.corr.fx <- function(year_sampled, fraction_modern) {
  ifelse(is.na(fraction_modern), NA, 1000*((fraction_modern*exp((-year_sampled + 1950)/8267))-1))
}

# d14C to fraction modern function
d14c.fm.fx <- function(year_sampled, d14c) {
  ((d14c / 1000) + 1) / exp(0.00012097 * (1950 - year_sampled))
  }

# define helper function for stripping out "NA" and non-Exploratories data
extra.clean.fx <- function(df) {
  df <- df[which(grepl("HEG", df$Probe) |
                   grepl("HEW", df$Probe) |
                   grepl("SEG", df$Probe) |
                   grepl("SEW", df$Probe)), ]
  if(length(grep("Experiment", names(df))) > 0) {
     df <- df[which(df$Experiment != "tme"), ] 
  }
  ix <- which(is.na(df$F14C) & is.na(df$`∆14C.(‰)`))
  if(length(ix) > 1) {
    df <- df[-ix, ]
  }
  return(df)
}

# extract storage duration data separately
tme.14c.fx <- function(df) {
  df$SampleName <- df$Probe
  df$Period <- "inc"
  if(length(grep("Experiment", names(df))) > 0) {
     df <- df[which(df$Experiment == "tme"), ] 
  }
  return(df)
}
```

6. Clean up ^14^C data and add external data points (tme experiment, Xplr control samples)
```{r clean up 14C data}
# Need to add SampleName col and variable columns Treatment, Period, and Experiment to all data frames in ams_results_ls 

# First add 14C and 13C data from Exploratories control samples to ams_results_ls
# note that neither 13C nor 14C was measured for the pre-incubation period for these samples
# get names for ams_results data frames
results.14c.nms <- names(ams_results_ls$`ams_jena_results-co2_2019-01-28`$`Beem-Miller_9.xlsx`)
# get names for relevant Experiment/Treatment/Period combination
arc_ctl_inc_SN <- as.character(all.14c.template[all.14c.template$Experiment == "arc" & all.14c.template$Treatment == "control" & all.14c.template$Period == "inc", "SampleName"])
tme_ctl_inc_xplr_SN <- unlist(lapply(ams_results_ls[1:2], function(x) unlist(lapply(x, "[[", "Probe"))))
tme_ctl_inc_xplr_SN <- tme_ctl_inc_xplr_SN[grep("HE", tme_ctl_inc_xplr_SN)]
if(length(tme_ctl_inc_xplr_SN) == 14) {
  tme_ctl_inc_xplr_SN <- unique(substr(tme_ctl_inc_xplr_SN, 1, nchar(tme_ctl_inc_xplr_SN)-2))
}

# add sublist for Exploratories 2011 (control) data and data frames for above subsets
ams_results_ls$exploratories_ischoening_14c.co2_2011$arc_ctl_inc <- data.frame(matrix(nrow = length(arc_ctl_inc_SN), ncol = length(results.14c.nms)))
ams_results_ls$exploratories_ischoening_14c.co2_2011$tme_ctl_inc_xplr <- data.frame(matrix(nrow = length(tme_ctl_inc_xplr_SN), ncol = length(results.14c.nms)))

# add names and fill SampleNames for new data frames
ams_results_ls$exploratories_ischoening_14c.co2_2011 <- lapply(ams_results_ls$exploratories_ischoening_14c.co2_2011, function(x) {
  names(x) <- results.14c.nms
  return(x)
  }) 
ams_results_ls$exploratories_ischoening_14c.co2_2011$arc_ctl_inc[ , c("SampleName","Probe")] <- arc_ctl_inc_SN
ams_results_ls$exploratories_ischoening_14c.co2_2011$tme_ctl_inc_xplr[ , c("SampleName","Probe")] <- tme_ctl_inc_xplr_SN

# fill 14C data
ams_results_ls$exploratories_ischoening_14c.co2_2011 <- lapply(ams_results_ls$exploratories_ischoening_14c.co2_2011, function(df) {
  df$`∆14C.(‰)` <- xplr.ctl.iso[match(df$Probe, xplr.ctl.iso$EP_Plotid), "14c_CO2"]
  df$Treatment <- "control"
  df$Period <- "inc"
  return(df)
})

# add Experiment
ams_results_ls$exploratories_ischoening_14c.co2_2011$arc_ctl_inc$Experiment <- "arc"
ams_results_ls$exploratories_ischoening_14c.co2_2011$tme_ctl_inc_xplr$Experiment <- "tme"


# `ams_jena_results-co2_2018-10-12` and `ams_jena_results-co2_2018-11-09`
# `Beem-Miller_4.xlsx`, `Beem-Miller_5.xlsx`; `Beem-Miller_6.xlsx`, `Beem-Miller_7.xlsx`
# Add treatment, experiment, period cols
# Experiment = tme
# Treatment = storage duration
# Period = inc
ams_results_ls[1:2] <- lapply(ams_results_ls[1:2], function(ls) lapply(ls, function(df) {
  df$Experiment <- "tme"
  df$Treatment <- "storage duration"
  df$Period <- "inc"
  return(df)
  })
)

# `ams_jena_results-co2_2019-01-28`
# `Beem-Miller_9.xlsx`
# Experiment = arc
# Treatment = air-dry + storage
# Period = pre
ams_results_ls$`ams_jena_results-co2_2019-01-28`$`Beem-Miller_9.xlsx`$SampleName <- ams_results_ls$`ams_jena_results-co2_2019-01-28`$`Beem-Miller_9.xlsx`$Probe
ams_results_ls$`ams_jena_results-co2_2019-01-28`$`Beem-Miller_9.xlsx`$Treatment <- "air-dry + storage"
ams_results_ls$`ams_jena_results-co2_2019-01-28`$`Beem-Miller_9.xlsx`$Period <- "pre"
ams_results_ls$`ams_jena_results-co2_2019-01-28`$`Beem-Miller_9.xlsx`$Experiment <- "arc"

# ams_results_ls$`ams_jena_results-co2_2019-02-25`$
# Archived_Incubations_30.xlsx
# Experiment = arc
# Treatment = air-dry + storage
# Period = inc
ams_results_ls$`ams_jena_results-co2_2019-02-25`$Archived_Incubations_30.xlsx$SampleName <- substr(ams_results_ls$`ams_jena_results-co2_2019-02-25`$Archived_Incubations_30.xlsx$Probe, 1, 9)
ams_results_ls$`ams_jena_results-co2_2019-02-25`$Archived_Incubations_30.xlsx$Treatment <- "air-dry + storage"
ams_results_ls$`ams_jena_results-co2_2019-02-25`$Archived_Incubations_30.xlsx$Period <- "inc"
ams_results_ls$`ams_jena_results-co2_2019-02-25`$Archived_Incubations_30.xlsx$Experiment <- "arc"

# ams_results_ls$`ams_jena_results-co2_2019-04-30`
# `Rost_31-2.xlsx`
# Experiment = arc
# Treatment = air-dry + storage
# Period = inc, pre (SEW11-2-2)
ams_results_ls$`ams_jena_results-co2_2019-04-30`$`Rost_31-2.xlsx`$SampleName <- substr(ams_results_ls$`ams_jena_results-co2_2019-04-30`$`Rost_31-2.xlsx`$Probe, 1, 9)
ams_results_ls$`ams_jena_results-co2_2019-04-30`$`Rost_31-2.xlsx`$Treatment <- "air-dry + storage"
ams_results_ls$`ams_jena_results-co2_2019-04-30`$`Rost_31-2.xlsx`$Experiment <- "arc"
ams_results_ls$`ams_jena_results-co2_2019-04-30`$`Rost_31-2.xlsx`[2:9, "Period"] <- "inc"
ams_results_ls$`ams_jena_results-co2_2019-04-30`$`Rost_31-2.xlsx`[10, "Period"] <- "pre"

# ams_results_ls$`ams_jena_results-co2_2020-02-13`
# Rost_47.xlsx 
# Experiment = rewet
# Treatment = control
# Period = pre
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_47.xlsx$SampleName <- substr(ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_47.xlsx$Probe, 1, 7)
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_47.xlsx$Treatment <- "control"
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_47.xlsx$Experiment <- "rewet"
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_47.xlsx$Period <- "pre"
# Rost_50.xlsx
# Experiment = rewet
# Treatment = control
# Period = inc
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_50.xlsx$SampleName <- substr(ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_50.xlsx$Probe, 1, 7)
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_50.xlsx$Treatment <- "control"
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_50.xlsx$Experiment <- "rewet"
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_50.xlsx$Period <- "inc"
# Rost_52.xlsx
# Experiment = rewet
# Treatment = air-dry
# Period = pre, inc
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_52.xlsx$SampleName <- substr(ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_52.xlsx$Probe, 1, 11)
pre.ix <- which(grepl("PRE", ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_52.xlsx$Probe))
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_52.xlsx$Treatment <- "air-dry"
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_52.xlsx$Experiment <- "rewet"
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_52.xlsx[pre.ix, "Period"] <- "pre"
ams_results_ls$`ams_jena_results-co2_2020-02-13`$Rost_52.xlsx[-pre.ix, "Period"] <- "inc"
```

7. Combine data.
```{r combine 14C soil obs data}
# flatten list and remove AMS blanks, NA samples, and non-Exploratory data
ams_results_arc_rewet <- lapply(ams_results_ls[3:7], function(x) bind_rows(x))
ams_results_arc_rewet.df <- bind_rows(lapply(ams_results_arc_rewet, function(x) extra.clean.fx(x)))

# flatten external data list
ams_results_tme <- lapply(ams_results_ls[c(1:2,7)], function(x) lapply(x, tme.14c.fx))
ams_results_tme.df <- bind_rows(lapply(ams_results_tme, function(x) bind_rows(x)))

# cut out non-relevant AMS results columns
nms.ams.vars <- c("F14C", "err", "∆14C.(‰)", "err.(‰)",
                 "SampleName", "Treatment", "Period", "Experiment")
ams_results_arc_rewet.df <- ams_results_arc_rewet.df[ , nms.ams.vars]
colnames(ams_results_arc_rewet.df) <- c("fm", "fm_err", "d14c", "d14c_err", nms.ams.vars[5:8])

ams_results_tme.df <- ams_results_tme.df[ , nms.ams.vars]
colnames(ams_results_tme.df) <- c("fm", "fm_err", "d14c", "d14c_err", nms.ams.vars[5:8])
ams_results_tme.df$ID <- substr(ams_results_tme.df$SampleName, 1, nchar(ams_results_tme.df$SampleName)-2)
ams_results_tme.df$Type <- ifelse(grepl("G", ams_results_tme.df$ID), "G", "F")

# merge arc and rewet 14C data with template
if(nrow(merge(all.14c.template, ams_results_arc_rewet.df, all.x = TRUE)) == nrow(all.14c.template)) {
  all.14c <- merge(all.14c.template, ams_results_arc_rewet.df, all.x = TRUE)
} else {
  cat("duplicate sample merge error")
}

# merge soil obs data with 14c data
if(nrow(merge(all.14c, ts.epsn.avg, by = c("SampleName", "Period"))) == nrow(all.14c)) {
  all.14c <- merge(all.14c, ts.epsn.avg, all.x = TRUE)
} else {
  cat("duplicate sample merge error")
}

# merge tme 14C data with template and ts.epsn.avg
tme.14c <- merge(tme.14c.template, ams_results_tme.df, all.x = TRUE)

# add ctl sample 14C data to tme.14c
# first remove NA values
tme.14c.clean <- tme.14c[which(!is.na(tme.14c$d14c) | !is.na(tme.14c$fm)), ]
# define merge fx
tme.mrg.fx <- function(x, y) {
  mrg.nms <- c("SampleName", "ID", "Treatment", "Period")
  var.nms <- c("YearSampled", "Site", "Horizon",  "d14c", "fm")
  if(length(grep("fm", names(y))) > 0) {
    df <- merge(x[ , mrg.nms], y[ , c(mrg.nms, var.nms)])
  } else {
    df <- merge(x[ , mrg.nms], y[ , c(mrg.nms, var.nms[1:4])])
    df$fm <- NA
  }
  return(df)
}

tme.14c.df <- bind_rows(lapply(list(tme.ctl.duk,
                               tme.ctl.flx.sra[which(!is.na(tme.ctl.flx.sra$d14c)), ],
                               tme.ctl.ornl,
                               tme.misc), function(i) tme.mrg.fx(tme.14c, i)))

# add storage duration data for treatment samples analyzed externally
tme.14c.df$Type <- "F"

# add YearSampled for tme.14c.clean samples
for(i in 1:nrow(tme.14c.clean)) {
  if(tme.14c.clean[i, "Site"] == "Duke") {
    tme.14c.clean[i, "YearSampled"] <- 2008
  } else if(tme.14c.clean[i, "Site"] == "Xplr") {
    tme.14c.clean[i, "YearSampled"] <- 2011
  } else if(tme.14c.clean[i, "Site"] == "Sierra") {
    tme.14c.clean[i, "YearSampled"] <- 2009
  } else {
    tme.14c.clean[i, "YearSampled"] <- 2004
  }
}

# Add Horizon to tme.14c.clean
tme.14c.clean$Horizon <- ifelse(tme.14c.clean$ID == "MB" | tme.14c.clean$ID == "SB", "B", "A")

# combine tme data
tme.14c <- bind_rows(tme.14c.clean[ , names(tme.14c.df)], tme.14c.df)
tme.14c$Experiment <- "tme"

# remove outlier (MA-1)
# *** Need to clarify method for determining that this is an outlier *** 
tme.14c <- tme.14c[-which(tme.14c$SampleName == "MA-1" | tme.14c$SampleName == "HEW49-1" | tme.14c$SampleName == "HEW26-1"), ]


# Add missing variable columns to all.14c
# add YearSampled column
all.14c$YearSampled <- ifelse(all.14c$Experiment == "arc", 2011, 2019)

# add Site and Horizon columns
all.14c$Site <- ifelse(grepl("HE", all.14c$ID), "Hainich", "Schorfheide")
all.14c$Horizon <- "A"


# combine tme.14c and all.14c into a list and calculate fraction modern from d14C as needed
all.14c.ls <- lapply(list(tme.14c, all.14c), function(df) {
  df$fm <- ifelse(is.na(df$fm),
                  d14c.fm.fx(df$YearSampled, df$d14c),
                  df$fm)
  df$d14c_corr <- decay.corr.fx(df$YearSampled, df$fm)
  return(df)
})

all.14c.sum <- lapply(all.14c.ls, function(df) {
  df.sum <- select(df, -SampleName) %>%
    group_by(ID, Type, Period, Treatment, Experiment, Site, Horizon, YearSampled) %>%
    summarise_all(list(mean = mean, max = max, min = min), na.rm = TRUE)
  df.sum[mapply(is.infinite, df.sum)] <- NA
  return(df.sum)
})
all.14c.sum <- rapply(all.14c.sum, f = function(x) ifelse(is.nan(x), NA , x), how = "replace")
all.14c.sum.df <- bind_rows(all.14c.sum)

# remove unnecessary objects
```

8. Count number of ^14^C observations for checking plots.
```{r counts for checking plots, include = FALSE}
# count !is.na(14C) cases by SampleName
all.14c[-which(is.na(all.14c$d14c_corr)),] %>%
  group_by(Experiment, Treatment, Period, Type) %>%
  count()
# by ID
all.14c[-which(is.na(all.14c$d14c_corr)),] %>%
  group_by(Experiment, Treatment, Period, Type) %>%
  distinct(ID) %>%
  count()
```

9. Plot pre-incubation period $\Delta$^14^C against equilibrium respiration period $\Delta$^14^C.
* Points are means of duplicate lab reps and error bars are min and max (except for the 2011 control samples, which were not replicated)
* Pre-incubation $\Delta$^14^C was not measured for the 2011 control samples.
* Relative outlier point is the very negative (mean = -51.1‰) HEW22 pre-incubation control samples from the 2019 air-dry experiment.
* Samples from three of the forest plots of the 2011 treatment samples failed to accumulate enough CO~2~ to measure ^14^C.
```{r pre vs inc 14c plot}
all.14c.pre.inc <- merge(all.14c.sum.df[all.14c.sum.df$Period == "pre", c("ID", "Experiment", "Type", "Treatment", "d14c_corr_mean", "d14c_corr_min", "d14c_corr_max")],
                         all.14c.sum.df[all.14c.sum.df$Period == "inc", c("ID", "Experiment", "Type", "Treatment", "d14c_corr_mean", "d14c_corr_min", "d14c_corr_max")],
                         by = c("ID", "Experiment", "Type", "Treatment"),
                         suffixes = c("_pre", "_inc"))
all.14c.pre.inc$treat.bi <- ifelse(all.14c.pre.inc$Treatment == "control", "control", "treatment")

all.14c.pre.inc %>%
  filter(Experiment != "tme") %>%
  ggplot(., aes(d14c_corr_mean_pre, d14c_corr_mean_inc, color = Type, shape = Treatment)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_vline(xintercept = 0, color = "gray") +
  geom_hline(yintercept = 0, color = "gray") +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = d14c_corr_min_inc, 
        ymax = d14c_corr_max_inc, 
        color = Type), 
    width = .25) +
  geom_errorbarh(
    aes(xmin = d14c_corr_min_pre, 
        xmax = d14c_corr_max_pre, 
        color = Type), 
    height = .9) +
  scale_color_manual(name = 'Landuse',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  scale_shape_manual(name = 'Treatment',
                     values = c("control" = 16,
                                "air-dry" = 1,
                                "air-dry + storage" = 0)) +
  scale_x_continuous(limits = c(-60, 115)) +
  scale_y_continuous(limits = c(-60, 115)) +
  xlab(expression('Pre-incubation '*Delta*''^14*'C (‰)')) +
  ylab(expression('Equilibrium respiration '*Delta*''^14*'C (‰)')) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.key.height=unit(.8, "cm"),
        aspect.ratio = 1)
```

10. Plot $\Delta$^14^C against proportion of soil C respired by experiment, land cover, and sampling period.
* First figure shows data averaged by plot
* (Note that pre-incubation $\Delta$^14^C was not measured for the 2011 control samples)
* Limits exclude outlier point (HEW22 control pre-incubation) for improved legibility
```{r plot 14c by c resp}
# collapse treatment variable into binary control/treatment
all.14c.sum.df$treat.bi <- ifelse(all.14c.sum.df$Treatment == "control", "control", "treatment")

# HEW22 control pre point excluded to make figure more readable 
all.14c.sum.df %>%
  filter(Experiment != "tme") %>%
  mutate(ID.treat = paste(ID, Treatment)) %>%
  mutate(ID.treat.Period = paste(ID.treat, Period)) %>%
  filter(ID.treat.Period != "HEW22 control pre") %>%
  mutate(treat.Period = paste(treat.bi, Period)) %>%
  ggplot(., aes(mgCO2.C_gC_mean, d14c_corr_mean, color = Type, shape = treat.Period)) +
  geom_point(size = 3) +
  geom_path(aes(group = ID.treat)) +
  geom_errorbar(
    aes(ymin = d14c_corr_min, 
        ymax = d14c_corr_max, 
        color = Type), 
    width = .25) +
  geom_errorbarh(
    aes(xmin = mgCO2.C_gC_min, 
        xmax = mgCO2.C_gC_max, 
        color = Type), 
    height = .9) +
  scale_color_manual(name = 'Landuse',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  scale_shape_manual(name = 'Sampling period (treatment)',
                     values =c(16, 17, 1, 2),
                     labels = c("control pre" = 'pre-incubation\n(control)', 
                                "control inc" = 'equilibrium respiration\n(control)',
                                "treatment pre" = 'pre-incubation\n(air-dry/air-dry + storage)',
                                "treatment inc" = 'equilibrium respiration\n(air-dry/air-dry + storage)')) +
  facet_grid(cols = vars(Type), rows = vars(YearSampled)) +
  xlab(expression('Respired C (mgCO'[2]*'-C'*' g soil C'^-1*')')) +
  ylab(expression('mean '*Delta*''^14*'C (‰)')) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.key.height=unit(.8, "cm"))
```

* Second figure shows data averaged by land use and treatment within sampling periods
```{r plot C resp vs 14C averaged by land use and treatment}
# average control and treatment data
# remove HEW22 control pre
# note that upper/lower limits = 2xSE
all.14c.sum2 <- all.14c.ls[[2]] %>%
  mutate(ID.treat.Period = paste(ID, Treatment, Period)) %>%
  filter(ID.treat.Period != "HEW22 control pre") %>%
  select(Type, Period, Treatment, Experiment, YearSampled, d14c_corr, mgCO2.C_gC) %>%
  group_by(Type, Period, Treatment, Experiment, YearSampled) %>%
  add_count() %>%
  summarize_all(list(mean = mean, sd = sd), na.rm = TRUE) %>%
  mutate(d14c_corr_se = d14c_corr_sd/n_mean,
         d14c_corr_u = d14c_corr_mean + d14c_corr_se*2,
         d14c_corr_l = d14c_corr_mean - d14c_corr_se*2,
         mgCO2.C_gC_se = mgCO2.C_gC_sd/n_mean,
         mgCO2.C_gC_u = mgCO2.C_gC_mean + mgCO2.C_gC_se*2,
         mgCO2.C_gC_l = mgCO2.C_gC_mean - mgCO2.C_gC_se*2) 

all.14c.sum2$treat.bi <- ifelse(all.14c.sum2$Treatment == "control", "control","treatment")
all.14c.sum2 %>%
  mutate(treat.Period = paste(treat.bi, Period)) %>%
  ggplot(., aes(mgCO2.C_gC_mean, d14c_corr_mean, color = Type, shape = treat.Period)) +
  geom_point(size = 3) +
  geom_path(aes(group = Treatment)) +
  geom_errorbar(
    aes(ymin = d14c_corr_u, 
        ymax = d14c_corr_l, 
        color = Type), 
    width = .25) +
  geom_errorbarh(
    aes(xmin = mgCO2.C_gC_u, 
        xmax = mgCO2.C_gC_l, 
        color = Type), 
    height = .9) +
  scale_color_manual(name = 'Landuse',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  scale_shape_manual(name = 'Sampling period (treatment)',
                     values =c(16, 17, 1, 2),
                     labels = c("control pre" = 'pre-incubation\n(control)', 
                                "control inc" = 'equilibrium respiration\n(control)',
                                "treatment pre" = 'pre-incubation\n(air-dry/air-dry + storage)',
                                "treatment inc" = 'equilibrium respiration\n(air-dry/air-dry + storage)')) +
  facet_grid(cols = vars(Type), rows = vars(YearSampled)) +
  xlab(expression('Respired C (mgCO'[2]*'-C'*' g soil C'^-1*')')) +
  ylab(expression('mean '*Delta*''^14*'C (‰)')) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.key.height=unit(.8, "cm"))
```

* Code for a third plot with outlier data is in code chunk below, but not is currently included in report
```{r plot 14c by c resp with outlier, include = FALSE}
# average control and treatment data
all.14c.sum3 <- select(all.14c.ls[[2]], Type, Period, Treatment, Experiment, YearSampled, d14c_corr, mgCO2.C_gC) %>%
  group_by(Type, Period, Treatment, Experiment, YearSampled) %>%
  add_count() %>%
  summarize_all(list(mean = mean, sd = sd), na.rm = TRUE) %>%
  mutate(d14c_corr_se = d14c_corr_sd/n_mean,
         d14c_corr_u = d14c_corr_mean + d14c_corr_se*2,
         d14c_corr_l = d14c_corr_mean - d14c_corr_se*2,
         mgCO2.C_gC_se = mgCO2.C_gC_sd/n_mean,
         mgCO2.C_gC_u = mgCO2.C_gC_mean + mgCO2.C_gC_se*2,
         mgCO2.C_gC_l = mgCO2.C_gC_mean - mgCO2.C_gC_se*2) 

all.14c.sum3$treat.bi <- ifelse(all.14c.sum3$Treatment == "control", "control","treatment")
all.14c.sum3 %>%
  mutate(treat.Period = paste(treat.bi, Period)) %>%
  ggplot(., aes(mgCO2.C_gC_mean, d14c_corr_mean, color = Type, shape = treat.Period)) +
  geom_point(size = 3) +
  geom_path(aes(group = Treatment)) +
  geom_errorbar(
    aes(ymin = d14c_corr_u, 
        ymax = d14c_corr_l, 
        color = Type), 
    width = .25) +
  geom_errorbarh(
    aes(xmin = mgCO2.C_gC_u, 
        xmax = mgCO2.C_gC_l, 
        color = Type), 
    height = .9) +
  scale_color_manual(name = 'Landuse',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  scale_shape_manual(name = 'Sampling period (treatment)',
                     values =c(16, 17, 1, 2),
                     labels = c("control pre" = 'pre-incubation\n(control)', 
                                "control inc" = 'equilibrium respiration\n(control)',
                                "treatment pre" = 'pre-incubation\n(air-dry/air-dry + storage)',
                                "treatment inc" = 'equilibrium respiration\n(air-dry/air-dry + storage)')) +
  facet_grid(cols = vars(Type), rows = vars(YearSampled)) +
  xlab(expression('Respired C (mgCO'[2]*'-C'*' g soil C'^-1*')')) +
  ylab(expression('mean '*Delta*''^14*'C (‰)')) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.key.height=unit(.8, "cm"))
```


12. Show treatment effects on observed equilibrium period ^14^C.
* Fig. 1 shows the direction of mean treatment effects over time in reference to the atmosphere (gray line). Points are means (n = 12 for 2011 treatment points, n = 9 for 2011 control points; n = 6 for all 2019 points); error bars = 2x std. err. of the mean.
```{r treatment effects by time}
# first load atmospheric data from ISRaD and bind future projection
Datm <- rbind(graven, future14C)
Datm <- Datm[Datm$Date > 1900, c("Date", "NHc14")]

# # code for plotting data by ID
# all.14c.sum %>%
#   filter(Period == "inc") %>%
#   filter(!grepl("S", ID)) %>%
#   mutate(year.ID = paste(YearSampled, ID)) %>%
#   ggplot(., aes(YearSampled, d14c_corr_mean, color = Type)) +
#   geom_point(aes(shape = Treatment), size = 3) +
#   geom_path(aes(group = year.ID), 
#             arrow = arrow(type = "open", length = unit(.1, "inches"), ends = "last"),
#             show.legend = FALSE) +
#   geom_path(data = Datm, aes(Date, NHc14), color = "darkgray") +
#   scale_color_manual(name = 'Land use',
#                      values =c('F'='#a35513','G'='#1361a3'),
#                      labels = c('Forest','Grassland')) +
#   scale_shape_manual(name = 'Treatment',
#                      values = c("air-dry" = 1, 
#                                 "air-dry + storage" = 0,
#                                 "control" = 16)) +
#   scale_x_continuous(limits = c(2010, 2020), breaks = c(2013, 2016, 2019)) +
#   scale_y_continuous(limits = c(-15, 100)) +
#   xlab('Year sampled') +
#   ylab(expression('mean '*Delta*''^14*'C (‰)')) +
#   facet_wrap(~ ID, ncol = 3) +
#   theme_bw() +
#   theme(panel.grid = element_blank())

# plot over time, averaged by Type
all.14c.sum2 %>%
  filter(Period == "inc") %>%
  mutate(year.Type = paste(YearSampled, Type)) %>%
  ggplot(., aes(YearSampled, d14c_corr_mean, color = Type)) +
  geom_point(aes(shape = Treatment), size = 3, stroke = 1.5) +
  geom_path(aes(group = year.Type), 
            arrow = arrow(type = "open", length = unit(.1, "inches"), ends = "first"),
            show.legend = FALSE,
            position = position_dodge(width = 2, preserve = "single"),
            size = 1.2) +
  geom_path(data = Datm, aes(Date, NHc14), color = "darkgray") +
  geom_errorbar(
    aes(ymin = d14c_corr_l, 
        ymax = d14c_corr_u, 
        color = Type), 
    width = .5,
    alpha = .3) +
  scale_color_manual(name = 'Land use',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  scale_shape_manual(name = 'Treatment',
                     values = c("air-dry" = 1, 
                                "air-dry + storage" = 0,
                                "control" = 16)) +
  scale_x_continuous(limits = c(2010, 2020), breaks = c(2013, 2016, 2019)) +
  scale_y_continuous(limits = c(-15, 100)) +
  xlab('Year sampled') +
  ylab(expression('mean '*Delta*''^14*'C (‰)')) +
  facet_grid(cols = vars(Type)) +
  theme_bw() +
  theme(panel.grid = element_blank())
```

* Code below plots treatment effects as means with inferential error bars (2x SE), but is not included in the report.
```{r side-by-side plot, include = FALSE}
# plot as side-by-side means w/ 2xSE
# will change to 95%CIs at some point, but need to calculate test statistic with appropriate df
all.14c.sum2 %>%
  filter(Period == "inc") %>%
  ggplot(., aes(treat.bi, d14c_corr_mean, color = Type)) +
  geom_point(aes(shape = Treatment), size = 3) +
  geom_errorbar(
    aes(ymin = d14c_corr_u, 
        ymax = d14c_corr_l, 
        color = Type), 
    width = .1) +
  scale_color_manual(name = 'Land use',
                     values =c('F'='#a35513','G'='#1361a3'),
                     labels = c('Forest','Grassland')) +
  scale_shape_manual(name = 'Treatment',
                     values = c("air-dry" = 1, 
                                "air-dry + storage" = 0,
                                "control" = 16)) +
  ylab(expression('mean '*Delta*''^14*'C (‰)')) +
  facet_grid(cols = vars(Type), rows = vars(YearSampled)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.title.x = element_blank())

```


15. Show overall effect of treatment on whole data set.
* Notes:
    + Control data shown on x-axis, treatment data shown on y-axis
    + Solid line is 1:1, dashed line is a 20 ‰ offset, and dotted line is a 40‰ offset (roughly equivalent to the atmospheric decline over five and 10 years respectively for the period 2000 to 2020)
    + Data from all three experiments conducted in this study are shown, as well as a handful of additional data points for which both control and treatment (i.e. after air-drying + storage) incubations were conducted in another laboratory (Harvard points)
    + Only A horizon data are shown here, as owing to sample availabilityonly three samples were analyzed for organic and B horizons (respectively) 
    + Points are means of replicates and error bars are min and max of replicates (number of replicates varies from 1 to 3); note that error bars are not visible at this scale for most points 
* Key messages:
    + Difference between control and treatment samples is within a 5 year range for the majority of points
    + Grassland samples tend to be above 1:1 line, forest samples below, regardless of site
    + The three German forest samples above 1:1 line were analyzed in 2019 (air-dry only treatment, this study), lending support to the interpretation that slow and fast soil C pool curves for northern hemisphere forest samples have only recently crossed or may have yet to cross.
```{r plot ctl vs trt}
# Filter to "inc", split by binary treatment variable, then merge
all.14c.ctl <- all.14c.sum.df[all.14c.sum.df$treat.bi == "control" & all.14c.sum.df$Period == "inc", ]
all.14c.ctl <- rename_at(all.14c.ctl, vars(contains('d14c')), list(~paste0(., "_ctl")))
all.14c.trt <- all.14c.sum.df[all.14c.sum.df$treat.bi == "treatment" & all.14c.sum.df$Period == "inc", ]
all.14c.trt <- rename_at(all.14c.trt, vars(contains('d14c')), list(~paste0(., "_trt")))
all.14c.ctl.trt <- inner_join(data.frame(all.14c.ctl) %>% 
                                select(ID, Experiment, Type, Site, Horizon, contains("d14c_corr")),
                              data.frame(all.14c.trt) %>% 
                                select(ID, Experiment, Type, Site, Horizon, contains("d14c_corr")),
                              by = c("ID", "Experiment", "Type", "Site", "Horizon"))

# Plot on 1:1 line w/ diagonal lines showing ‰ offsets
# first collapse sites for better visuals

loc <- list(
    "Duke" = "Duke, NC",
    "Howland" = "Howland, ME",
    "Harvard" = "Harvard, MA",
    "Hainich" = "Central Germany",
    "Schorfheide" = "Central Germany",
    "Musick" = "Sierra Nevada, CA",
    "Shaver" = "Sierra Nevada, CA",
    "Walker Branch" = "Oak Ridge, TN",
    "Tennessee Valley" = "Oak Ridge, TN")

all.14c.ctl.trt %>%
  filter(Horizon == "A") %>%
  mutate(Location = recode(Site, !!!loc)) %>%
  ggplot(., aes(d14c_corr_mean_ctl, d14c_corr_mean_trt, color = Location, shape = Type)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_abline(slope = 1, intercept = -20, linetype = "dashed") +
  geom_abline(slope = 1, intercept = 20, linetype = "dashed") +
  geom_abline(slope = 1, intercept = -40, linetype = "dotted") +
  geom_abline(slope = 1, intercept = 40, linetype = "dotted") +
  geom_vline(xintercept = 0, color = "gray") +
  geom_hline(yintercept = 0, color = "gray") +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = d14c_corr_min_trt, 
        ymax = d14c_corr_max_trt, 
        color = Location), 
    width = .25) +
  geom_errorbarh(
    aes(xmin = d14c_corr_min_ctl, 
        xmax = d14c_corr_max_ctl, 
        color = Location), 
    height = .9) +
  scale_color_manual(values = c('Duke, NC' = '#01BB97',
                                'Sierra Nevada, CA' = '#FFC107',
                                'Central Germany' = 'black',
                                'Harvard, MA' = '#1E88E5',
                                'Oak Ridge, TN' = '#D81B60')) +
  scale_shape_manual(name = 'Land use',
                     values = c('F'= 17,'G' = 16),
                     labels = c('Forest','Grassland')) +
  xlab(expression('Control '*Delta*''^14*'C (‰)')) +
  ylab(expression('Treatment '*Delta*''^14*'C (‰)')) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.key.height=unit(.8, "cm"),
        aspect.ratio = 1)

# for plotting with color = Type (currently suppressed)
# all.14c.ctl.trt %>%
#   filter(Horizon == "A") %>%
#   mutate(Location = factor(recode(Site, !!!loc), exclude = TRUE)) %>%
#   ggplot(., aes(d14c_corr_mean_ctl, d14c_corr_mean_trt, color = Type, shape = Location)) +
#   geom_abline(slope = 1, intercept = 0) +
#   geom_abline(slope = 1, intercept = -20, linetype = "dashed") +
#   geom_abline(slope = 1, intercept = 20, linetype = "dashed") +
#   geom_abline(slope = 1, intercept = -40, linetype = "dotdash") +
#   geom_abline(slope = 1, intercept = 40, linetype = "dotdash") +
#   geom_vline(xintercept = 0, color = "gray") +
#   geom_hline(yintercept = 0, color = "gray") +
#   geom_point(size = 3) +
#   geom_errorbar(
#     aes(ymin = d14c_corr_min_trt, 
#         ymax = d14c_corr_max_trt, 
#         color = Type), 
#     width = .25) +
#   geom_errorbarh(
#     aes(xmin = d14c_corr_min_ctl, 
#         xmax = d14c_corr_max_ctl, 
#         color = Type), 
#     height = .9) +
#   scale_color_manual(name = 'Land use',
#                      values =c('F'='#a35513','G'='#1361a3'),
#                      labels = c('Forest','Grassland')) +
#   xlab(expression('Control '*Delta*''^14*'C (‰)')) +
#   ylab(expression('Treatment '*Delta*''^14*'C (‰)')) +
#   theme_bw() +
#   theme(panel.grid = element_blank(),
#         legend.key.height=unit(.8, "cm"),
#         aspect.ratio = 1)
```


16. Show the effect of storage duration by plotting the difference between control and treatment ^14^C as a function of storage duration. However, the expected Ctl-Trt difference would also likely have a trend over time. Perhaps plot an expected trend...? Not sure how much this matters.
* As previously, dashed and dotted lines are plotted to give a sense of error introduced by the treatment by showing a difference of 20‰ and 40‰ (roughly equivalent to the atmospheric decline over five and 10 years respectively for the period 2000 to 2020)
* Position of points jittered to avoid overplotting; storage duration has been rounded down to the nearest whole year
* In general, the trend in the differences due to treatment for the highly enriched samples from Oak Ridge, TN suggest losses of the most recently fixed carbon over the duration of storage
    + These samples were included primarily because it was assumed that they would be more sensitive to potential losses of recently fixed carbon, as the label should only be present in this pool of soil C
    + However, in contrast, there does not seem to be evidence for a storage duration effect in the samples that only contain bomb-C
```{r}
# First add YearSampled and YearMeasured columns
all.14c.ctl.trt$YearSampled <- unlist(all.14c.sum.df[match(all.14c.ctl.trt$ID, all.14c.sum.df$ID), "YearSampled"])
# correct archive sample YearSampled
all.14c.ctl.trt[all.14c.ctl.trt$Experiment == "arc" & all.14c.ctl.trt$Site == "Hainich", "YearSampled"] <- 2011

# add year measured column, modify for tme.misc samples, rewet samples
all.14c.ctl.trt$YearMeasured <- 2018
all.14c.ctl.trt[all.14c.ctl.trt$Experiment == "rewet", "YearMeasured"] <- 2019
all.14c.ctl.trt[which(all.14c.ctl.trt$ID %in% tme.misc$ID), "YearMeasured"] <- 2009

# calculate storage duration numerically and as factor
all.14c.ctl.trt$tme <- unlist(all.14c.ctl.trt$YearMeasured - all.14c.ctl.trt$YearSampled)
all.14c.ctl.trt$tme.f <- factor(all.14c.ctl.trt$tme)

# add difference column
# (expressed as difference of means)
all.14c.ctl.trt$ctl.trt.d14 <- all.14c.ctl.trt$d14c_corr_mean_ctl - all.14c.ctl.trt$d14c_corr_mean_trt

# # add atm 14C for year sampled and calculate delta-delta
# all.14c.ctl.trt$atm.yr.sampled <- Datm[match(all.14c.ctl.trt$YearSampled, floor(Datm$Date)), "NHc14"]
# all.14c.ctl.trt$dd.ctl <- all.14c.ctl.trt$d14c_corr_mean_ctl - all.14c.ctl.trt$atm.yr.sampled
# all.14c.ctl.trt$dd.trt <- all.14c.ctl.trt$d14c_corr_mean_trt - all.14c.ctl.trt$atm.yr.sampled
# all.14c.ctl.trt$ctl.trt.dd14 <- all.14c.ctl.trt$dd.ctl - all.14c.ctl.trt$dd.trt

# Plot storage duration
all.14c.ctl.trt %>%
  filter(Horizon == "A") %>%
  filter(Experiment != "rewet") %>%
  select(Type, tme.f, Site, ctl.trt.d14, tme) %>%
  group_by(Type, tme.f, Site) %>%
  add_tally() %>%
  summarise_all(list(mean = mean, sd = sd), na.rm = TRUE) %>%
  mutate(dif_se = ctl.trt.d14_sd/n_mean,
         dif_u = ctl.trt.d14_mean + dif_se*2,
         dif_l = ctl.trt.d14_mean - dif_se*2) %>%
  mutate(Location = factor(recode(Site, !!!loc), exclude = TRUE)) %>%
  ggplot(., aes(tme_mean, ctl.trt.d14_mean, color = Location, shape = Type)) +
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = -20, linetype = "dashed") +
  geom_hline(yintercept = 20, linetype = "dashed") +
  geom_hline(yintercept = 40, linetype = "dotted") +
  geom_hline(yintercept = -40, linetype = "dotted") +
  geom_point(size = 3, position = position_dodge2(width = .7, preserve = "single")) +
  geom_errorbar(
    aes(ymin = dif_l,
        ymax = dif_u, 
        color = Location),
    width = .7,
    position = position_dodge2(width = .7, preserve = "single")) +
  scale_color_manual(values = c('Duke, NC' = '#01BB97',
                                'Sierra Nevada, CA' = '#FFC107',
                                'Central Germany' = 'black',
                                'Harvard, MA' = '#1E88E5',
                                'Oak Ridge, TN' = '#D81B60')) +
  scale_shape_manual(name = 'Land use',
                     values = c('F'= 17,'G' = 16),
                     labels = c('Forest','Grassland')) +
  ylab(expression('Control - Treatment '*Delta*''^14*'C (‰)')) +
  xlab("Storage duration (years)") +
  theme_bw() +
  theme(panel.grid = element_blank())
```

